---
title: "Biogeographic Provinces"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
---

```{r setup, message=F, warning=F}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',message = FALSE, warning = FALSE)

library(raster)
library(tidyverse)
library(sf)
library(rredlist)
library(furrr)
library(doParallel)
library(parallel)
library(foreach)
library(paletteer)

emLab_project_dir <- "~/Google Drive File Stream/Shared drives/emlab/projects/current-projects/ocean-conservation-priorities"

emLab_shrd_data_dir <- "~/Google Drive File Stream/Shared drives/emlab/data/"
```

# Create ocean and land masks

Here we use a 1km2 resolution raster in Mollweide projection provided by NCEAS. We resample this file to a lower resolution (50x50km) and also create a unprojected (WGS84) version. 

```{r, eval = F}
## Get 1km2 resolution ocean raster in Mollweide projection from NCEAS and resample to 50x50km

ocean <- raster(file.path(emLab_shrd_data_dir, "ohi-supplement", "ocean.tif"))

empty_lr_raster <- raster(crs = crs(ocean), 
                          ext = extent(ocean),
                          res = 50000)

ocean_low_res_moll <- projectRaster(ocean, empty_lr_raster, method = 'ngb', over = T)

#Remove cells with 100% land

land_50 <- rnaturalearth::ne_download(scale = 50, type = 'land', category = 'physical', returnclass = "sf")

land_50_moll <- land_50 %>% 
  st_transform(crs = projection(ocean_low_res_moll))

land_50_raster_moll <- rasterize(land_50_moll, ocean_low_res_moll, getCover = T)

ocean_low_res_moll <- ocean_low_res_moll %>% 
  mask(land_50_raster_moll, maskvalue = 1)

# make wgs84 versions

empty_wgs84_raster <- raster(crs = crs(land_50), 
                             ext = extent(land_50),
                             res = 0.5)

ocean_low_res_wgs84 <- projectRaster(ocean_low_res_moll, empty_wgs84_raster, method = "ngb", over = T)

land_50_raster_wgs84 <- rasterize(land_50, empty_wgs84_raster, getCover = T)

ocean_low_res_wgs84 <- ocean_low_res_wgs84 %>% 
  mask(land_50_raster_wgs84, maskvalue = 1) 
```

```{r, eval = F}
writeRaster(ocean_low_res_moll, 
            file.path(emLab_project_dir, "data","02_processed", "masks", "ocean_low_res_moll.tif"), overwrite = T)

writeRaster(ocean_low_res_wgs84, 
            file.path(emLab_project_dir, "data","02_processed", "masks", "ocean_low_res_wgs84.tif"), overwrite = T)

writeRaster(land_50_raster_wgs84, 
            file.path(emLab_project_dir, "data","02_processed", "masks", "land_low_res_wgs84.tif"), overwrite = T)

writeRaster(land_50_raster_moll, 
            file.path(emLab_project_dir, "data","02_processed", "masks", "land_low_res_moll.tif"), overwrite = T)

sf::st_write(land_50_moll,
             file.path(emLab_project_dir,"data", "02_processed", "masks", "land_50_moll.shp"))

sf::st_write(land_50,
             file.path(emLab_project_dir, "data", "02_processed", "masks", "land_50.shp"))
```

# Biogeographic provinces

## MEOW - PPOW

Marine ecoregions and pelagic provinces of the world (Spalding 2007, 2012)

```{r}
meow <- sf::st_read(file.path(emLab_shrd_data_dir, "marine-ecoregions-and-provinces", "01_Data", "WCMC-036-MEOW-PPOW-2007-2012-NoCoast.shp"))
```

```{r manual_fix}
# There is a regions that is not a valid polygons and need manual fixing

manual_fix <- sf::st_read(file.path(emLab_shrd_data_dir, "marine-ecoregions-and-provinces", "01_Data", "fixed_ecoregion.shp"))

meow <- meow %>% 
  filter(PROVINC != "North Central Pacific Gyre") 

meow <- rbind(meow, manual_fix %>% select(-fid))
```

```{r simplify_meows}
#simplify shapefile and make valid
meow_simpler <- rmapshaper::ms_simplify(meow, keep = 0.1)

valid <- meow_simpler[sf::st_is_valid(meow_simpler), ]

invalid <- meow_simpler[sf::st_is_valid(meow_simpler) == FALSE, ]

fixed_invalid <- invalid %>% 
  lwgeom::st_make_valid()

meow_simpler <- rbind(valid, 
                   fixed_invalid) 
```

### Group at provinces level

```{r group_meows}
#group at the province level
provinces <- meow_simpler %>% 
  dplyr::rename("PROVINCE" = PROVINC) %>% 
  group_by(PROVINCE, TYPE, BIOME, REALM) %>% 
  summarise() %>% 
  ungroup() %>% 
  mutate(ID = as.factor(row_number())) %>% 
  select(ID, everything())

provinces_moll <- provinces %>% 
  st_wrap_dateline(options = c("WRAPDATELINE=YES", "DATELINEOFFSET=180"), quiet = TRUE) %>% 
  st_transform("+proj=moll")
```

```{r }
color = grDevices::colors()[grep('gr(a|e)y', grDevices::colors(), invert = T)]

cols = sample(color, 99)

(provinces_plot <- provinces %>% 
  st_wrap_dateline(options = c("WRAPDATELINE=YES", "DATELINEOFFSET=180"), quiet = TRUE) %>% 
  st_transform("+proj=moll") %>%
  st_cast("MULTIPOLYGON") %>% 
  ggplot()+
  geom_sf(aes(fill = ID, label1 = REALM, label2= PROVINCE), col = 'transparent', show.legend = F) +
  scale_fill_manual(values = cols)+
  geom_sf(data = land_50_moll, fill = 'grey', col = 'grey', lwd = 0.1)+
  my_theme_map())

ggsave(provinces_plot, 
       filename = file.path(emLab_project_dir, "figures", "supplement", "coastal_and_pelagic_provinces.png"),
       dpi = 300,  height = 4, width = 8)
```

```{r, eval = F}
provinces %>% 
  st_set_geometry(NULL) %>% 
  group_by(TYPE) %>% 
  count() # 62 coastal provinces + 37 pelagic provinces
```

#### rasterize'em

```{r, eval = F}
# Calculate the number of cores
no_cores <- detectCores() - 1

# Split features in n parts
n <- nrow(provinces_moll)

# Initiate cluster (after loading all the necessary object to R environment
cl <- makeCluster(no_cores, type = "FORK")

print(cl)

clusterExport(cl, varlist = c("ocean_low_res_moll", "provinces_moll", "n"), envir=environment())

prov_rasters <- parLapply(cl = cl, X = 1:n, 
                          fun = function(x) mask(rasterize(provinces_moll[x,],
                                                           ocean_low_res_moll, getCover = T), 
                                                 ocean_low_res_moll))

stopCluster(cl)

# Merge all raster parts

provinces_stack <- stack(prov_rasters)

names(provinces_stack) <- provinces_moll$PROVINCE

write_rds(provinces_stack, path = file.path(emLab_project_dir, "data", "02_processed", "biogeography", "provinces_stack.rds"))
```

```{r}
names(provinces_stack) <- str_replace_all(names(provinces_stack), "\\.", "_")

unstacked_provs <- raster::unstack(provinces_stack) 

output_paths <- file.path(emLab_project_dir, "data", "02_processed", "biogeography", paste0(names(provinces_stack), ".tif"))

for(i in seq_along(unstacked_provs)){
  writeRaster(unstacked_provs[[i]], file = output_paths[i], overwrite=TRUE)
  }
```

## GOODS provinces 

### Abysal

```{r}
GOODS_abysal_provinces <- sf::st_read(file.path(emLab_shrd_data_dir, "bathyal-abyssal-biogeographic-provs", "GOODSprovinces_abyssal.shp"))

GOODS_abysal_dissolved <- GOODS_abysal_provinces %>% 
  st_wrap_dateline(options = c("WRAPDATELINE=YES", "DATELINEOFFSET=180"), quiet = TRUE) %>%
  group_by(Province) %>% 
  summarize(Name = unique(Name)) 

GOODS_abysal_dissolved <- GOODS_abysal_dissolved %>% 
  rmapshaper::ms_simplify(keep = 0.1) 

GOODS_abysal_map <- GOODS_abysal_dissolved %>% 
  st_wrap_dateline(options = c("WRAPDATELINE=YES", "DATELINEOFFSET=180"), quiet = TRUE) %>% 
  st_transform("+proj=moll") %>% 
  ggplot()+
  geom_sf(aes(fill = as.factor(Province)), col = 'transparent', show.legend = F) + 
  my_theme_map()+
  scale_fill_manual(values = cols)+
  geom_sf(data = land_50_moll, fill = 'grey', col = 'grey', lwd = 0.1)

ggsave(GOODS_abysal_map, 
       filename =  file.path(emLab_project_dir, "figures", "supplement", "abyssal_provinces.png"),
       dpi = 300,  height = 4, width = 8)
```

#### rasterize

```{r}
n_abysal_provs <- length(unique(GOODS_abysal_provinces$Province))

registerDoParallel() #register cores

foreach(p = seq(1:n_abysal_provs)) %dopar% { 
  
  GOODS_abysal_provinces %>% 
    filter(Province == p) %>% 
    st_wrap_dateline(options = c("WRAPDATELINE=YES", "DATELINEOFFSET=180"), quiet = TRUE) %>% 
    st_transform("+proj=moll") %>% 
    rasterize(ocean_low_res_moll,  getCover = T) %>% 
    mask(ocean_low_res_moll) %>% 
    raster::writeRaster(filename = paste0(file.path(emLab_project_dir, "data", "02_processed", "biogeography"), "/GOODS_abysal_provinces_", p ,".tif"),
                        overwrite=TRUE)
}
```

### Bathyal

```{r}
GOODS_bathyal_provinces <- sf::st_read(file.path(emLab_shrd_data_dir, "bathyal-abyssal-biogeographic-provs", "GOODSprovinces_bathyal.shp"))

GOODS_bathyal_dissolved <- GOODS_bathyal_provinces %>% 
  st_wrap_dateline(options = c("WRAPDATELINE=YES", "DATELINEOFFSET=180"), quiet = TRUE) %>%
  group_by(Province) %>% 
  summarize(Name = unique(Name)) %>% 
  rmapshaper::ms_simplify(keep = 0.1) 

bathyal_map <- GOODS_bathyal_dissolved %>% 
  st_wrap_dateline(options = c("WRAPDATELINE=YES", "DATELINEOFFSET=180"), quiet = TRUE) %>% 
  st_transform("+proj=moll") %>% 
  ggplot()+
  geom_sf(aes(fill = as.factor(Province)), col = 'transparent', show.legend = F) + 
  my_theme_map()+
  scale_fill_manual(values = cols)+
  geom_sf(data = land_50_moll, fill = 'grey', col = 'grey', lwd = 0.1)

ggsave(bathyal_map, 
       filename = file.path(emLab_project_dir, "figures", "supplement", "bathyal_provinces.png"),
       dpi = 300,  height = 4, width = 8)
```

#### rasterize

```{r}
n_bathyal_provs <- length(unique(GOODS_bathyal_provinces$Province))

registerDoParallel() #register cores

foreach(p = seq(1:n_bathyal_provs)) %dopar% { 

  GOODS_bathyal_provinces %>% 
  filter(Province == p) %>% 
  st_wrap_dateline(options = c("WRAPDATELINE=YES", "DATELINEOFFSET=180"), quiet = TRUE) %>% 
  st_transform("+proj=moll") %>% 
  rasterize(ocean_low_res_moll,  getCover = T) %>% 
  mask(ocean_low_res_moll) %>% 
  raster::writeRaster(filename = paste0(file.path(emLab_project_dir, "data", "02_processed", "biogeography"), "/GOODS_bathyal_provinces_", p ,".tif"),
                      overwrite=TRUE)}
```


