---
title: "Seamounts"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
---

```{r setup, message=F,warning=F}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',message = FALSE, warning = FALSE)

library(raster)
library(tidyverse)
library(sf)
library(rredlist)
library(furrr)
library(parallel)
library(foreach)
library(doParallel)

source(here::here("common.R"))
```

# Seamounts

## Clark et al

Clark et al classify the world's bathyal seamounts (depth < 3500) into 194 classes. 

```{r}
clark_seamounts <- readxl::read_xls(file.path(emLab_shrd_data_dir, "seamounts-clark", "Clarketal_2011_OCM_All_seamounts_values.xls"))

clark_seamounts <- clark_seamounts %>% 
  janitor::clean_names() %>% 
  select(id, long, lat, depth, class = smt_class_poc) %>% 
  filter(depth <= 3500) 

n_distinct(clark_seamounts$class) # 194
```

```{r}
clark_seamounts_sf_moll <- sf::st_as_sf(clark_seamounts, 
             coords = c("long", "lat"), 
             agr = "constant", 
             crs = 4326) %>% 
  st_transform(crs = projection(ocean_low_res_moll)) 
```

```{r}
color = grDevices::colors()[grep('gr(a|e)y', grDevices::colors(), invert = T)]

cols = sample(color, 194)
```

```{r}
ggplot()+
  geom_sf(data = land_50_moll, fill = 'grey', col = 'grey', lwd = 0.1)+
  geom_sf(data = clark_seamounts_sf_moll, 
          aes(col = factor(class), stroke = 0), 
          size = 0.7,
          show.legend = F)+
  my_theme_map()+
  scale_color_manual(values  = cols)+
  ggsave(filename = file.path(emLab_project_dir, "figures", "supplement", "seamounts_by_class.png"),
         height = 4, width = 8)
```

```{r}
clark_seamounts %>% 
  group_by(class) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n)) 
```

```{r}
clark_seamounts %>% 
  group_by(class) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n)) %>% 
  ggplot()+
  geom_histogram(aes(n))+
  theme_minimal()+
  ggsave(filename = file.path(emLab_project_dir, "figures", "exploratory", "seamounts_histogram.png"))
```

We can also use the RFMO data to classify them as fishable or non-fishable. Let's use the max depth across all RFMOs 2200m

```{r}
clark_seamounts <- clark_seamounts %>% 
  mutate(fishable = depth < 2200)
```

```{r}
clark_fishable_seamounts_sf_moll <- sf::st_as_sf(clark_seamounts %>% 
                                                   filter(fishable), 
                                                 coords = c("long", "lat"), 
                                                 agr = "constant", 
                                                 crs = 4326) %>% 
  st_transform(crs = projection(ocean_low_res_moll)) 
```

```{r}
ggplot()+
  geom_sf(data = land_50_moll, fill = 'black', col = 'black', lwd = 0.1)+
  geom_sf(data = clark_fishable_seamounts_sf_moll, aes(col = factor(class), stroke = 0), show.legend = F)+
  my_theme_map()+
  scale_color_manual(values  = cols)+
  labs(caption = "Clark's classification of global seamounts (depth < 2200m)")+
  ggsave(filename = file.path(emLab_project_dir, "figures", "exploratory", "seamounts_by_class_fishable.png"), height = 4, width = 8)
```

### rasterize'em

Let's rasterize these data and produce one raster for each class of seamount. 

```{r, eval = F}
registerDoParallel(10) #register cores

seamounts_classes <- unique(clark_seamounts$class)

foreach(seamount_class = seamounts_classes) %dopar% { 
  
  smt_data <- clark_seamounts_sf_moll %>% 
    filter(class == seamount_class) 
  
  smt_raster <- rasterize(smt_data, 
                          ocean_low_res_moll, field = 'id' ,fun = 'count')
  
  smt_raster <- mask(smt_raster, ocean_low_res_moll)
  
  raster::writeRaster(smt_raster, 
                      file.path(emLab_project_dir, "data", "02_processed", "seamounts/", paste0("smt_class_", seamount_class, ".tif")))
}
```

```{r, eval = F}
#registerDoParallel(10) #register cores
# 
# foreach(seamount_class = seamounts_classes) %dopar% { 
#   
#   smt_data <- clark_fishable_seamounts_sf_moll %>% 
#     filter(class == seamount_class) 
#   
#   smt_raster <- rasterize(smt_data, 
#                           ocean_low_res_moll, field = 'id' ,fun = 'count')
#   
#   smt_raster <- mask(smt_raster, ocean_low_res_moll)
#   
#   raster::writeRaster(smt_raster, 
#                       file.path(emLab_project_dir, "data", "02_processed", "seamounts/", paste0("fishable_smt_class_", seamount_class, ".tif")))
# }
```

