---
title: "Functional & Evolutionary Distinctiveness"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
---

```{r setup, message=F,warning=F}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',message = FALSE, warning = FALSE)

library(raster)
library(sf)
library(tidyverse)
library(rredlist)
library(furrr)
library(taxize)
library(treemap)

source(here::here("common.R"))
```

Let's examine the data compiled by David Moulliot of Functional and Evolutionary Distinctiveness of vertebrates

# Evolutionary Distinctiveness

David complied phylogenetic distinctiveness for bony fishes. It includes mean and SD values for 14,200 marine bony fish and 1192 sharks and rays. Uncertainty is due to the phylogenetic resolution that produces many Trees with the same likelihood.

## load datasets

```{r read data}
load(file.path(emLab_project_dir, "data", "01_raw", "Res_ED_isaac_fish"))

ED_bony_fish <- Res_ED_isaac %>% 
  rownames_to_column(var = "sci_name") %>% 
  rename("ED_mean" = "ED_isaac_mean",
         "ED_sd" = "ED_isaac_sd")

load(file.path(emLab_project_dir, "data", "01_raw", "Res_ED_shark"))

ED_shark <- Res_ED_shark %>% 
  rownames_to_column(var = "sci_name") %>% 
  rename("ED_mean" = "ED_shark_mean",
         "ED_sd" = "ED_shark_sd")

load(file.path(emLab_project_dir, "data", "01_raw", "Res_ED_isaac_corals"))

ED_corals <- Res_ED_isaac_corals %>% 
  rownames_to_column(var = "sci_name") %>% 
  rename("ED_mean" = "ED_isaac_mean",
         "ED_sd" = "ED_isaac_sd") %>% 
  mutate(sci_name = str_replace_all(sci_name, "_", " ")) %>% 
  mutate(sci_name = word(sci_name, 2,3))

load(file.path(emLab_project_dir, "data", "01_raw", "Res_ED_isaac_MMs"))

ED_mammals <- Res_ED_isaac_MMs %>% 
  rownames_to_column(var = "sci_name") %>% 
  rename("ED_mean" = "ED_isaac_mean",
         "ED_sd" = "ED_isaac_sd") %>% 
  mutate(sci_name = str_replace_all(sci_name, "_", " "))

load(file.path(emLab_project_dir, "data", "01_raw", "Res_ED_isaac_seabirds"))

ED_seabirds <- Res_ED_isaac_seabirds %>% 
  rownames_to_column(var = "sci_name") %>% 
  rename("ED_mean" = "ED_isaac_mean",
         "ED_sd" = "ED_isaac_sd") %>% 
  mutate(sci_name = str_replace_all(sci_name, "_", " "))
```

```{r}
ED_dataset <- bind_rows(ED_shark, ED_bony_fish, ED_corals, ED_mammals, ED_seabirds) %>% 
  mutate(sci_name = str_replace_all(sci_name, "_", " ")) %>% 
  arrange(sci_name)
```

### resolve scientific names

This fixes misspellings 

```{r}
sources <- c(worms = 9, fishbase = 155, eol = 12, col = 1, gbif = 11)

ED_resolved_names <- sources %>% 
  map(~ taxize::gnr_resolve(data_source_ids = .x, 
                            names = ED_dataset$sci_name,
                            best_match_only = T, 
                            fields = c("all"), 
                            canonical = T ))

ED_resolved_names_df <- ED_resolved_names %>% 
  map(~ .x %>% 
        filter(!match_value %in% c("Could only match genus") & str_count(matched_name2, "\\w+") >= 2) %>% 
        select(user_supplied_name, matched_name2, taxon_id)) %>% 
  reduce(full_join, by = "user_supplied_name") %>% 
  set_names(c("user_supplied_name", "worms_sci_name", "worms_id", "fishbase_sci_name", "fishbase_id", "eol_sci_name", "eol_id", 
              "col_sci_name", "col_id","gbif_sci_name", "gbif_id"))

ED_resolved_names <- ED_resolved_names_df %>% 
  mutate(resolved_scientific_name = case_when(
    !is.na(worms_sci_name) ~ worms_sci_name,
    !is.na(fishbase_sci_name) ~ fishbase_sci_name,
    !is.na(eol_sci_name) ~ eol_sci_name,
    !is.na(col_sci_name) ~ col_sci_name,
    TRUE ~ user_supplied_name
  )) %>% 
  select(user_supplied_name, resolved_scientific_name, everything()) 
```

```{r}
ED_dataset <- ED_dataset %>% 
  left_join(ED_resolved_names %>% 
              select(user_supplied_name, resolved_sci_name = resolved_scientific_name),
            by = c("sci_name" = "user_supplied_name"))

ED_dataset$resolved_sci_name[ED_dataset$sci_name == "Glaucostegus spinosus"] <- "Rhinobatos spinosus"
ED_dataset$resolved_sci_name[ED_dataset$sci_name == "Okamejei porosa"] <- "Okamejei kenojei"

ED_dataset <- ED_dataset %>%   
  mutate(resolved_sci_name = coalesce(resolved_sci_name, sci_name))
```

### validate names and get taxonomic info from WoRMS

```{r}
get_worms_info <- function(sci_name){
  
  worrms::wm_records_name(name = sci_name) %>% 
    mutate(provided_sci_name = sci_name) %>% 
    select(provided_sci_name, scientificname, status, valid_name, kingdom, phylum, class, order, family, genus)
  
}
```

```{r}
plan(multiprocess)

ED_worms_info <- ED_dataset %>%
  pull(resolved_sci_name) %>% 
  furrr::future_map(safely(get_worms_info)) %>% 
  transpose()

ED_worms_info <- ED_worms_info$result %>% 
  compact() %>% 
  bind_rows()
```

Lets take those entries that have no NAs, where the name queried matches the name supplied, and when there are only one valid name per entry

```{r}
valid_ED_worms_info <- ED_worms_info %>% 
  filter(!is.na(scientificname), !is.na(valid_name)) %>% # delete entries with NAs
  group_by(provided_sci_name) %>% 
  filter(provided_sci_name == scientificname) %>% # make sure to keep only rows that match the sci_name we supplied
  filter(n_distinct(valid_name, na.rm = T) == 1) %>%  # get those for which there is only one valid name
  ungroup() %>% 
  group_by(provided_sci_name) %>% 
  summarise_at(vars(valid_name, kingdom, phylum, class, order, family, genus), .funs = unique)
```

Now, from those with multiple valid names per row, lets get only the row for which status == "accepted", which means that the supplied name was already valid

```{r}
accepted_ED_worms_info <- ED_worms_info %>% 
  filter(!is.na(scientificname), !is.na(valid_name)) %>% # delete entries with NAs
  group_by(provided_sci_name) %>% 
  filter(provided_sci_name == scientificname) %>% # make sure to keep only rows that match the sci_name we supplied
  filter(n_distinct(valid_name, na.rm = T) > 1) %>% 
  filter(status == "accepted") %>% 
  group_by(provided_sci_name) %>% 
  summarise_at(vars(valid_name, kingdom, phylum, class, order, family, genus), .funs = unique)
```

```{r}
ED_dataset <- ED_dataset %>% 
  left_join(bind_rows(accepted_ED_worms_info,
                      valid_ED_worms_info),
            by = c("resolved_sci_name" = "provided_sci_name")
)

ED_dataset <- ED_dataset %>% 
  mutate(valid_sci_name = coalesce(valid_name, resolved_sci_name)) %>% 
  select(sci_name, resolved_sci_name,  valid_sci_name, everything(), -valid_name)
```

### fill taxonomic  gaps

```{r}
get_taxonomy <- function (sci_name, db) {
  
  taxize::classification(sci_name, db, accepted = T, ask = F)[[1]] %>% 
    select(-id) %>% 
    mutate(rank = stringr::str_to_lower(rank)) %>% 
    filter(rank %in% c("genus","family",  "order", "class","phylum","kingdom")) %>% 
    spread(rank, name) %>% 
    mutate(sci_name = sci_name) %>% 
    janitor::clean_names() %>% 
    select(sci_name, genus, family, order, class, phylum, kingdom)
  
}
```

```{r}
taxonomy_gaps_ED <- ED_dataset %>% 
  filter(is.na(kingdom)) %>% 
  select(valid_sci_name)
```

#### COL

```{r}
plan(multiprocess)

ED_taxonomy_col <- taxonomy_gaps_ED %>% 
  pull(valid_sci_name) %>% 
  furrr::future_map(safely(get_taxonomy), db = "col") %>% 
  transpose()

ED_taxonomy_col <- bind_rows(ED_taxonomy_col$result)
```

#### ITIS

```{r}
plan(multiprocess)

ED_taxonomy_itis <- taxonomy_gaps_ED %>% 
  filter(!valid_sci_name %in% ED_taxonomy_col$sci_name) %>% 
  pull(valid_sci_name) %>% 
  furrr::future_map(safely(get_taxonomy), db = "itis") %>% 
  transpose()

ED_taxonomy_itis <- bind_rows(ED_taxonomy_itis$result)
```

#### GBIF

```{r}
plan(multiprocess)

ED_taxonomy_gbif <- taxonomy_gaps_ED %>% 
  filter(!valid_sci_name %in% ED_taxonomy_col$sci_name) %>% 
  filter(!valid_sci_name %in% ED_taxonomy_itis$sci_name) %>% 
  pull(valid_sci_name) %>% 
  furrr::future_map(safely(get_taxonomy), db = "gbif") %>% 
  transpose()

ED_taxonomy_gbif <- bind_rows(ED_taxonomy_gbif$result)
```

```{r}
filled_taxonomy_gaps_ED <- bind_rows(ED_taxonomy_col, ED_taxonomy_itis, ED_taxonomy_gbif)

filled_taxonomy_gaps_ED <- filled_taxonomy_gaps_ED %>% 
  mutate(valid_sci_name = sci_name) %>% 
  group_by(valid_sci_name) %>% 
  summarise_at(vars(kingdom, phylum, class, order, family, genus), .funs = unique)
```

```{r}
ED_dataset <- ED_dataset %>% 
  left_join(filled_taxonomy_gaps_ED, by = c("valid_sci_name" = "valid_sci_name")) %>% 
  mutate(genus = coalesce(genus.x, genus.y),
         family = coalesce(family.x, family.y),
         order = coalesce(order.x, order.y),
         class = coalesce(class.x, class.y),
         phylum = coalesce(phylum.x, phylum.y),
         kingdom = coalesce(kingdom.x, kingdom.y)
         ) %>% 
  select(-contains(".x"), -contains(".y")) 
```


```{r}
ED_dataset %>% 
  group_by(class) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n))
```

Fix classes 

```{r}
ED_dataset <- ED_dataset %>% 
  mutate(class = case_when(
    class == "Holocephali" ~ "Chondrichthyes",
    class == "Elasmobranchii" ~ "Chondrichthyes",
    class == "Teleostei" ~ "Actinopterygii",
    TRUE  ~ class
  ))
```

```{r}
ED_dataset %>% 
  write_csv(file.path(emLab_project_dir, "data", "02_processed", "distinctiveness" ,"ED_dataset_with_taxonomy.csv"))
```

# Functional distinctiveness

David has also calculated Ecological Distinctiveness, for 3871 species. He used the function distinctiveness from the funrar package, which basically estimates the functional distance between each species and the rest of the species pool. The traits used for classification are: 

  - Loo (cm)	asymptotic maximum length
  - K (yr-1)	individual growth
  - Winfinity (g)	asymptotic maximum mass
  - tmax (yr)	maximum age
  - tm (yr)	age at maturity
  - M (yr-1)	natural mortality rate
  - Lm (cm)	length at maturity

## load datasets

```{r}
fish_FD <- read.table(file.path(emLab_project_dir, "data", "01_raw", "fishes_FD.txt"), header = T)

fish_FD <- fish_FD %>% 
  set_names(c("sci_name", "FD")) %>% 
  mutate(sci_name = str_replace(sci_name, "_", " ")) %>% 
  arrange(sci_name) %>% 
  select(sci_name, FD)
```

```{r}
corals_FD <- read.table(file.path(emLab_project_dir, "data", "01_raw", "corals_FD.txt")) %>% 
  rownames_to_column()

corals_FD <- corals_FD %>% 
  set_names(c("sci_name", "FD")) %>% 
  mutate(sci_name = as.character(sci_name))
```

```{r}
seabirds_FD <- read.table(file.path(emLab_project_dir, "data", "01_raw", "seabirds_FD.txt"), header = T) %>% 
  rownames_to_column()

seabirds_FD <- seabirds_FD %>% 
  set_names(c("sci_name", "FD")) %>% 
  mutate(sci_name = str_replace(sci_name, "_", " ")) 
```

```{r}
mammals_FD <- read.table(file.path(emLab_project_dir, "data", "01_raw", "mammals_FD.txt"), header = T)

mammals_FD <- mammals_FD %>% 
  set_names(c("sci_name", "FD")) %>% 
  mutate(sci_name = str_replace(sci_name, "_", " "))
```

```{r}
FD_dataset <- bind_rows(fish_FD, seabirds_FD, corals_FD, mammals_FD) %>% 
  filter(!is.na(FD))
```

### resolve scientific names

This fixes misspellings 

```{r}
sources <- c(worms = 9, fishbase = 155, eol = 12, col = 1, gbif = 11)

FD_resolved_names <- sources %>% 
  map(~ taxize::gnr_resolve(data_source_ids = .x, 
                            names = FD_dataset$sci_name,
                            best_match_only = T, 
                            fields = c("all"), 
                            canonical = T ))

FD_resolved_names_df <- FD_resolved_names %>% 
  map(~ .x %>% 
        filter(!match_value %in% c("Could only match genus") & str_count(matched_name2, "\\w+") >= 2) %>% 
        select(user_supplied_name, matched_name2, taxon_id)) %>% 
  reduce(full_join, by = "user_supplied_name") %>% 
  set_names(c("user_supplied_name", "worms_sci_name", "worms_id", "fishbase_sci_name", "fishbase_id", "eol_sci_name", "eol_id", "col_sci_name", "col_id","gbif_sci_name", "gbif_id"))

FD_resolved_names <- FD_resolved_names_df %>% 
  mutate(resolved_scientific_name = case_when(
    !is.na(worms_sci_name) ~ worms_sci_name,
    !is.na(fishbase_sci_name) ~ fishbase_sci_name,
    !is.na(eol_sci_name) ~ eol_sci_name,
    !is.na(col_sci_name) ~ col_sci_name,
    TRUE ~ user_supplied_name
  )) %>% 
  select(user_supplied_name, resolved_scientific_name, everything()) 
```

```{r}
FD_dataset <- FD_dataset %>% 
  left_join(FD_resolved_names %>% 
              select(user_supplied_name, resolved_sci_name = resolved_scientific_name),
            by = c("sci_name" = "user_supplied_name"))

FD_dataset <- FD_dataset %>%   
  mutate(resolved_sci_name = coalesce(resolved_sci_name, sci_name)) %>% 
  filter(!is.na(FD))
```

```{r}
FD_dataset$resolved_sci_name[FD_dataset$sci_name == "Pagrus auratus"] <- 'Chrysophrys auratus'
```

### validate names and get taxonomic info from WoRMS

```{r}
plan(multiprocess)

FD_worms_info <- FD_dataset %>%
  pull(resolved_sci_name) %>% 
  furrr::future_map(safely(get_worms_info)) %>% 
  transpose()

FD_worms_info <- FD_worms_info$result %>% 
  compact() %>% 
  bind_rows()
```

Lets take those entries that have no NAs, where the name queried matches the name supplied, and when there are only one valid name per entry

```{r}
valid_FD_worms_info <- FD_worms_info %>% 
  filter(!is.na(scientificname), !is.na(valid_name)) %>% # delete entries with NAs
  group_by(provided_sci_name) %>% 
  filter(provided_sci_name == scientificname) %>% # make sure to keep only rows that match the sci_name we supplied
  filter(n_distinct(valid_name, na.rm = T) == 1) %>%  # get those for which there is only one valid name
  ungroup() %>% 
  group_by(provided_sci_name) %>% 
  summarise_at(vars(valid_name, kingdom, phylum, class, order, family, genus), .funs = unique)
```

Now, from those with multiple valid names per row, lets get only the row for which status == "accepted", which means that the supplied name was already valid

```{r}
accepted_FD_worms_info <- FD_worms_info %>% 
  filter(!is.na(scientificname), !is.na(valid_name)) %>% # delete entries with NAs
  group_by(provided_sci_name) %>% 
  filter(provided_sci_name == scientificname) %>% # make sure to keep only rows that match the sci_name we supplied
  filter(n_distinct(valid_name, na.rm = T) > 1) %>% 
  filter(status == "accepted") %>% 
  group_by(provided_sci_name) %>% 
  summarise_at(vars(valid_name, kingdom, phylum, class, order, family, genus), .funs = unique)
```

```{r}
FD_dataset <- FD_dataset %>% 
  left_join(bind_rows(accepted_FD_worms_info,
                      valid_FD_worms_info),
            by = c("resolved_sci_name" = "provided_sci_name")
)

FD_dataset <- FD_dataset %>% 
  mutate(valid_sci_name = coalesce(valid_name, resolved_sci_name)) %>% 
  select(sci_name, resolved_sci_name,  valid_sci_name, everything(), -valid_name)
```

### fill taxonomic  gaps

```{r}
taxonomy_gaps_FD <- FD_dataset %>% 
  filter(is.na(kingdom)) %>% 
  select(valid_sci_name)
```

#### COL

```{r}
plan(multiprocess)

FD_taxonomy_col <- taxonomy_gaps_FD %>% 
  pull(valid_sci_name) %>% 
  furrr::future_map(safely(get_taxonomy), db = "col") %>% 
  transpose()

FD_taxonomy_col <- bind_rows(FD_taxonomy_col$result)
```

#### ITIS

```{r}
plan(multiprocess)

FD_taxonomy_itis <- taxonomy_gaps_FD %>% 
  filter(!valid_sci_name %in% FD_taxonomy_col$sci_name) %>% 
  pull(valid_sci_name) %>% 
  furrr::future_map(safely(get_taxonomy), db = "itis") %>% 
  transpose()

FD_taxonomy_itis <- bind_rows(FD_taxonomy_itis$result)
```

#### GBIF

```{r}
plan(multiprocess)

FD_taxonomy_gbif <- taxonomy_gaps_FD %>% 
  filter(!valid_sci_name %in% FD_taxonomy_col$sci_name) %>% 
  filter(!valid_sci_name %in% FD_taxonomy_itis$sci_name) %>% 
  pull(valid_sci_name) %>% 
  furrr::future_map(safely(get_taxonomy), db = "gbif") %>% 
  transpose()

FD_taxonomy_gbif <- bind_rows(FD_taxonomy_gbif$result)
```

```{r}
filled_taxonomy_gaps_FD <- bind_rows(FD_taxonomy_col, FD_taxonomy_itis, FD_taxonomy_gbif)

filled_taxonomy_gaps_FD <- filled_taxonomy_gaps_FD %>% 
  mutate(valid_sci_name = sci_name) %>% 
  group_by(valid_sci_name) %>% 
  summarise_at(vars(kingdom, phylum, class, order, family, genus), .funs = unique)
```

```{r}
FD_dataset <- FD_dataset %>% 
  left_join(filled_taxonomy_gaps_FD, by = c("valid_sci_name" = "valid_sci_name")) %>% 
  mutate(genus = coalesce(genus.x, genus.y),
         family = coalesce(family.x, family.y),
         order = coalesce(order.x, order.y),
         class = coalesce(class.x, class.y),
         phylum = coalesce(phylum.x, phylum.y),
         kingdom = coalesce(kingdom.x, kingdom.y)
         ) %>% 
  select(-contains(".x"), -contains(".y")) 
```

```{r}
FD_dataset %>% 
  group_by(class) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n))
```

```{r}
FD_dataset %>% 
  write_csv(file.path(emLab_project_dir, "data", "02_processed", "distinctiveness" ,"FD_dataset_with_taxonomy.csv"))
```

# Coverage

```{r}
FD_dataset <- read_csv(file.path(emLab_project_dir, "data", "02_processed", "distinctiveness" ,"FD_dataset_with_taxonomy.csv"))

ED_dataset <- read_csv(file.path(emLab_project_dir, "data", "02_processed", "distinctiveness" ,"ED_dataset_with_taxonomy.csv"))

included_aquamaps_spp_info <- read_csv(file.path(emLab_project_dir, "data", "02_processed", "species_list","included_aquamaps_spp_list.csv"))

included_seabirds_spp_info <-  read_csv(file.path(emLab_project_dir, "data", "02_processed", "species_list","included_birdlife_spp_list.csv"))
```

```{r}
full_included_spp_list <- bind_rows(included_aquamaps_spp_info %>% 
                                      select(valid_sci_name, genus, order, family, class, phylum, kingdom, iucn_category, file_id = aquamaps_id),
                                    included_seabirds_spp_info %>% 
                                      mutate(birdlife_id = as.character(birdlife_id)) %>% 
                                      select(valid_sci_name, genus, order, family, class, phylum, kingdom, iucn_category, file_id = birdlife_id))
```

##  Join ED info

```{r}
full_included_spp_list <- full_included_spp_list %>% 
  left_join(ED_dataset %>% 
              group_by(valid_sci_name) %>% 
              filter(n_distinct(ED_mean) == 1) %>% 
              ungroup() %>% 
              distinct(valid_sci_name, ED_mean, ED_sd), by = "valid_sci_name")

ED_dupes <- ED_dataset %>% 
  group_by(valid_sci_name) %>% 
  filter(n_distinct(ED_mean) > 1) %>% 
  ungroup() %>% 
  filter(sci_name == valid_sci_name) %>% 
  distinct(valid_sci_name, ED_mean, ED_sd)

full_included_spp_list <- full_included_spp_list %>% 
  left_join(ED_dupes, 
            by = "valid_sci_name") %>% 
  mutate(ED_mean = coalesce(ED_mean.x, ED_mean.y),
         ED_sd = coalesce(ED_sd.x, ED_sd.y)) %>% 
  select(-contains(".x"), -contains(".y"))
```

##  Join FD info

```{r}
full_included_spp_list <- full_included_spp_list %>% 
  left_join(FD_dataset %>% 
              group_by(valid_sci_name) %>% 
              filter(n_distinct(FD) == 1) %>% 
              ungroup() %>% 
              distinct(valid_sci_name, FD))

FD_dupes <- FD_dataset %>% 
  group_by(valid_sci_name) %>% 
  filter(n_distinct(FD) > 1) %>% 
  ungroup() %>% 
  filter(sci_name == valid_sci_name) %>% 
  distinct(valid_sci_name, FD)

full_included_spp_list <- full_included_spp_list %>% 
  left_join(FD_dupes, 
            by = "valid_sci_name") %>% 
  mutate(FD = coalesce(FD.x, FD.y)) %>% 
  select(-contains(".x"), -contains(".y"))
```

Of included species, what % have FD and ED info?

```{r}
full_included_spp_list %>%
  summarize(n_spp = n(),
            spp_with_ED = sum(!is.na(ED_mean))/n_spp,
            spp_with_FD = sum(!is.na(FD))/n_spp)
```

Of vertebrates, what % have FD and ED info?

```{r}
full_included_spp_list %>% 
  filter(phylum == "Chordata") %>% 
  summarize(n_spp = n(),
            spp_with_ED = sum(!is.na(ED_mean))/n_spp,
            spp_with_FD = sum(!is.na(FD))/n_spp)
```

Of vertebrates, what is our covergae by class?

```{r}
full_included_spp_list %>% 
  filter(phylum == "Chordata") %>% 
  group_by(class) %>% 
  summarize(n_spp = n(),
            spp_wit_ED = sum(!is.na(ED_mean))/n_spp,
            spp_wit_FD = sum(!is.na(FD))/n_spp) %>% 
  arrange(desc(n_spp))
```

For which groups are we missing the most?

```{r}
full_included_spp_list %>%
  group_by(class) %>% 
  summarize(n_spp = n(),
            fraction_spp_sans_ED = sum(is.na(ED_mean))/n_spp,
            fraction_spp_sans_FD = sum(is.na(FD))/n_spp) %>% 
  arrange(desc(n_spp)) 
```


```{r}
full_included_spp_list %>%
  filter(phylum != "Chordata") %>% 
  group_by(class) %>% 
  summarize(n_spp = n(),
            fraction_spp_sans_ED = sum(is.na(ED_mean))/n_spp,
            fraction_spp_sans_FD = sum(is.na(FD))/n_spp) %>% 
  arrange(desc(n_spp))
```

# Distributions

```{r}
tmp <- full_included_spp_list %>% 
  filter(!is.na(FD)| !is.na(ED_mean)) %>% 
  group_by(class) %>% 
  summarise(n = n()) %>% 
  filter(n > 1, !is.na(class)) %>% 
  arrange(desc(n))

cols <- c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#fdbf6f','#ff7f00','#cab2d6','#6a3d9a','#ffff99')
cols <- c('#7F3C8D','#11A579','#3969AC','#F2B701','#E73F74','#80BA5A','#E68310','#008695','#CF1C90','#f97b72','#4b4b8f')
names(cols) <- tmp$class
```


```{r}
full_included_spp_list %>% 
  filter(!is.na(class)) %>% 
  ggplot()+
  geom_density(aes(x = ED_mean, col = class))+
  scale_color_manual(values = cols)+
  theme_minimal()+
  labs(title = "Evolutionary distinctiveness by class")+
  ggsave(filename = file.path(emLab_project_dir, "figures", "exploratory", "ED_distributions_by_taxa.png"),
          dpi = 300, width = 8, height = 4)
```

```{r}
full_included_spp_list %>% 
  ggplot()+
  geom_density(aes(x = FD, col = class))+
  scale_color_manual(values = cols)+
  theme_minimal()+
  labs(title = "Functional distinctiveness by class")+
  ggsave(filename = file.path(emLab_project_dir, "figures", "exploratory", "FD_distributions_by_taxa.png"),
         dpi = 300, width = 8, height = 4)
```

```{r}
full_included_spp_list %>% 
  ggplot()+
  geom_density(aes(x = FD, col = class))+
  scale_color_manual(values = cols)+
  theme_minimal()+
  scale_x_log10()+
  labs(title = "Functional distinctiveness by class")+
  ggsave(filename = file.path(emLab_project_dir, "figures", "exploratory", "FD_distributions_by_taxa_log.png"),
         dpi = 300, width = 8, height = 4)
```

# Fill in gaps

```{r}
full_included_spp_list_gf <- full_included_spp_list %>% 
  mutate(class = case_when(
    class == "Holocephali" ~ "Chondrichthyes",
    class == "Elasmobranchii" ~ "Chondrichthyes",
    class == "Teleostei" ~ "Actinopterygii",
    TRUE  ~ class
  )) %>%  
  group_by(class) %>% 
  mutate(class_ED = mean(ED_mean, na.rm = T),
         class_FD = mean(FD, na.rm = T)) %>% 
  mutate(ED = coalesce(ED_mean, class_ED),
         FD = coalesce(FD, class_FD)) %>% 
  ungroup() %>% 
  mutate(ED = if_else(is.na(ED), mean(ED, na.rm = T), ED),
         FD = if_else(is.na(FD), mean(FD, na.rm = T), FD)) %>% 
  select(-ED_mean, -ED_sd, -class_ED, -class_FD)
```

# Rescale 

```{r}
full_included_spp_list_gf %>% 
  arrange(desc(FD))
```

```{r}
full_included_spp_list_gf <- full_included_spp_list_gf %>% 
  ungroup() %>% 
  mutate(ED = scales::rescale(ED, to = c(0,1)),
         FD = scales::rescale(FD, to = c(0,1)))
```

## Export files

```{r}
included_aquamaps_spp_info %>% 
  left_join(full_included_spp_list_gf %>% 
              select(valid_sci_name, ED, FD),
            by =  "valid_sci_name") %>% 
  write_csv(file.path(emLab_project_dir, "data", "02_processed", "species_list", "included_aquamaps_spp_list_with_ED_FD.csv"))
```

```{r}
included_seabirds_spp_info %>% 
  left_join(full_included_spp_list_gf %>% 
              select(valid_sci_name, ED, FD), 
            by = "valid_sci_name")  %>% 
    write_csv(file.path(emLab_project_dir, "data", "02_processed", "species_list", "included_birdlife_spp_list_with_ED_FD.csv"))
```

# Visualize Weights

```{r}
included_aquamaps_spp_info <- read_csv(file.path(emLab_project_dir, "data", "02_processed", "species_list", "included_aquamaps_spp_list_with_ED_FD.csv"))

included_seabirds_spp_info <- read_csv(file.path(emLab_project_dir, "data", "02_processed", "species_list", "included_birdlife_spp_list_with_ED_FD.csv"))

full_included_spp_list <- bind_rows(included_aquamaps_spp_info %>% 
                                      select(valid_sci_name, genus, order, family, class, phylum, kingdom, iucn_category, ED, FD),
                                    included_seabirds_spp_info %>% 
                                      select(valid_sci_name, genus, order, family, class, phylum, kingdom, iucn_category,  ED, FD))
```

```{r}
full_included_spp_list <- full_included_spp_list %>% 
  mutate(iucn_weight = case_when(
    iucn_category == "CR" ~ 1,
    iucn_category == "EN" ~ .75,
    iucn_category == "VU" ~ .5, 
    iucn_category == "NT" ~ .25, 
    iucn_category == "DD" ~ .25,
    iucn_category == "LC" ~ .125,
    is.na(iucn_category) ~ .25
  )) %>% 
  mutate(full_weight = iucn_weight*(ED + FD))
```

```{r}
full_included_spp_list %>% 
  arrange(desc(full_weight))
```

```{r}
n <- n_distinct(full_included_spp_list$class)

col_25 <- c("#396735",
           "#ba45de",
           "#57b946",
           "#5b33c4",
           "#9eb03b",
           "#8369e1",
           "#d3953b",
           "#543492",
           "#db4f2b",
           "#6681d5",
           "#954c27",
           "#d13fb5",
           "#64b58d",
           "#d94a82",
           "#58a6c3",
           "#cc4450",
           "#3f5883",
           "#a09859",
           "#d47ac9",
           "#605126",
           "#933277",
           "#d38b86",
           "#613766",
           "#b290c2",
           "#773039")
```

```{r}
png(filename =  file.path(emLab_project_dir, "figures","supplement", "included_spp_treemap.png"),
    width = 800, height=800)
full_included_spp_list %>% 
  filter(!is.na(class)) %>% 
  group_by(class) %>% 
  summarise(weight = n()) %>% 
  mutate(label = paste(class, weight, sep = "\n")) %>% 
  treemap(
    index = c("label"),
    vSize = "weight",
    type = "index",
    title = "",
    palette = col_25,
    fontsize.labels = c(12)
  ) 
dev.off()
```

```{r}
png(filename =  file.path(emLab_project_dir, "figures","exploratory", "included_spp_treemap_iucn_weights.png"),
    width = 800, height = 800)
full_included_spp_list %>% 
  filter(!is.na(class)) %>% 
  group_by(class) %>% 
  summarise(weight = sum(iucn_weight)) %>% 
  mutate(f_weight = paste(round(100*weight/sum(weight), 1), "%", sep = " "),
         label = paste(class, f_weight, sep = "\n")) %>% 
  treemap(
    index = c("label"),
    vSize = "weight",
    type = "index",
    title = "",
    palette = col_25,
    fontsize.labels=c(12)
  ) 
dev.off()
```

```{r}
full_included_spp_list %>% 
  filter(!is.na(class)) %>% 
  group_by(class) %>% 
  summarise(weight = sum(iucn_weight)) %>% 
  mutate(f_weight = round(100*weight/sum(weight)))
```


```{r}
png(filename = file.path(emLab_project_dir, "figures","supplement", "included_spp_treemap_full_weights.png"),
    width = 800, height=800)
full_included_spp_list %>% 
  filter(!is.na(class)) %>% 
  group_by(class) %>% 
  summarise(weight = sum(full_weight)) %>% 
  mutate(f_weight = paste(round(100*weight/sum(weight), 1), "%", sep = " "),
         label = paste(class, f_weight, sep = "\n")) %>% 
  treemap(
    index = c("label"),
    vSize = "weight",
    type = "index",
    title = "",
    palette = col_25,
    fontsize.labels=c(12)
  ) 
dev.off()
```

