---
title: "Blue Carbon Stocks"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
---

```{r setup, message=F,warning=F}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',message = FALSE, warning = FALSE)

library(raster)
library(tidyverse)
library(sf)
library(rredlist)
library(furrr)

source(here::here("common.R"))
```

# Read data 

```{r}
carbon_raster <- raster(file.path(emLab_shrd_data_dir, "global-marine-sedimentary-carbon", "Mean carbon_stock.tif"))

total_carbon <- raster::cellStats(carbon_raster, sum)/10^6 # 3117263 millions of Tons in the raw data

carbon_map_high_res <- tmap::tm_shape(land_50)+
  tmap::tm_fill(col = "black")+
  tmap::tm_shape(log(carbon_raster))+
  tmap::tm_raster(title =  expression(paste("Mg C ") * km^2),
                  palette  = viridis::plasma(n = 10),
                  style = 'cont',
                  legend.is.portrait = T, legend.reverse = T)+
  tmap::tm_layout(title = "",
                  title.position = c("center", .95),
                  inner.margins = c(0.12, 0, 0.08, 0.1), 
                  frame = F,
                  legend.position = c(.92, "center"))

tmap::tmap_save(carbon_map_high_res, 
                filename = file.path(emLab_project_dir, "figures", "exploratory", "carbon_stocks_high_res.png"),
                dpi = 300,  height = 4, width = 8)
```


```{r, eval = F}
carbon_hr <- raster(file.path(emLab_shrd_data_dir, "global-marine-sedimentary-carbon", "Mean carbon_stock.tif"))

area_hr <- area(carbon_hr)

plot(area_hr)

total_carbon_hr <- cellStats(area_hr*carbon_hr, sum) # 2.32 x10^12

total_carbon_hr/10^12

carbon_hr <- carbon_hr %>% 
  projectRaster(ocean_hr, "bilinear") %>% 
  mask(ocean_hr)

raster::writeRaster(carbon_hr, 
                    file.path(emLab_project_dir, 
                              "data", "02_processed", "impacts", "bottom_trawling", "carbon_hr.tif"), overwrite = T)
```

# Resample

How can we aggregate this high res raster?

One approach is to aggregate the resolution to an equivalent 0.5ยบ degrees, an summing all carbon within each pixel. 

```{r}
agg_fact <- round(res(ocean_low_res_wgs84)[1]/res(carbon_raster)[1])

carbon_raster_lr_sum <- aggregate(carbon_raster, 
                                  fact = agg_fact, 
                                  fun = sum, na.rm = T) %>% 
  projectRaster(ocean_low_res_wgs84) %>% 
  mask(ocean_low_res_wgs84)

raster::cellStats(carbon_raster_lr_sum, sum)/10^6 # 3117263 millions of Tons, same as in the raw data

carbon_raster_lr_sum %>% 
  raster::writeRaster(file.path(emLab_project_dir, "data", "02_processed","carbon","carbon_lr_sum_wgs84.tif"), 
                      overwrite = T)
```

```{r}
carbon_map_lr_sum <- tmap::tm_shape(land_50)+
  tmap::tm_fill(col = "black")+
  tmap::tm_shape(carbon_raster_lr_sum)+
  tmap::tm_raster(title = "Mg Carbon @ 1m depth",
                  palette  = viridis::plasma(n = 10),
                  style = 'cont', legend.is.portrait = T, legend.reverse = T)+
  tmap::tm_layout(title = "",
                  title.position = c("center", "top"),
                  frame = FALSE,
                  legend.position = c("right","bottom"),
                  legend.width = 10)

tmap::tmap_save(carbon_map_lr_sum, 
                filename = file.path(emLab_project_dir, "figures", "exploratory", "carbon_stocks_low_res_sum.png"),
                dpi = 300,  height = 4, width = 8)
```

However, this approach results in a wrong pattern of lower total carbon along the coastline because the function sums too many land values that have NA. 

If instead we take the mean of a cell we retain the spatial patterns

```{r}
carbon_raster_lr_mean <- aggregate(carbon_raster, 
                                   fact = agg_fact, 
                                   fun = mean,
                                   na.rm = T) %>% 
  projectRaster(ocean_low_res_wgs84) %>% 
  mask(ocean_low_res_wgs84)

carbon_map_lr_mean <- tmap::tm_shape(land_50)+
  tmap::tm_fill(col = "black")+
  tmap::tm_shape(carbon_raster_lr_mean)+
  tmap::tm_raster(title = expression(paste("Avg Mg C ") * km^2),
                  palette  = viridis::plasma(n = 10),
                  style = 'cont', legend.is.portrait = T, legend.reverse = T)+
  tmap::tm_layout(title = "",
                  title.position = c("center", "top"),
                  frame = FALSE,
                  legend.position = c("right","bottom"),
                  legend.width = 10)

tmap::tmap_save(carbon_map_lr_mean, 
                filename = file.path(emLab_project_dir, "figures", "exploratory", "carbon_stocks_low_res_mean.png"),
                dpi = 300,  height = 4, width = 8)

carbon_raster_lr_mean %>% 
  raster::writeRaster(file.path(emLab_project_dir, "data", "02_processed","carbon","carbon_lr_mean_wgs84.tif"), 
                      overwrite = T)

raster::cellStats(carbon_raster_lr_mean, sum)/10^6 
```

The patterns seen here appear to be correct. How much carbon does the map shows? To estimate this we need to multiply the average carbon in each cell time its area. But, we need to reproject the raster in the first place in order to get an accurate estimate of each pixels area.

```{r}
raster::cellStats(raster::area(carbon_raster_lr_mean)*carbon_raster_lr_mean, sum)/10^6/total_carbon 
```

```{r}
# reprojecting to mollweide

carbon_raster_lr_mean_moll <- projectRaster(carbon_raster_lr_mean, 
                                            ocean_low_res_moll, method = 'bilinear') %>% 
  mask(ocean_low_res_moll) 

raster::cellStats(raster::area(carbon_raster_lr_mean_moll)/10^6*carbon_raster_lr_mean_moll, sum)/10^6/total_carbon # 2615847 million tons of carbon

carbon_raster_lr_mean_moll %>% 
  mask(ocean_low_res_moll) %>% 
  raster::writeRaster(file.path(emLab_project_dir, "data", "02_processed","carbon","carbon_lr_mean_moll.tif"), 
                      overwrite = T)

plot((carbon_raster_lr_mean_moll))
```

```{r}
carbon_low_res <- raster::raster(file.path(emLab_project_dir, "data", "02_processed","carbon","carbon_lr_mean_moll.tif"))

carbon_low_res %>% 
  as.data.frame(xy = T) %>% 
  ggplot()+
  geom_raster(aes(x, y, fill = carbon_lr_mean_moll), show.legend = T)+
  theme_void()+
  geom_sf(data = land_50_moll, fill = 'grey', col = 'grey', lwd = 0.1)+
  scale_fill_gradientn(colors  = pals::viridis(n = 30),
                       na.value = "white",
                       trans = "log10")+
  labs(fill = "Avg Mg Carbon per km2\n @ 1m depth") +
  ggsave(filename = file.path(emLab_project_dir, "figures", "supplement", "mean_carbon_stocks.png"), 
         dpi = 300,  height = 4, width = 8)
```






