---
title: "Estimating bottom trawling footprint"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
editor_options:
  chunk_output_type: inline
---

```{r setup}
library(raster)
library(tidyverse)
library(DBI)
library(bigrquery)
library(sf)

source(here::here("common.R"))
```

```{r}
BQ_connection <-  dbConnect(bigquery(), 
                            project = 'emlab-gcp',
                            dataset = "carbon_trawl",
                            billing = "emlab-gcp", # your billing account name
                            use_legacy_sql = FALSE) # specify we are using Standard SQL

knitr::opts_chunk$set(connection = "BQ_connection")
```

# Vessels of interest

```{sql, connection = "BQ_connection", output.var = "gfw_vessels", eval = F}
SELECT
  year,
  ssvid,
  ais_identity.n_shipname_mostcommon.value AS ship_name,
  registry_info.best_known_imo AS imo,
  registry_info.best_known_callsign AS callsign,
  best.best_flag AS flag,
  best.best_length_m AS length,
  best.best_engine_power_kw AS engine_power,
  best.best_vessel_class AS vessel_class,
  activity.active_hours AS active_hours,
  activity.fishing_hours AS fishing_hours,
  activity.avg_depth_fishing_m AS avg_fishing_depth
FROM
  `world-fishing-827.gfw_research.vi_ssvid_byyear_v20200312` AS vessels
INNER JOIN
  `world-fishing-827.gfw_research.fishing_vessels_ssvid_v20200312` AS FILTER
USING
  (year, ssvid)
WHERE
  best.best_vessel_class IN ("trawlers", "dredge_fishing")
  AND year > 2015
  AND year < 2020
```

# Refine gear type

## EU list

```{sql, connection = "BQ_connection", output.var = "EU_gear_types", eval = F}
SELECT
  DISTINCT CAST(mmsi AS string)ssvid,
  shipname_matched,
  callsign_matched,
  imo_matched,
  registry.geartype_original,
  regexp_extract(registry.geartype_original,"^[^-]*[^ -]") as main_gear,  
  regexp_extract(registry.geartype_original,"[^ -][^-]*$") as sec_gear
FROM
  `world-fishing-827.vessel_database_staging.matched_eu_v20180613`,
  UNNEST(registry) AS registry
WHERE
  geartype_original IS NOT NULL
  AND mmsi IS NOT NULL
  and matched
  AND REGEXP_CONTAINS(geartype_original, "DRB|DTS|HMD|OTB|OTT|PTB|TBB|OTM|PTM|TM")
ORDER BY
  ssvid
```

## CCAMLR

```{sql, connection = "BQ_connection", output.var = "CCAMLR_OTM", eval = F}
SELECT
  DISTINCT CAST(mmsi AS string)ssvid,
  shipname_matched,
  callsign_matched,
  imo_matched,
  registry.geartype_original
FROM
  `world-fishing-827.vessel_database_staging.matched_ccamlr_v20200301` ,
  UNNEST(registry) AS registry
WHERE
 REGEXP_CONTAINS(geartype_original, "MIDWATER")
```

```{r, eval = F}
write_rds(gfw_vessels, file.path(emLab_project_dir, "data", "02_processed","impacts", "bottom_trawling", "gfw_vessels.rds"))
write_rds(CCAMLR_OTM, file.path(emLab_project_dir, "data", "02_processed","impacts", "bottom_trawling", "gfw_CCAMLR_vessels.rds"))
write_rds(EU_gear_types, file.path(emLab_project_dir, "data", "02_processed","impacts", "bottom_trawling", "gfw_EU_vessels.rds"))
```

## join lists

```{r}
gfw_vessels <- read_rds(file.path(emLab_project_dir, "data", "02_processed","impacts", "bottom_trawling", "gfw_vessels.rds"))
CCAMLR_OTM <- read_rds(file.path(emLab_project_dir, "data", "02_processed","impacts", "bottom_trawling", "gfw_CCAMLR_vessels.rds"))
EU_gear_types <- read_rds(file.path(emLab_project_dir, "data", "02_processed","impacts", "bottom_trawling", "gfw_EU_vessels.rds"))
```

```{r}
EU_gear_group <- EU_gear_types %>% 
  mutate(gear_group = case_when(
    main_gear %in% c("TM", "OTM", "PTM") ~ "OTM",
    main_gear %in% c("HMD")  ~ "HD",
    main_gear %in% c("DRB") ~ "TD",
    main_gear %in% c("TBB")  ~ "BT", 
    main_gear %in% c("OTB", "OTT", "DTS", "PTB") ~ "OT",
    sec_gear %in% c("TM", "OTM", "PTM") ~ "OTM",
    sec_gear %in% c("HMD")  ~ "HD",
    sec_gear %in% c("DRB") ~ "TD",
    sec_gear %in% c("TBB")  ~ "BT", 
    sec_gear %in% c("OTB", "OTT", "DTS", "PTB") ~ "OT"
  )) %>%
  group_by(ssvid) %>% 
  filter(n_distinct(gear_group) == 1) %>%
  distinct(ssvid, gear_group)
```

If there is no gear group we assume they are otter trawls as they are the most common. We also remove all midwater trawlers. 

```{r}
vessels_of_interest <- gfw_vessels %>% 
  left_join(EU_gear_group) %>% 
  mutate(gear_group = case_when(
    !is.na(gear_group) ~ gear_group,
    vessel_class == "trawlers" ~ "OT",
    vessel_class == "dredge_fishing" ~ "TD"
  )) %>% 
  filter(!ssvid %in% CCAMLR_OTM$ssvid) %>%
  filter(gear_group != "OTM") 

vessels_of_interest %>% 
  group_by(gear_group) %>% 
  summarize(n_ssvid = n_distinct(ssvid))
```

# Gear width

Using the estimates from [Eigaard et al](https://academic.oup.com/icesjms/article/73/suppl_1/i27/2573989#supplementary-data) we can estimate gear width as:

  - Dredgers: operate at 2-2.5 kts, very shallow, and use 0.72-3m width dredges. 

$$W = 0.3142*LOA^{1.2454}$$
  - Otter trawl (OT): operate at 2-4 kts, 25-250 m between doors width, between 10-2500 m deep. Equation from OT_MIX group,  representative of most number of species 

$$W = 10.6608*KW^{0.2921}$$
  - Beam trawl (BT): operate usually two beam with a total widht between 4-12m, speed between 2.5-7 knts,  and shallower than 100m. Equation from TBB_DMF group, representative of most number of species 

$$W = 0.6601*KW^{0.5078}$$

```{r}
vessels_of_interest <- vessels_of_interest %>% 
  mutate(gear_width = case_when(
    gear_group == "BT" ~ 0.6601*engine_power^(0.5078),
    gear_group == "OT" ~ 10.6608*engine_power^(0.2921),
    gear_group %in% c("TD", "HD") ~ 0.3142*length^(1.2454)
  )) 

vessels_of_interest %>% 
  ggplot()+
  geom_density(aes(gear_width,  fill = gear_group))+
  ggsave(file.path(emLab_project_dir, "figures", "exploratory", "bottom_trawling_gear_width_distributions.png"), 
         dpi = 300, height = 4, width = 8)
```

```{r, eval = F}
DBI::dbWriteTable(conn = BQ_connection,
                  name = "ssvid_of_interest",
                  value = vessels_of_interest[c("year", "ssvid", "gear_group", "gear_width")],
                  fields = as_bq_fields(vessels_of_interest[c("year", "ssvid", "gear_group", "gear_width")]),
                  overwrite = F
                  )
```

# Effort

```{sql, connection = "BQ_connection", output.var = "binned_effort_by_ssvid", eval = F}
WITH
  --------------------------------------------------------
  -- Use a slightly more agressive filter for good segments
  --------------------------------------------------------
  good_segments AS (
  SELECT
    seg_id
  FROM
    `world-fishing-827.gfw_research.pipe_production_v20190502_segs` 
  WHERE
    good_seg
    AND positions > 10
    AND NOT overlapping_and_short ),
  --------------------------------------------------------
  -- Likely gear
  -------------------------------------------------------
  likely_gear AS (
  SELECT
    ssvid
  FROM
    `world-fishing-827.gfw_research.vi_ssvid_v20190430`
  WHERE 
   REGEXP_CONTAINS(ais_identity.shipname_mostcommon.value, r"(.*)([\\s]+[0-9]+%)$")
    OR REGEXP_CONTAINS(ais_identity.shipname_mostcommon.value, r"[0-9].[0-9]V")
    OR REGEXP_CONTAINS(ais_identity.shipname_mostcommon.value, r"[0-9]*\\ [0-9]V")
    OR REGEXP_CONTAINS(ais_identity.shipname_mostcommon.value, r"(.*)[@]+([0-9]+V[0-9]?)$")
    OR REGEXP_CONTAINS(ais_identity.shipname_mostcommon.value, r"BOUY")
    OR REGEXP_CONTAINS(ais_identity.shipname_mostcommon.value, r"BUOY")
    OR REGEXP_CONTAINS(ais_identity.shipname_mostcommon.value, r"NET MARK")
    OR REGEXP_CONTAINS(ais_identity.shipname_mostcommon.value, r"NETMARK")
    OR REGEXP_CONTAINS(ais_identity.shipname_mostcommon.value, r"^[0-9]*-[0-9]*$")
    OR REGEXP_CONTAINS(ais_identity.shipname_mostcommon.value, r"NET FISHING")
    OR REGEXP_CONTAINS(ais_identity.shipname_mostcommon.value, r"NETFISHING")),
  ---------------------------------------------------
  -- nast_ssvid are ssvid that are likely offsetting because they have
  -- lots of positions more than 3000 km from the satellite
  ------------------------------------------------------
  nast_ssvid AS (
  SELECT
    ssvid,
    SUM( positions) positions
  FROM
    `world-fishing-827.gfw_research.pipe_production_v20190502_segs`
  WHERE
    ((dist_avg_pos_sat_vessel_km > 3000 AND sat_positions_known > 5) )
    AND first_timestamp < TIMESTAMP("2018-12-31")
    AND last_timestamp > TIMESTAMP("2016-01-01")
  GROUP BY
    ssvid
  HAVING
    positions > 50 ),
  --------------------------------------------------------
  -- good_ssvid here are ones that are definitely not spoofing
  -- and which have a at least 10 days of activity
  -------------------------------------------------------
  good_ssvid AS (
  SELECT
   *
  FROM
    `emlab-gcp.carbon_trawl.ssvid_of_interest` 
  WHERE
    ssvid NOT IN (SELECT ssvid FROM nast_ssvid)
    AND ssvid NOT IN (SELECT ssvid FROM likely_gear)
    AND CAST(ssvid AS int64) not in (SELECT ssvid FROM `world-fishing-827.gfw_research.bad_mmsi` CROSS JOIN UNNEST(ssvid) AS ssvid)),
  -----------------
  -- raw position data
  ----------------
  position_table AS (
  SELECT
    distinct extract(year from timestamp) as year, 
    ssvid,
    lon, 
    lat,
    hours,
    timestamp, 
    nnet_score, 
    speed_knots,
    implied_speed_knots,
    elevation_m
  FROM
    `world-fishing-827.gfw_research.pipe_production_v20190502_fishing`
  WHERE
    timestamp BETWEEN TIMESTAMP("2016-01-01")
    AND TIMESTAMP("2018-12-31")
    AND seg_id IN (SELECT seg_id FROM good_segments)
    and ssvid in (select ssvid from good_ssvid))
  --------------------------------------------------
  -- estimate fishing effort by ssvid, lon, lat
  -------------------------------------------------
SELECT
year, 
ssvid,
FLOOR(lat/0.5)*0.5 + 0.5*0.5 lat_bin,
FLOOR(lon/0.5)*0.5 + 0.5*0.5 lon_bin,
gear_group,
gear_width,
  case 
    when gear_group = "OT" then SUM(IF(speed_knots >= 2 and speed_knots <= 4 and elevation_m >= -2000, hours, 0))
    when gear_group = "BT" then SUM(IF(speed_knots >= 2.5 and speed_knots <= 7 and elevation_m >= -100, hours, 0))  
    when gear_group in ("TD", "HD") then SUM(IF(speed_knots >= 2 and speed_knots <= 2.5, hours, 0))  
    end as fishing_hours,
  case 
    when gear_group = "OT" then SUM(IF(speed_knots >= 2 and speed_knots <= 4 and elevation_m >= -2000, hours*speed_knots, 0))
    when gear_group = "BT" then SUM(IF(speed_knots >= 2.5 and speed_knots <= 7 and elevation_m >= -100, hours*speed_knots, 0))  
    when gear_group in ("TD", "HD") then SUM(IF(speed_knots >= 2 and speed_knots <= 2.5, hours*speed_knots, 0))  
    end as distance_fished_nm,
  sum(if(elevation_m >= -2200, hours*4, 0)) as old_distance_fished_nm
FROM
  position_table
INNER JOIN
  good_ssvid
USING
  (year, ssvid)
WHERE 
  nnet_score > 0.5
GROUP BY
  year, 
  ssvid,
  lat_bin,
  lon_bin,
  gear_group, 
  gear_width
```

```{r, eval = F}
write_rds(binned_effort_by_ssvid, file.path(emLab_project_dir, "data", "02_processed","impacts", "bottom_trawling", "gfw_binned_effort_by_ssvid.rds"))
```


```{r}
binned_effort_by_ssvid <- read_rds(file.path(emLab_project_dir, "data", "02_processed","impacts", "bottom_trawling", "gfw_binned_effort_by_ssvid.rds"))

binned_effort_by_ssvid <- binned_effort_by_ssvid %>% 
  mutate(swept_area_m2 = gear_width*distance_fished_nm*1852,
         old_swept_area_m2 = 200*old_distance_fished_nm*1852) # previous fast approach using 200m widths an 4 knots speed

binned_effort_by_ssvid %>% 
  summarize(sum(swept_area_m2, na.rm = T)/sum(old_swept_area_m2, na.rm = T)) # informed approach minimizes swept area by 85%
```

```{r}
swept_area_raster <- binned_effort_by_ssvid %>%
  filter(fishing_hours != 0) %>% 
  group_by(lon_bin, lat_bin) %>% 
  summarise(avg_swept_area_km2 = 1/3*sum(swept_area_m2, na.rm = T)/10^6) %>% 
  raster::rasterFromXYZ(crs = ocean_low_res_wgs84) %>% 
  projectRaster(ocean_low_res_wgs84) %>%
  mask(ocean_low_res_wgs84)

SAR_raster <- swept_area_raster/raster::area(swept_area_raster)

plot(SAR_raster)
```
 
```{r}
c_raster_05 <- raster(file.path(emLab_project_dir, "data", "02_processed", "carbon","carbon_lr_mean_wgs84.tif")) %>% 
  projectRaster(ocean_low_res_wgs84) %>% 
  mask(ocean_low_res_wgs84)

total_c_05 <- cellStats(c_raster_05, sum)/10^6

SAR_raster_reclass <-  SAR_raster %>% 
  reclassify(c(1, Inf, 1))

c_disturbed_raster <- SAR_raster_reclass*c_raster_05

total_c_disturbed <- c_disturbed_raster %>% 
  raster::cellStats(stat = sum, na.rm = T)/10^6

100*(total_c_disturbed)/total_c_05
```
How does this compare with doing the previous uninformed approach? 

```{r}
effort_raster_05 <- binned_effort_by_ssvid %>%
  group_by(lon_bin, lat_bin) %>%
  summarise(fishing_hours = sum(fishing_hours, na.rm = T)) %>%
  raster::rasterFromXYZ(crs = ocean_low_res_wgs84) %>%
  projectRaster(ocean_low_res_wgs84) %>%
  mask(ocean_low_res_wgs84)

SAR_raster_v2 <- 1/3*(effort_raster_05*4*1.852*0.2)/(raster::area(effort_raster_05))

SAR_raster_v2_reclass <- SAR_raster_v2 %>% 
  reclassify(c(1, Inf, 1)) 

c_disturbed_raster_v2 <- SAR_raster_v2_reclass*c_raster_05

total_c_disturbed_v2 <- c_disturbed_raster_v2 %>% 
  raster::cellStats(stat = sum, na.rm = T)/10^6

100*(total_c_disturbed_v2)/total_c_05  
```


```{r}
#How about against the old analyis? The difference is that this one ignored dredgers

# SAR_raster_old <-  raster("~/gcs/git-annex/global_blue_carbon/files/processed_files/GFW_SAR_05_rez.tif") %>%
#   projectRaster(ocean_low_res_wgs84)
# 
# c_disturbed_raster_old <- SAR_raster_old*c_raster_05
# 
# total_c_disturbed_old<- c_disturbed_raster_old %>% 
#   raster::cellStats(stat = sum, na.rm = T)/10^6
# 
# 100*(total_c_disturbed_old)/total_c_05
```

# Penetration depths

We can use the average penetration depth by gear type from hiddink et al as follows

```{r}
p_depth_by_gear <- tibble(gear_group = c("OT", "BT", "TD", "HD"),
                          p_depth = c(2.44, 2.72, 5.47, 16.11)/100)
```

```{r}
swept_volume_raster <- binned_effort_by_ssvid %>%
  filter(fishing_hours > 0) %>% 
  left_join(p_depth_by_gear) %>% 
  group_by(lon_bin, lat_bin) %>% 
  summarise(avg_vol = 1/3*sum(p_depth*swept_area_m2, na.rm = T)/10^6) %>% 
  raster::rasterFromXYZ(crs = ocean_low_res_wgs84) %>% 
  projectRaster(ocean_low_res_wgs84, over = T) %>%
  mask(ocean_low_res_wgs84)

swept_volume_raster[swept_volume_raster < 0] <- 0

SVR_raster <- swept_volume_raster/raster::area(swept_area_raster)

plot(SVR_raster)

# % carbon disturbed 
100*(raster::cellStats((SVR_raster*c_raster_05), stat = sum, na.rm = T)/10^6)/total_c_05
```

# Sediment types

```{r}
sediments_raster <- raster::raster(file.path(emLab_shrd_data_dir, "seafloor-lithology", "seabed_lithology_v1.nc"))

reclass_df <- tibble(sediment_type = unique(sediments_raster)) %>% 
  mutate(sediment_group = case_when(
    sediment_type %in% c(1, 11) ~ 1, # gravel
    sediment_type %in% c(2, 13) ~ 2, # "sand"
    sediment_type %in% c(3, 4, 12) ~ 3, #"mud"
    sediment_type %in% c(5,6,7,8,9,10) ~ 4, #"biogenic"
    ))

reclassed_sedi_ras <- reclassify(sediments_raster, as.matrix(reclass_df))

reclass_sedi_ras_list <- map(.x = unique(reclassed_sedi_ras), 
    .f =  function(sediment_type) {
             aggregate(reclassed_sedi_ras, fact = c(5,5), fun = function(vals, na.rm = T) {
               sum(vals == sediment_type, na.rm = T)/sum(!is.na(vals))
             }) %>% 
        projectRaster(ocean_low_res_wgs84, method = "ngb")
           })

names(reclass_sedi_ras_list) <- c("gravel", "sand", "mud", "bio")

sedi_stack <- stack(reclass_sedi_ras_list)

sediment_risk_factor <- overlay(sedi_stack, fun = function(g, s, m, b){
  if_else(g > 0.5, 1, 
          if_else(m + b > 0.5, 3,
                  if_else(s > 0.5 | (g + s >= 0.5 & m + b <= 0.5) | (g + m + b >= 0.5 & s <= 0.5) | (m + b + s >= 0.5 & g <= 0.5), 2, 
                          ifelse(is.na(g) & is.na(m) & is.na(s) & is.na(b), 1000, 999))))
  })

sediment_risk_factor_map <- tmap::tm_shape(sediment_risk_factor)+
  tmap::tm_raster(title = "",
                  style = "cat", 
                  palette  = pals::brewer.set2(n = 4),
                  legend.reverse = T)

tmap::tmap_save(sediment_risk_factor_map, 
                filename = file.path(emLab_project_dir, "figures", "exploratory", "sediment_risk_factor.png"),
                dpi = 300,  height = 4, width = 8)
```

Are there data gaps? 

```{r}
overlay(sediment_risk_factor, c_raster_05, fun = function(x, y){is.na(x) & !is.na(y)}) %>% 
  plot()
```

Assign empty sediment type width moderate (2) risk

```{r}
sediment_risk_factor_gf <- sediment_risk_factor

sediment_risk_factor_gf[is.na(sediment_risk_factor_gf)] <- 2

sediment_risk_factor_gf <- sediment_risk_factor_gf %>% 
  mask(ocean_low_res_wgs84)

plot(sediment_risk_factor_gf)
```

```{r}
SVR_raster_gf <- SVR_raster

SVR_raster_gf[is.na(SVR_raster_gf)] <- 0

SVR_raster_gf <- SVR_raster_gf %>%
  mask(ocean_low_res_wgs84) %>% 
  mask(SVR_raster_gf > 0 | is.na(SVR_raster_gf), maskvalue = 0) 

plot(SVR_raster_gf)
```

```{r}
Ia_carbon <- sediment_risk_factor_gf*SVR_raster_gf

Ia_carbon_rescaled <- (Ia_carbon - cellStats(Ia_carbon, 'min')) / (cellStats(Ia_carbon,'max') - cellStats(Ia_carbon, 'min'))

Ia_carbon_rescaled[is.na(Ia_carbon_rescaled[])] <- 0 

Ia_carbon_rescaled <- Ia_carbon_rescaled %>% 
  projectRaster(ocean_low_res_moll, method = "bilinear") %>% 
  mask(ocean_low_res_moll) 

plot(Ia_carbon_rescaled)
hist(Ia_carbon_rescaled)

raster::writeRaster(Ia_carbon_rescaled, 
                    file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling", "abatable_impact_carbon_svr.tif"), overwrite = T)
```

```{r}
Ia_carbon_log <- log(Ia_carbon)

Ia_carbon_log_rescaled <- (Ia_carbon_log - cellStats(Ia_carbon_log, 'min')) / (cellStats(Ia_carbon_log,'max') - cellStats(Ia_carbon_log, 'min'))

Ia_carbon_log_rescaled[is.na(Ia_carbon_log_rescaled[])] <- 0 

Ia_carbon_log_rescaled <- Ia_carbon_log_rescaled %>% 
  projectRaster(ocean_low_res_moll, method = "bilinear") %>% 
  mask(ocean_low_res_moll) 
  
plot(Ia_carbon_log_rescaled)
hist(Ia_carbon_log_rescaled)

raster::writeRaster(Ia_carbon_log_rescaled, 
                    file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling", "abatable_impact_carbon_svr_log.tif"), overwrite = T)
```

# Visualize

```{r}
abatable_impact_carbon_svr <- raster::raster(file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling", "abatable_impact_carbon_svr_log.tif"))

Ia_carbon_map <- abatable_impact_carbon_svr %>% 
  as.data.frame(xy = T) %>% 
  ggplot()+
  geom_raster(aes(x, y, fill = abatable_impact_carbon_svr_log), show.legend = T)+
  geom_sf(data = land_50_moll, fill = 'grey', col = 'grey', lwd = 0.1)+
  my_theme_map()+
  labs(fill = "relative risk") %>% 
  scale_fill_gradientn(colors  = pals::ocean.thermal(n = 30),
                        na.value = "white")
  
Ia_carbon_hist <- abatable_impact_carbon_svr %>% 
  as.data.frame(xy = T) %>% 
  ggplot()+
  geom_histogram(aes(abatable_impact_carbon_svr_log))+ 
  theme_classic()+
  labs(x = "Relative remineralization risk", y = "")

cowplot::plot_grid(Ia_carbon_map, Ia_carbon_hist, labels = NULL, ncol = 1, rel_heights = c(2,1))+
  cowplot::ggsave2(dpi = 300, filename = file.path(emLab_project_dir, "figures", "supplement", "abatable_impacts_carbon.png"), height = 4, width = 6)
```
