---
title: "Estimating bottom trawling footprint"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
editor_options:
  chunk_output_type: inline
---

```{r setup}
library(raster)
library(tidyverse)
library(DBI)
library(bigrquery)
library(sf)
library(connections)

source(here::here("common.R"))
```

```{r}
BQ_connection <-  dbConnect(bigquery(),
                            project = 'emlab-gcp',
                            dataset = "carbon_trawl",
                            billing = "emlab-gcp", # your billing account name
                            use_legacy_sql = FALSE) # specify we are using Standard SQL

gfw_research <- connection_open(bigquery(),
                                project = "world-fishing-827",  # Note how I connect to the GFW project
                                dataset = "gfw_research",
                                billing = "world-fishing-827",          # But the billing stays the same
                                use_legacy_sql = FALSE,
                                allowLargeResults = TRUE)

#knitr::opts_chunk$set(connection = "BQ_connection")

options(scipen = 20)
```

# Vessels of interest

```{r, eval = F}
gfw_fishing_vessels <- tbl(gfw_research, "fishing_vessels_ssvid_v20200410")

bad_mmsi <- tbl(gfw_research, "bad_mmsi") %>% 
  collect() %>% 
  unnest() %>% 
  transmute(ssvid = as.character(ssvid)) 

likely_gear <- tbl(gfw_research, "vi_ssvid_v20200410") %>% 
  filter(on_fishing_list_best) %>% 
  mutate(shipname_mostcommon = ais_identity$shipname_mostcommon$value) %>%
  filter(REGEXP_CONTAINS(shipname_mostcommon, "(.*)([\\s]+[0-9]+%)$")|
           REGEXP_CONTAINS(shipname_mostcommon, "[0-9].[0-9]V")|
           REGEXP_CONTAINS(shipname_mostcommon, "[0-9]*\\ [0-9]V")|
           REGEXP_CONTAINS(shipname_mostcommon, "(.*)[@]+([0-9]+V[0-9]?)$")|
           REGEXP_CONTAINS(shipname_mostcommon, "BOUY")|
           REGEXP_CONTAINS(shipname_mostcommon, "BUOY")|
           REGEXP_CONTAINS(shipname_mostcommon, "NET MARK")|
           REGEXP_CONTAINS(shipname_mostcommon, "NETMARK")|
           REGEXP_CONTAINS(shipname_mostcommon, "^[0-9]*-[0-9]*$")|
           REGEXP_CONTAINS(shipname_mostcommon, "NET FISHING")|
           REGEXP_CONTAINS(shipname_mostcommon, "NETFISHING")) %>% 
  select(ssvid) 

nasty_ssvid <- tbl(gfw_research, "pipe_v20190502_segs") %>% 
  filter(dist_avg_pos_sat_vessel_km > 3000,
         sat_positions_known > 5,
         first_timestamp < TIMESTAMP("2018-12-31"),
         last_timestamp > TIMESTAMP("2016-01-01")) %>% 
  group_by(ssvid) %>% 
  summarize(positions = sum(positions, na.rm = T)) %>% 
  ungroup() %>% 
  filter(positions > 50) %>% 
  select(ssvid)

gfw_trawlers <- tbl(gfw_research, "vi_ssvid_byyear_v20200410") %>% 
  filter(on_fishing_list_best &
           best$best_vessel_class %in% c("trawlers", "dredge_fishing") & 
           year > 2015 &
           year < 2020) %>% 
  filter(!ssvid %in% !!bad_mmsi$ssvid) %>% 
  mutate(ship_name = ais_identity$n_shipname_mostcommon$value,
         imo = registry_info$best_known_imo,
         callsign = registry_info$best_known_callsign,
         flag = best$best_flag,
         length = best$best_length_m,
         engine_power = best$best_engine_power_kw,
         vessel_class = best$best_vessel_class,
         active_hours = activity$active_hours,
         fishing_hours = activity$fishing_hours,
         avg_fishing_depth = activity$avg_depth_fishing_m) %>% 
  select(year, ssvid, ship_name, imo, callsign, flag, length, engine_power, vessel_class, active_hours, fishing_hours, avg_fishing_depth) %>% 
  inner_join(gfw_fishing_vessels) %>% 
  anti_join(full_join(nasty_ssvid, 
                      likely_gear)) %>% 
  collect()
```

# Refine gear type

## EU list

```{sql, connection = "BQ_connection", output.var = "EU_gear_types", eval = F}
SELECT
  DISTINCT CAST(mmsi AS string)ssvid,
  shipname_matched,
  callsign_matched,
  imo_matched,
  registry.geartype_original,
  regexp_extract(registry.geartype_original,"^[^-]*[^ -]") as main_gear,  
  regexp_extract(registry.geartype_original,"[^ -][^-]*$") as sec_gear
FROM
  `world-fishing-827.vessel_database_staging.matched_eu_v20180613`,
  UNNEST(registry) AS registry
WHERE
  geartype_original IS NOT NULL
  AND mmsi IS NOT NULL
  and matched
  AND REGEXP_CONTAINS(geartype_original, "DRB|DTS|HMD|OTB|OTT|PTB|TBB|OTM|PTM|TM")
ORDER BY
  ssvid
```

## CCAMLR

```{sql, connection = "BQ_connection", output.var = "CCAMLR_OTM", eval = F}
SELECT
  DISTINCT CAST(mmsi AS string)ssvid,
  shipname_matched,
  callsign_matched,
  imo_matched,
  registry.geartype_original
FROM
  `world-fishing-827.vessel_database_staging.matched_ccamlr_v20200301` ,
  UNNEST(registry) AS registry
WHERE
 REGEXP_CONTAINS(geartype_original, "MIDWATER")
```

```{r, eval = F}
write_rds(gfw_trawlers, file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling","gfw_trawlers.rds"))
write_rds(CCAMLR_OTM, file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling", "gfw_CCAMLR_vessels.rds"))
write_rds(EU_gear_types, file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling", "gfw_EU_vessels.rds"))
```

## join lists

```{r}
gfw_trawlers <- read_rds(file.path(emLab_project_dir, "data", "02_processed","impacts", "bottom_trawling","gfw_trawlers.rds"))
CCAMLR_OTM <- read_rds(file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling","gfw_CCAMLR_vessels.rds"))
EU_gear_types <- read_rds(file.path(emLab_project_dir, "data", "02_processed","impacts", "bottom_trawling", "gfw_EU_vessels.rds"))
```

```{r}
EU_gear_group <- EU_gear_types %>% 
  mutate(gear_group = case_when(
    main_gear %in% c("TM", "OTM", "PTM") ~ "OTM",
    main_gear %in% c("HMD")  ~ "HD",
    main_gear %in% c("DRB") ~ "TD",
    main_gear %in% c("TBB")  ~ "BT", 
    main_gear %in% c("OTB", "OTT", "DTS", "PTB") ~ "OT",
    sec_gear %in% c("TM", "OTM", "PTM") ~ "OTM",
    sec_gear %in% c("HMD")  ~ "HD",
    sec_gear %in% c("DRB") ~ "TD",
    sec_gear %in% c("TBB")  ~ "BT", 
    sec_gear %in% c("OTB", "OTT", "DTS", "PTB") ~ "OT"
  )) %>%
  group_by(ssvid) %>% 
  filter(n_distinct(gear_group) == 1) %>%
  distinct(ssvid, gear_group)
```

If there is no gear group we assume they are otter trawls as they are the most common. We also remove all midwater trawlers. 

```{r}
vessels_of_interest <- gfw_trawlers %>% 
  left_join(EU_gear_group) %>% 
  mutate(gear_group = case_when(
    !is.na(gear_group) ~ gear_group,
    vessel_class == "trawlers" ~ "OT",
    vessel_class == "dredge_fishing" ~ "TD"
  )) %>% 
  filter(!ssvid %in% CCAMLR_OTM$ssvid) %>%
  filter(gear_group != "OTM") 

vessels_of_interest %>% 
  group_by(gear_group) %>% 
  summarize(n_ssvid = n_distinct(ssvid))
```

# Gear width

Using the estimates from [Eigaard et al](https://academic.oup.com/icesjms/article/73/suppl_1/i27/2573989#supplementary-data) we can estimate gear width as:

  - Dredgers: operate at 2-2.5 kts, very shallow, and use 0.72-3m width dredges. 

$$W = 0.3142*LOA^{1.2454}$$
  - Otter trawl (OT): operate at 2-4 kts, 25-250 m between doors width, between 10-2500 m deep. Equation from OT_MIX group,  representative of most number of species 

$$W = 10.6608*KW^{0.2921}$$
  - Beam trawl (BT): operate usually two beam with a total widht between 4-12m, speed between 2.5-7 knts,  and shallower than 100m. Equation from TBB_DMF group, representative of most number of species 

$$W = 0.6601*KW^{0.5078}$$

```{r}
vessels_of_interest <- vessels_of_interest %>% 
  mutate(gear_width = case_when(
    gear_group == "BT" ~ 0.6601*engine_power^(0.5078),
    gear_group == "OT" ~ 10.6608*engine_power^(0.2921),
    gear_group %in% c("TD", "HD") ~ 0.3142*length^(1.2454)
  )) 

vessels_of_interest %>% 
  ggplot()+
  geom_density(aes(gear_width,  fill = gear_group))+
  ggsave(file.path(emLab_project_dir, "figures", "exploratory", "bottom_trawling_gear_width_distributions.png"), 
         dpi = 300, height = 4, width = 8)
```

```{r, eval = F}
DBI::dbWriteTable(conn = BQ_connection,
                  name = "ssvid_of_interest",
                  value = vessels_of_interest[c("year", "ssvid", "gear_group", "gear_width")],
                  fields = as_bq_fields(vessels_of_interest[c("year", "ssvid", "gear_group", "gear_width")]),
                  overwrite = T)
```

# Query Effort

```{r, eval = F}
vessels_of_interest <- gfw_research %>% 
    tbl("emlab-gcp.carbon_trawl.ssvid_of_interest")

good_segs <- tbl(gfw_research, "pipe_v20190502_segs")%>% 
  filter(positions > 10, 
         overlapping_and_short == FALSE, 
         good_seg == TRUE) %>% 
  select(seg_id)

res <- 0.5

ais_positions <- gfw_research %>% 
    tbl("pipe_v20190502_fishing") %>% 
    mutate(year = year(date)) %>% 
    select(year, date, timestamp, ssvid, seg_id, lat, lon, speed_knots, hours, nnet_score, regions, elevation_m) %>% 
    filter(timestamp < TIMESTAMP("2020-01-01"),
           timestamp > TIMESTAMP("2016-01-01"),
           nnet_score == 1,
           speed_knots > 0) %>% 
  inner_join(vessels_of_interest, by = c("ssvid", "year")) %>% 
  inner_join(good_segs) 

binned_effort_by_ssvid <-  ais_positions %>% 
  mutate(lat_bin = floor(lat/res)*res + 0.5*res,
         lon_bin = floor(lon/res)*res + 0.5*res) %>% 
  mutate(hours = case_when(gear_group == "OT" & speed_knots >= 2 & speed_knots <= 4 & elevation_m >= -2000 ~ hours,
                           gear_group == "BT" & speed_knots >= 2.5 & speed_knots <= 7 & elevation_m >= -100 ~ hours,
                           gear_group %in% c("TD", "HD")  & speed_knots >= 2 & speed_knots <= 2.5 ~ hours,
                           TRUE ~ 0),
         gear_width = round(gear_width, 2)) %>% 
  group_by(year, ssvid, gear_group, gear_width, lat_bin, lon_bin) %>% 
  summarize(fishing_hours = round(sum(hours, na.rm = T), 2),
            distance_fished_nm = round(sum(hours*speed_knots, na.rm = T), 2)) %>%
  ungroup() %>% 
  mutate(swept_area_m2 = gear_width*distance_fished_nm*1852) %>% 
  filter(fishing_hours > 0) %>% 
  collect()

write_rds(binned_effort_by_ssvid, 
          file.path(emLab_project_dir, "data", "02_processed","impacts", "bottom_trawling", "gfw_binned_effort_by_ssvid_updated.rds"))
```

```{r}
binned_effort_by_ssvid <- read_rds(file.path(emLab_project_dir, 
                                             "data", "02_processed","impacts", "bottom_trawling", "gfw_binned_effort_by_ssvid_updated.rds"))

binned_effort_by_ssvid %>%
  filter(fishing_hours != 0) %>% 
  group_by(year) %>% 
  summarize(n_distinct(ssvid))
```

# Four Metrics:

## Swept Area

```{r}
swept_area_raster <- binned_effort_by_ssvid %>%
  filter(fishing_hours != 0) %>% 
  group_by(lon_bin, lat_bin) %>% 
  summarise(avg_swept_area_km2 = 1/4*sum(swept_area_m2, na.rm = T)/10^6) %>% 
  raster::rasterFromXYZ(crs = ocean_low_res_wgs84) %>% 
  projectRaster(to = ocean_low_res_wgs84) %>% 
  mask(ocean_low_res_wgs84)

swept_area_raster[swept_area_raster < 0] <- 0

SAR_raster <- swept_area_raster/raster::area(swept_area_raster)

SAR_raster <- SAR_raster %>% 
  projectRaster(to = ocean_low_res_moll, method = "bilinear") %>% 
  mask(ocean_low_res_moll)

SAR_raster_reclass <-  SAR_raster %>% 
  reclassify(c(1, Inf, 1))
```
 
```{r}
carbon_lr <- raster(file.path(emLab_project_dir, "data", "02_processed", "carbon", "carbon_lr_mean_moll.tif")) 

total_carbon_lr <- cellStats(carbon_lr, sum)

disturbed_carbon <- SAR_raster_reclass*carbon_lr

total_disturbed_carbon <- raster::cellStats(disturbed_carbon, stat = sum, na.rm = T)

# fraction carbon disturbed

total_disturbed_carbon/total_carbon_lr # aprox 1.5%

# total C disturbed
total_disturbed_carbon/10^6 # 14.14 

# max remineralized CO2

3.67*total_disturbed_carbon/10^6 # 51.89
```

## Swept Volume 

```{r}
p_depth_by_gear <- tibble(gear_group = c("OT", "BT", "TD", "HD"),
                          p_depth = c(2.44, 2.72, 5.47, 16.11)/100)

swept_volume_raster <- binned_effort_by_ssvid %>%
  filter(fishing_hours > 0) %>% 
  left_join(p_depth_by_gear) %>% 
  group_by(lon_bin, lat_bin) %>% 
  summarise(avg_vol_km3 = 1/4*sum(p_depth*swept_area_m2, na.rm = T)/10^9) %>% 
  raster::rasterFromXYZ(crs = ocean_low_res_wgs84) %>% 
  projectRaster(ocean_low_res_wgs84, over = T) %>%
  mask(ocean_low_res_wgs84)

swept_volume_raster[swept_volume_raster < 0] <- 0

svr_raster <- swept_volume_raster/(raster::area(swept_area_raster)*0.001)

svr_raster[is.na(svr_raster)] <- 0

svr_raster <- svr_raster %>%
  projectRaster(ocean_low_res_moll) %>% 
  mask(ocean_low_res_moll)

svr_raster <- svr_raster %>% 
  mask(svr_raster > 0 | is.na(svr_raster), maskvalue = 0) 

plot(svr_raster)

raster::writeRaster(svr_raster, 
                    file.path(emLab_project_dir, 
                              "data", "02_processed", "impacts", "bottom_trawling", "svr.tif"), overwrite = T)
```

```{r}
svr_disturbed_carbon <- svr_raster*carbon_lr

total_svr_disturbed_carbon <- raster::cellStats(svr_disturbed_carbon, stat = sum, na.rm = T)

# fraction carbon disturbed

total_svr_disturbed_carbon/total_carbon_lr # 0.00048

# total C disturbed
total_svr_disturbed_carbon/10^6 # 0.48

# max remineralized CO2

3.67*total_svr_disturbed_carbon/10^6 # 1.76
```

## Remineralization risk 

```{r, eval = F}
sediments_raster <- raster::raster(file.path(emLab_shrd_data_dir, "seafloor-lithology", "seabed_lithology_v1.nc"))

reclass_df <- tibble(sediment_type = unique(sediments_raster)) %>% 
  mutate(sediment_group = case_when(
    sediment_type %in% c(1, 11) ~ 1, # gravel
    sediment_type %in% c(2, 13) ~ 2, # "sand"
    sediment_type %in% c(3, 4, 12) ~ 3, #"mud"
    sediment_type %in% c(5,6,7,8,9,10) ~ 4, #"biogenic"
    ))

reclassed_sedi_ras <- reclassify(sediments_raster, as.matrix(reclass_df))

reclass_sedi_ras_list <- map(.x = unique(reclassed_sedi_ras), 
    .f =  function(sediment_type) {
             aggregate(reclassed_sedi_ras, fact = c(5,5), fun = function(vals, na.rm = T) {
               sum(vals == sediment_type, na.rm = T)/sum(!is.na(vals))
             }) %>% 
        projectRaster(ocean_low_res_wgs84, method = "ngb")
           })

names(reclass_sedi_ras_list) <- c("gravel", "sand", "mud", "bio")

sedi_stack <- stack(reclass_sedi_ras_list)

sediment_risk_factor <- overlay(sedi_stack, fun = function(g, s, m, b){
  if_else(g > 0.5, 1, 
          if_else(m + b > 0.5, 3,
                  if_else(s > 0.5 | (g + s >= 0.5 & m + b <= 0.5) | (g + m + b >= 0.5 & s <= 0.5) | (m + b + s >= 0.5 & g <= 0.5), 2, 
                          ifelse(is.na(g) & is.na(m) & is.na(s) & is.na(b), 1000, 999))))
  })

sediment_risk_factor_map <- tmap::tm_shape(sediment_risk_factor)+
  tmap::tm_raster(title = "",
                  style = "cat", 
                  palette  = pals::brewer.set2(n = 4),
                  legend.reverse = T)

tmap::tmap_save(sediment_risk_factor_map, 
                filename = file.path(emLab_project_dir, "figures", "exploratory", "sediment_risk_factor.png"),
                dpi = 300,  height = 4, width = 8)
```

Assign empty sediment type width moderate (2) risk

```{r, eval = F}
sediment_risk_factor_gf <- sediment_risk_factor

sediment_risk_factor_gf[is.na(sediment_risk_factor_gf)] <- 2

sediment_risk_factor_gf %>% 
  projectRaster(ocean_low_res_moll, method = "ngb") %>% 
  mask(ocean_low_res_moll) %>% 
  raster::writeRaster(file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling", "sediment_risk_factor_gf.tif"), overwrite=TRUE)
```

```{r, eval = F}
# sediment_risk_factor_gf <- raster(file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling", "sediment_risk_factor_gf.tif"))
# 
# svr_disturbed_carbon %>% 
#   as.data.frame(xy = T) %>% 
#   drop_na() %>% 
#   left_join(sediment_risk_factor_gf %>% 
#               as.data.frame(xy = T)) %>% 
#   group_by(sediment_risk_factor_gf) %>% 
#   summarize(total_c = sum(layer, na.rm = T)) %>% 
#   mutate(f = total_c/sum(total_c))
```

```{r, eval = F}
Ia_carbon <- sediment_risk_factor_gf*SVR_raster_gf

Ia_carbon_rescaled <- (Ia_carbon - cellStats(Ia_carbon, 'min')) / (cellStats(Ia_carbon,'max') - cellStats(Ia_carbon, 'min'))

Ia_carbon_rescaled[is.na(Ia_carbon_rescaled[])] <- 0 

Ia_carbon_rescaled <- Ia_carbon_rescaled %>% 
  projectRaster(ocean_low_res_moll, method = "bilinear") %>% 
  mask(ocean_low_res_moll) 

plot(Ia_carbon_rescaled)
hist(Ia_carbon_rescaled)

raster::writeRaster(Ia_carbon_rescaled, 
                    file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling", "abatable_impact_carbon_svr.tif"), 
                    overwrite = F)
```

```{r,  eval = F}
Ia_carbon_log <- log(Ia_carbon)

Ia_carbon_log_rescaled <- (Ia_carbon_log - cellStats(Ia_carbon_log, 'min')) / (cellStats(Ia_carbon_log,'max') - cellStats(Ia_carbon_log, 'min'))

Ia_carbon_log_rescaled[is.na(Ia_carbon_log_rescaled[])] <- 0 

Ia_carbon_log_rescaled <- Ia_carbon_log_rescaled %>% 
  projectRaster(ocean_low_res_moll, method = "bilinear") %>% 
  mask(ocean_low_res_moll) 
  
plot(Ia_carbon_log_rescaled)
hist(Ia_carbon_log_rescaled)

raster::writeRaster(Ia_carbon_log_rescaled, 
                    file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling", "abatable_impact_carbon_svr_log.tif"),
                    overwrite = F)
```

```{r, eval = F}
abatable_impact_carbon_svr <- raster::raster(file.path(emLab_project_dir, 
                                                       "data", "02_processed", "impacts", "bottom_trawling", "abatable_impact_carbon_svr_log.tif"))

Ia_carbon_map <- abatable_impact_carbon_svr %>% 
  as.data.frame(xy = T) %>% 
  ggplot()+
  geom_raster(aes(x, y, fill = abatable_impact_carbon_svr_log), show.legend = T)+
  geom_sf(data = land_50_moll, fill = 'grey', col = 'grey', lwd = 0.1)+
  my_theme_map()+
  labs(fill = "relative risk") %>% 
  scale_fill_gradientn(colors  = pals::ocean.thermal(n = 30),
                        na.value = "white")
  
Ia_carbon_hist <- abatable_impact_carbon_svr %>% 
  as.data.frame(xy = T) %>% 
  ggplot()+
  geom_histogram(aes(abatable_impact_carbon_svr_log))+ 
  theme_classic()+
  labs(x = "Relative remineralization risk", y = "")

cowplot::plot_grid(Ia_carbon_map, Ia_carbon_hist, labels = NULL, ncol = 1, rel_heights = c(2,1))+
  cowplot::ggsave2(dpi = 300, 
                   filename = file.path(emLab_project_dir, "figures", "supplement", "abatable_impacts_carbon.png"), 
                   height = 4, width = 6)
```

## Remineralized CO2

```{r, eval = F}
p_labile <- sediment_risk_factor_gf

p_labile[p_labile == 1] <- 0.04
p_labile[p_labile == 2] <- 0.286
p_labile[p_labile == 3] <- 0.7

plot(p_labile)

p_labile %>%
  projectRaster(ocean_low_res_moll, method = "ngb") %>% 
  raster::writeRaster(file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling", "p_labile.tif"), overwrite=TRUE)
```

```{r, eval = F}
seas_shp <- sf::st_read(file.path(emLab_shrd_data_dir, "World_Seas_IHO_v3", "World_Seas_IHO_v3.shp")) 

seas_lut <- read_csv(file.path(emLab_project_dir, "data", "01_raw", "sea_regions.csv")) %>% 
  janitor::clean_names()

seas_shp <- seas_shp %>% 
  left_join(seas_lut, by = c("ID" = "iso_id")) %>% 
  mutate(project_region = trim(project_region),
         k = case_when(project_region == c("Atlantic") | iso_name == "Baltic Sea" ~ 3.62*10^(-1),
                       project_region == "Arctic" ~ 2.75*10^(-1),
                       project_region == "Gulf of Mexico" ~ 1.68*10^(1),
                       project_region == "Indian" ~ 4.76*10^(0),
                       project_region == "Mediterranean" ~ 1.23*10^(1),
                       project_region == "North Pacific" ~ 1.67*10^(0),
                       project_region == "South Pacific" ~ 3.84*10^(0))) 

k_decay <- rasterize(seas_shp, ocean_low_res_wgs84, method = "ngb", field = "k") 

k_decay %>% 
  projectRaster(ocean_low_res_moll, method = "ngb") %>% 
  raster::writeRaster(file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling", "k_decay.tif"), 
                      overwrite=TRUE)

plot(k_decay)
```

```{r}
p_crd <- 0.87

svr_raster_gf <- raster(file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling", "svr_gf.tif"))

k_decay <- raster(file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling", "k_decay.tif")) 

p_labile <- raster(file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling", "p_labile.tif")) 

remineralized_carbon <- carbon_lr*(svr_raster_gf*p_crd*p_labile)*(1 - exp(-k_decay))

total_remineralized_carbon <- raster::cellStats(remineralized_carbon, stat = sum, na.rm = T)

# fraction carbon remineralized

total_remineralized_carbon/total_carbon_lr # 0.0001374601

# total C remineralized
total_remineralized_carbon/10^6 # 0.1363386

# remineralized to CO2

3.67*total_remineralized_carbon/10^6 # 0.5003627

# total C in ocean

total_1km_carbon <-  3117263*10^6

# total remineralized C

3.67*total_1km_carbon*(total_remineralized_carbon/total_carbon_lr)/10^9 # 1.57x10^9 Mg CO2

# Money?

50*3.67*total_1km_carbon*(total_remineralized_carbon/total_carbon_lr)/10^9 # 78.62961 billion usd at 50$ per ton of CO2

plot(round(remineralized_carbon*total_1km_carbon/total_carbon_lr))
```


```{r}
carbon_lr_sum <- raster(file.path(emLab_project_dir, "data", "02_processed", "carbon", "carbon_lr_sum_wgs84.tif")) %>% 
   projectRaster(ocean_low_res_moll, 'bilinear') 

carbon_lr_sum[carbon_lr_sum<0] <- NA

sum_total_carbon <- raster::cellStats(carbon_lr_sum, stat = sum, na.rm = T)

rem_carbon_v2 <- (carbon_lr_sum*SVR_raster_gf*p_crd*p_labile) - (carbon_lr_sum*SVR_raster_gf*p_crd*p_labile)*exp(-k_decay)

# fraction oftotal C remineralized

raster::cellStats(rem_carbon_v2, stat = sum, na.rm = T)/sum_total_carbon # 0.0001270932

# total remineralized to CO2

3.67*raster::cellStats(rem_carbon_v2, stat = sum, na.rm = T)/10^9 # 1.35

# Money?

50*raster::cellStats(rem_carbon_v2, stat = sum, na.rm = T)*3.67/10^9  # 68 billion usd at 50$ per ton of CO2
```

### validation

```{r, eval = F}
validation_df <- tibble(name = c("Westerchlede", "White Oak",  "Thermaic gulf","Sweden", "Aarhus Bay"),
                        obs_mol_co2_km2 = c(3.55E+07, 3.99E+06, 8.74E+06, 4.39E+06,  1.26E+07))

validation_df$poly[validation_df$name == "Westerchlede"] <- "POLYGON((3.348461 51.549847, 4.311777 51.549847, 4.311777 51.310398, 3.348461 51.310398, 3.348461 51.549847))"
validation_df$poly[validation_df$name == "White Oak"] <- "POLYGON((-77.150693 34.770435, -77.067296 34.770435, -77.067296 34.691695, -77.150693 34.691695, -77.150693 34.770435))"
validation_df$poly[validation_df$name == "Thermaic gulf"] <- "POLYGON((22.5323 40.6552, 23.3282 40.6552, 23.3282 39.831, 22.5323 39.831, 22.5323 40.6552))"
validation_df$poly[validation_df$name == "Sweden"] <- "POLYGON((11.666691 57.702016, 11.758501 57.702016, 11.758501 57.660455, 11.666691 57.660455, 11.666691 57.702016))"
validation_df$poly[validation_df$name == "Aarhus Bay"] <- "POLYGON((10.1941 56.3014, 10.6975 56.3014, 10.6975 55.7348, 10.1941 55.7348, 10.1941 56.3014))"

validation_sf <- validation_df %>% 
  sf::st_as_sf(wkt = "poly", crs = 4326) %>%
  st_transform(crs = crs(ocean_low_res_moll)) %>% 
  rownames_to_column("region_id")
```

```{r}
validation_df_lr <- validation_sf %>% 
  left_join(rem_carbon_v2 %>% 
              raster::extract(validation_sf, df = T) %>% 
              set_names(c("region_id", "rem_carbon")) %>% 
              mutate(region_id = as.character(region_id))) %>% 
  filter(!is.na(rem_carbon)) %>% 
  group_by(name) %>% 
  summarize(n_pixels = n(),
            obs_mol_co2_km2 = mean(obs_mol_co2_km2, na.rm = T),
            rem_co2 = 3.67*sum(rem_carbon, na.rm = T),
            rem_mol_co2 = rem_co2*22722, #convert from Mg CO2 to mol CO2
            rem_mol_co2_km2 = rem_mol_co2/(n_pixels*25000))    # divide by the area of each cell. We can do this because we are in mollweide here

validation_df_lr %>% 
  ggplot()+
  geom_point(aes(x = log(obs_mol_co2_km2), y = log(rem_mol_co2_km2)))
```

### High Res

```{r, eval = F}
carbon_hr <- raster(file.path(emLab_shrd_data_dir, "global-marine-sedimentary-carbon", "Mean carbon_stock.tif"))

res <- 0.01

binned_effort_by_ssvid_hr <-  ais_positions %>% 
  mutate(lat_bin = floor(lat/res)*res + 0.5*res,
         lon_bin = floor(lon/res)*res + 0.5*res) %>% 
  mutate(hours = case_when(gear_group == "OT" & speed_knots >= 2 & speed_knots <= 4 & elevation_m >= -2000 ~ hours,
                           gear_group == "BT" & speed_knots >= 2.5 & speed_knots <= 7 & elevation_m >= -100 ~ hours,
                           gear_group %in% c("TD", "HD")  & speed_knots >= 2 & speed_knots <= 2.5 ~ hours,
                           TRUE ~ 0),
         gear_width = round(gear_width, 2)) %>% 
  group_by(year, ssvid, gear_group, gear_width, lat_bin, lon_bin) %>% 
  summarize(fishing_hours = round(sum(hours, na.rm = T), 2),
            distance_fished_nm = round(sum(hours*speed_knots, na.rm = T), 2)) %>%
  ungroup() %>% 
  mutate(swept_area_m2 = gear_width*distance_fished_nm*1852) %>% 
  filter(fishing_hours > 0) %>% 
  collect()

write_rds(binned_effort_by_ssvid_hr, 
          file.path(emLab_project_dir, "data", "02_processed","impacts", "bottom_trawling", "gfw_binned_effort_by_ssvid_updated_hr.rds"))

swept_volume_raster_hr <- binned_effort_by_ssvid_hr %>%
  filter(fishing_hours > 0) %>% 
  left_join(p_depth_by_gear) %>% 
  group_by(lon_bin, lat_bin) %>% 
  summarise(avg_vol = 1/4*sum(p_depth*swept_area_m2, na.rm = T)/10^6) %>% 
  raster::rasterFromXYZ(crs = carbon_hr) %>% 
  projectRaster(carbon_hr, over = T) %>%
  mask(carbon_hr)

swept_volume_raster_hr[swept_volume_raster_hr < 0] <- 0

SVR_raster_hr <- swept_volume_raster_hr/raster::area(swept_volume_raster_hr)

#SVR_raster_hr[is.na(SVR_raster_hr)] <- 0

# SVR_raster_gf <- SVR_raster_gf %>%
#   projectRaster(ocean_low_res_moll) %>% 
#   mask(ocean_low_res_moll)
# 
# SVR_raster_gf <- SVR_raster_gf%>% 
#   mask(SVR_raster_gf > 0 | is.na(SVR_raster_gf), maskvalue = 0) 

plot(SVR_raster_hr)

SVR_raster_hr %>% 
  raster::writeRaster(file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling", "svr_hr.tif"), 
                      overwrite=TRUE)
```


```{r, eval = F}
p_labile_hr <- p_labile %>% 
  projectRaster(carbon_hr, "ngb")

p_labile_hr %>% 
  raster::writeRaster(file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling", "p_labile_hr.tif"), overwrite=TRUE)
  
k_decay_hr <- k_decay %>% 
  projectRaster(carbon_hr, "ngb")

k_decay_hr %>% 
  raster::writeRaster(file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling", "k_decay_hr.tif"), overwrite=TRUE)
```

```{r}
carbon_hr <- raster(file.path(emLab_shrd_data_dir, "global-marine-sedimentary-carbon", "Mean carbon_stock.tif"))

k_decay_hr <- raster(file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling", "k_decay_hr.tif"))

p_labile_hr <- raster(file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling", "p_labile_hr.tif"))

SVR_raster_hr <- raster(file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling", "svr_hr.tif"))

SVR_raster_hr <- raster("../../../Google Drive File Stream/Shared drives/emlab/projects/current-projects/carbon-trawl/data/03_output/carbon_disturbed_svr_per_km2.tif")
```

```{r, eval = F}
rem_carbon_hr <- (carbon_hr*SVR_raster_hr*p_crd*p_labile_hr)*(1 - exp(-k_decay_hr))

rem_carbon_hr %>%
  raster::writeRaster(file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling", "rem_carbon_hr.tif"), overwrite=TRUE)
```

```{r}
rem_carbon_hr <- raster(file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling", "rem_carbon_hr.tif"))

validations_hr <- validation_sf %>% 
  st_transform(crs = crs(rem_carbon_hr))

tmp <- validations_hr %>% 
  left_join(rem_carbon_hr %>% 
              raster::extract(validations_hr, df = T) %>% 
              set_names(c("region_id", "rem_carbon")) %>% 
              mutate(region_id = as.character(region_id))) %>% 
  filter(rem_carbon > 0) %>% 
  group_by(name) %>% 
  summarize(obs_mol_co2_km2 = mean(obs_mol_co2_km2, na.rm = T),
            model_Mg_co2 = sum(rem_carbon*3.67, na.rm = T),
            model_mol_co2 = sum(model_Mg_co2*22722, na.rm = T))
  
areas_hr <- raster::area(rem_carbon_hr) %>% 
  mask(rem_carbon_hr)

t <- rem_carbon_hr*areas_hr

total_audra <- cellStats(t, sum)

total_audra/10^9

plot(areas_hr)

fishing_area <- validations_hr %>% 
  left_join(areas_hr %>% 
              raster::extract(validations_hr, df = T) %>% 
              set_names(c("region_id", "fishing_area")) %>% 
              mutate(region_id = as.character(region_id))) %>% 
  filter(fishing_area > 0) %>% 
  group_by(name) %>% 
  summarize(fishing_area_km2 = sum(fishing_area, na.rm = T)) %>% 
  st_drop_geometry()

validation_df_hr <- tmp %>% 
  left_join(fishing_area) %>% 
  mutate(model_mol_co2_km2 = model_mol_co2/fishing_area_km2) 


validation_df_hr %>% 
  mutate(model_mol_co2_km2/obs_mol_co2_km2)



  ggplot()+
  geom_point(aes(x = log(model_mol_co2_km2), y = log(obs_mol_co2_km2)))+
  geom_abline(slope = 1, intercept = 0)
```
