---
title: "Carbon in Juanation"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
---

```{r}
library(raster)
library(sf)
library(tidyverse)
library(patchwork)

source(here::here("common.R"))

variant_name <- "06_carbon"

variant_dir <- file.path(emLab_project_dir,"data", "03_output", variant_name)

dir.create(variant_dir)
```

# Load features data 

```{r}
carbon_features_info <- tibble(filepath = file.path(emLab_project_dir, "data", "02_processed", "carbon", "carbon_lr_mean_moll.tif"),
                               w = 1)
```

```{r, eval = F}
carbon_features_df <- stack(carbon_features_info$filepath) %>%
  raster::as.data.frame(xy = T) %>%
  inner_join(ocean_matrix) %>%
  select(x, y, cell_id, everything()) %>%
  rename(carbon = layer) %>% 
  as_tibble()

data.table::fwrite(carbon_features_df, 
                   file.path(emLab_project_dir, "data", "02_processed", "features_df", "carbon_df.csv"))
```

```{r}
features_df <- data.table::fread(file.path(emLab_project_dir, "data", "02_processed", "features_df", "carbon_df.csv")) %>% 
  as_tibble()
```

## Normalize and trim

```{r}
features_matrix <- features_df %>% 
  select(-x,-y,-cell_id) %>% 
  as.matrix()

stopifnot(
  sum(map_lgl(features_df, is.numeric)) == ncol(features_df) 
  ) # ALL numeric?

Norm_features_matrix <- sweep(features_matrix, 2, colSums(features_matrix, na.rm = T), FUN = "/")

stopifnot(
  sum(colSums(Norm_features_matrix, na.rm = T)) == ncol(features_matrix)
  )  # Normalized OK?

rownames(Norm_features_matrix) <- features_df$cell_id

Norm_features_matrix <- Norm_features_matrix[rowSums(is.na(Norm_features_matrix)) != ncol(Norm_features_matrix), ] 

Norm_features_matrix <- as.matrix(Norm_features_matrix)

colnames(Norm_features_matrix) <- "carbon"

stopifnot(
  identical(colnames(Norm_features_matrix), 
          features_df %>% 
            select(-x,-y,-cell_id) %>% 
            colnames())
  )  # Is the order of the features mantained?
```

# Current MPAs

```{r}
MPAs <- raster(file.path(emLab_project_dir, "data", "02_processed", "masks", "fully_highly_protected_reviewed_MPAs.tif"))

MPAs_df <- MPAs %>% 
  raster::as.data.frame(xy = T) %>%
  inner_join(ocean_matrix) %>% 
  select(x, y, cell_id, mpa = fully_highly_protected_reviewed_MPAs) %>%
  as_tibble() 

fraction_in_MPA <- sum(!is.na(MPAs_df$mpa))/nrow(MPAs_df)
  
MPA_vect <- rownames(Norm_features_matrix) %>% 
  enframe() %>% 
  select(cell_id = value) %>% 
  left_join(MPAs_df) %>% 
  mutate(mpa =  !is.na(mpa)) %>% pull(mpa)
```

# Ranking 

## Scorched Earth

### Initial conditions

```{r}
z <- 1

protected_cells <- matrix(MPA_vect, nrow = 1, ncol = nrow(Norm_features_matrix))

colnames(protected_cells) <- rownames(Norm_features_matrix)

unprotected_matrix <- Norm_features_matrix[!protected_cells, ,drop = F]

baseline_state <- protected_cells%*%Norm_features_matrix

protected_matrix <- Norm_features_matrix[protected_cells, , drop = T]

max_slopes <- z*matrixStats::colMaxs(unprotected_matrix, na.rm = T)^(z - 1)
```

### Run algorithm

```{r}
ranking <- list()

current_pick <- 0

current_state <- baseline_state

start_time <- Sys.time()

while(nrow(unprotected_matrix) > 0){
  
  current_pick <- sum(current_pick, 1)
  
  slopes <- pmin(max_slopes, z*current_state^(z - 1))
  
  delta <- unprotected_matrix%*%as.matrix(slopes)
  
  best_cell_indeces <- doBy::which.maxn(delta, n = 100)
  
  best_cells <- delta[best_cell_indeces, ] %>% 
    enframe() %>% 
    set_names(c('cell_id', "delta")) %>% 
    mutate(pick_order = current_pick)
  
  ranking <- bind_rows(ranking, best_cells)
  
  current_state <- unprotected_matrix[best_cell_indeces, , drop = F] %>% 
    colSums() +
    current_state
  
  unprotected_matrix <- unprotected_matrix[-best_cell_indeces, ,drop = F]
  
}

end_time <- Sys.time()

end_time - start_time

write_rds(ranking, 
          file.path(variant_dir, "ranking_results_se.rds"))
```

## With impacts

### Apply Impacts 

```{r}
abatable_impacts_df <- raster(file.path(emLab_project_dir, 
                                            "data", "02_processed", "impacts", "bottom_trawling", "abatable_impact_carbon_svr_log.tif")) %>% 
  raster::as.data.frame(xy = T) %>% 
  set_names(c("x", "y", "Ia")) %>% 
  inner_join(ocean_matrix) %>% 
  as_tibble() %>% 
  replace_na(list(Ia = 0)) %>% 
  filter(cell_id %in% rownames(Norm_features_matrix))

abatable_impacts_df <- abatable_impacts_df %>% 
  mutate(Ia = if_else(Ia == 0, 10^-6, Ia))

Norm_features_matrix_BAU <- Norm_features_matrix %>% 
  sweep(1, (1 - abatable_impacts_df$Ia), FUN = "*") 

remains_BAU <- colSums(Norm_features_matrix_BAU, na.rm = T)

1- remains_BAU

Norm_features_matrix_MPAs <- Norm_features_matrix %>% 
  sweep(1, 1, FUN = "*")

remains_MPA <- colSums(Norm_features_matrix_MPAs, na.rm = T)

Norm_features_matrix_diff <- Norm_features_matrix %>% 
  sweep(1, (abatable_impacts_df$Ia), FUN = "*")
```

### Initial conditions

```{r}
z <- 1

protected_cells <- matrix(MPA_vect, nrow = 1, ncol = nrow(Norm_features_matrix_diff))

colnames(protected_cells) <- rownames(Norm_features_matrix_diff)

unprotected_matrix <- Norm_features_matrix_diff[!protected_cells, ,drop = F]

baseline_state <- remains_BAU + protected_cells%*%Norm_features_matrix_diff

protected_matrix <- Norm_features_matrix_diff[protected_cells, ,drop =F]

max_slopes <- z*baseline_state^(z - 1)
```

### Run algorithm

```{r}
ranking <- list()

current_pick <- 0

current_state <- baseline_state

start_time <- Sys.time()

while(nrow(unprotected_matrix) > 0){
  
  current_pick <- sum(current_pick, 1)
  
  slopes <- pmin(max_slopes, z*current_state^(z - 1))
  
  delta <- unprotected_matrix%*%t(as.matrix(slopes))
  
  best_cell_indeces <- doBy::which.maxn(delta, n = 100)
  
  best_cells <- delta[best_cell_indeces, ] %>% 
    enframe() %>% 
    set_names(c('cell_id', "delta")) %>% 
    mutate(pick_order = current_pick)
  
  ranking <- bind_rows(ranking, best_cells)
  
  current_state <- unprotected_matrix[best_cell_indeces, , drop = F] %>% 
    colSums() +
    current_state
  
  unprotected_matrix <- unprotected_matrix[-best_cell_indeces, , drop = F]
  
}

end_time <- Sys.time()

end_time - start_time

write_rds(ranking, 
          file.path(variant_dir, "ranking_results.rds"))

ranking %>% 
  mutate(f = 1/n(),
         f = cumsum(f)) %>% 
  ggplot()+
  geom_line(aes(x = f, y = cumsum(delta)))
```

# Results

## Benefit Curves

### Scorched Earth 

```{r}
ranking_se <- read_rds(file.path(variant_dir, "ranking_results_se.rds"))

rep_curves_se <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(ranking_se) %>% 
  mutate(is_mpa = is.na(delta)) %>% 
  full_join(ocean_matrix) %>% 
  left_join(Norm_features_matrix %>% 
              as_tibble(rownames = NA) %>% 
              rownames_to_column("cell_id")) %>% 
  replace(is.na(.), 0) %>% 
  select(-x, -y, -cell_id, -delta, -pick_order) %>% 
  mutate_at(vars(-is_mpa), .funs = cumsum)

benefit_curve_se <- rep_curves_se %>% 
  select(-is_mpa) %>% 
  mutate_all(.funs = ~(.)^z) %>% 
  as.matrix() %>% 
  rowSums() %>% 
  as_tibble() %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) 

rep_curves_se %>% 
  select(is_mpa) %>% 
  bind_cols(benefit_curve_se) %>% 
  filter(!is_mpa) %>%
  ggplot()+
  geom_line(aes(x = fraction_protected, y = V/max(V))) +
  geom_vline(xintercept = c(0.1, 0.2, 0.3),
             linetype = c("dashed","dotted","dotdash"))+
  labs(x = "Fraction Ocean Protected",
       y = "Carbon Benefit",
       caption = "Carbon, scorched-earth")+
  theme_classic()+
  scale_y_continuous(limits = c(0, 1), expand = c(0,0))+
  scale_x_continuous(limits = c(0, 1), expand = c(0,0))+
  geom_rect(aes(xmin = 0, xmax = fraction_in_MPA, ymin = 0, ymax = 1), fill = "#69B578", alpha = 0.01)+
  ggsave(file.path(variant_dir, "benefit_curve_se.png"),
         width = 7, height = 4)
```

```{r}
avg_rep_curve_se <- rep_curves_se %>% 
  select(-is_mpa) %>% 
  as.matrix() %>% 
  rowSums() %>% 
  as_tibble() %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) 

rep_curves_se %>% 
  select(is_mpa) %>% 
  bind_cols(avg_rep_curve_se) %>% 
  filter(!is_mpa) %>%
  ggplot()+
  geom_line(aes(x = fraction_protected, y = V)) +
  geom_vline(xintercept = c(0.1, 0.2, 0.3),
             linetype = c("dashed","dotted","dotdash"))+
  labs(x = "Fraction Ocean Protected",
       y = "Average feature representation",
       caption = "Carbon, Scorched Earth")+
  theme_classic()+
  scale_y_continuous(limits = c(0, 1), expand = c(0,0))+
  scale_x_continuous(limits = c(0, 1), expand = c(0,0))+
  geom_rect(aes(xmin = 0, xmax = fraction_in_MPA, ymin = 0, ymax = 1), fill = "#69B578", alpha = 0.01)+
  ggsave(file.path(variant_dir, "rep_curve_se.png"),
         width = 7, height = 5)
```


### With Impacts 

```{r}
ranking<- read_rds(file.path(variant_dir, "ranking_results.rds"))

benefit_curve <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(ranking) %>% 
  mutate(is_mpa = is.na(delta))  %>% 
  full_join(ocean_matrix) %>% 
  left_join(Norm_features_matrix_diff %>% 
              as_tibble(rownames = NA) %>% 
              rownames_to_column("cell_id")) %>% 
  replace(is.na(.), 0) %>% 
  select(-x, -y, -delta, -pick_order, -is_mpa) %>% 
  mutate(cell_id = as.numeric(cell_id)) %>% 
  mutate_at(vars(-cell_id),.funs = cumsum) %>% 
  sweep(MARGIN = 2, c(0, remains_BAU), `+`) %>%
  mutate_at(vars(-cell_id), .funs = ~(.)^z) 

write_rds(benefit_curve, 
          file.path(variant_dir, "carbon_benefit.rds"))

benefit_curve %>% 
  select(-cell_id) %>% 
  rowSums() %>% 
  as_tibble() %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) %>% 
  filter(fraction_protected > fraction_in_MPA) %>% 
  ggplot()+
  geom_line(aes(x = fraction_protected, y = V)) +
  geom_vline(xintercept = c(0.1, 0.2, 0.3),
             linetype = c("dashed","dotted","dotdash"))+
  labs(x = "Fraction Ocean Protected",
       y = "Carbon Benefit",
       caption = "Carbon, with impacts")+
  theme_classic()+
  scale_x_continuous(limits = c(0,1), expand = c(0,0))+
  geom_rect(aes(xmin = 0, 
                xmax = fraction_in_MPA, 
                ymin = min(benefit_curve$V), 
                ymax = max(benefit_curve$V)),
                fill = "#69B578", alpha = 0.01)+
  ggsave(file.path(variant_dir, "benefit_curve.png"),
         width = 7, height = 4)
```

```{r}
rep_curves <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(ranking) %>% 
  mutate(is_mpa = is.na(delta)) %>% 
  full_join(ocean_matrix) %>% 
  left_join(Norm_features_matrix %>% 
              as_tibble(rownames = NA) %>% 
              rownames_to_column("cell_id")) %>% 
  replace(is.na(.), 0) %>% 
  select(-x, -y, -cell_id, -delta, -pick_order, -is_mpa) %>% 
  mutate_all(.funs = cumsum) 
  
rep_curves %>% 
  as.matrix() %>% 
  rowSums() %>% 
  as_tibble() %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) %>% 
  filter(fraction_protected > fraction_in_MPA) %>% 
  ggplot()+
  geom_line(aes(x = fraction_protected, y = V)) +
  geom_vline(xintercept = c(0.1, 0.2, 0.3),
             linetype = c("dashed","dotted","dotdash"))+
  labs(x = "Fraction Ocean Protected",
       y = "Average feature representation",
       caption = "carbon, with impacts")+
  theme_classic()+
  scale_x_continuous(limits = c(0,1), expand = c(0,0))+
  geom_rect(aes(xmin = 0, 
                xmax = fraction_in_MPA, 
                ymin = 0,
                ymax = 1),
            fill = "#69B578", alpha = 0.01)+
  ggsave(file.path(variant_dir, "rep_curve.png"),
         width = 7, height = 4)
```


## Priority Maps

```{r}
ranking_raster_se <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(ranking_se) %>% 
  mutate(is_mpa = is.na(delta)) %>% 
  left_join(ocean_matrix) %>% 
  tibble::rowid_to_column("ID") %>% 
  mutate(rank = 1- scales::rescale(ID)) %>% 
  select(x, y, rank) %>% 
  rasterFromXYZ(crs = crs(ocean_low_res_moll)) %>% 
  projectRaster(ocean_low_res_moll) %>% 
  mask(ocean_low_res_moll) 

writeRaster(ranking_raster_se, 
            file.path(variant_dir, "ranking_raster_se.tif"), 
            overwrite = T)

ranking_raster <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(ranking) %>% 
  mutate(is_mpa = is.na(delta)) %>% 
  left_join(ocean_matrix) %>% 
  tibble::rowid_to_column("ID") %>% 
  mutate(rank = 1- scales::rescale(ID)) %>% 
  select(x, y, rank) %>% 
  rasterFromXYZ(crs = crs(ocean_low_res_moll)) %>% 
  projectRaster(ocean_low_res_moll) %>% 
  mask(ocean_low_res_moll) 

writeRaster(ranking_raster, 
            file.path(variant_dir, "ranking_raster.tif"), 
            overwrite = T)
```

```{r}
dv_raster <- benefit_curve %>% 
  select(-cell_id) %>% 
  as_tibble() %>% 
  set_names("V") %>% 
  bind_cols(benefit_curve %>% 
              transmute(cell_id = as.character(cell_id))) %>% 
  left_join(ocean_matrix) %>% 
  mutate(dV = V - lag(V)) %>% 
  replace_na(list(dV = 0)) %>% 
  select(x,y,dV) %>% 
  rasterFromXYZ(crs = crs(ocean_low_res_moll)) %>% 
  projectRaster(ocean_low_res_moll) %>% 
  mask(ocean_low_res_moll) 

writeRaster(dv_raster, 
            file.path(variant_dir, "delta_v_raster.tif"), 
            overwrite = T)
```

```{r}
ranking_raster_se <- raster(file.path(variant_dir, "ranking_raster_se.tif"))

ranking_map_se <- tmap::tm_shape(ranking_raster_se)+
  tmap::tm_raster(title = "",
                  palette = viridis::viridis(n = 8),
                  breaks = priorities_map_pal$breaks,
                  labels = priorities_map_pal$labels,
                  legend.is.portrait = T, 
                  legend.reverse = T)+
  tmap::tm_shape(land_50_moll)+
  tmap::tm_fill(col = "grey", border.col = "transparent")+
  tmap::tm_credits("Scorched earth") +
  tmap::tm_shape(MPAs)+
  tmap::tm_raster(col = "fully_highly_protected_reviewed_MPAs", palette = "#00A600FF", legend.show	= F)+
  tmap::tm_layout(title = "Carbon Conservation Priorities",
                  title.position = c("center", .95),
                  inner.margins = c(0.12, 0, 0.08, 0.04), 
                  frame = F,
                  legend.position = c(.99, "center")) 

tmap::tmap_save(ranking_map_se, 
                filename = file.path(variant_dir, "priority_map_se.png"),
                dpi = 300,  height = 4, width = 8 )

tmap::tmap_save(ranking_map_se, 
                filename = file.path(emLab_project_dir, "figures", "supplement", "carbon_priorities_scorched_earth.png"),
                dpi = 300,  height = 4, width = 8 )

ranking_raster <- raster(file.path(variant_dir, "ranking_raster.tif"))

ranking_map <- tmap::tm_shape(ranking_raster)+
  tmap::tm_raster(title = "",
                  palette  = viridis::viridis(n = 8),
                  breaks = priorities_map_pal$breaks,
                  labels = priorities_map_pal$labels,
                  legend.is.portrait = T, 
                  legend.reverse = T)+
  tmap::tm_shape(land_50_moll)+
  tmap::tm_fill(col = "grey", border.col = "transparent")+
  tmap::tm_credits("With impacts") +
  tmap::tm_shape(MPAs)+
  tmap::tm_raster(col = "fully_highly_protected_reviewed_MPAs", palette = "#00A600FF", legend.show	= F)+
  tmap::tm_layout(title = "Carbon Conservation Priorities",
                  title.position = c("center", .95),
                  inner.margins = c(0.12, 0, 0.08, 0.04), 
                  frame = F,
                  legend.position = c(.99, "center")) 

tmap::tmap_save(ranking_map, 
                filename = file.path(variant_dir, "priority_map.png"),
                dpi = 300,  height = 4, width = 8 )
```

```{r}
ranking_diff <- ranking_raster_se - ranking_raster

ranking_diff %>% 
  as.data.frame(xy = TRUE, centroids = T) %>% 
  set_names(c("lon", "lat", "diff_rank")) %>% 
  filter(!is.na(diff_rank)) %>% 
  ggplot()+
  geom_raster(aes(x = lon, y = lat, fill = diff_rank))+
  geom_raster(data = MPAs_df %>% 
                filter(!is.na(mpa)), aes(x = x, y = y), fill = "lightblue")+
  geom_sf(data = land_50_moll, col = "transparent", fill = "grey", inherit.aes = FALSE)+
  scale_fill_continuous_diverging(rev = T,
                                  guide = guide_colorbar(title = "ranking difference",
                                                         title.position = 'bottom',
                                                         direction = "horizontal",
                                                         title.hjust = 0.5,
                                                         label.hjust = -.02,
                                                         keywidth = 0.3))+
  theme(legend.text = element_text(size = 6),
        legend.text.align = 0,
        legend.position = "bottom",
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_line(color = "#ebebe5", size = 0.2),
        panel.grid.minor = element_blank(),
        plot.background = element_rect(fill = NA, color = NA), 
        panel.background = element_rect(fill = NA, color = NA), 
        legend.background = element_rect(fill = NA, color = NA),
        panel.border = element_blank())+
  ggsave(filename = file.path(emLab_project_dir, "figures", "supplement", "carbon_priorities_diff_scorched_earth.png"), 
         dpi = 300,  height = 9, width = 8)
```

# What does it take to get 90% of the benefits?

```{r}
benefit_curve <- read_rds(file.path(variant_dir, "carbon_benefit.rds"))

carbon_benefit_bau <- benefit_curve %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected),
         V_rescaled = scales::rescale(carbon)) %>% 
  slice(which.min(abs(fraction_protected - fraction_in_MPA))) %>% 
  pull(V_rescaled)

ninety_percent_delta_carbon <- carbon_benefit_bau + 0.9*(1 - carbon_benefit_bau)

(f_to_90 <- benefit_curve %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected),
         V_rescaled = scales::rescale(carbon)) %>% 
  slice(which.min(abs(V_rescaled - ninety_percent_delta_carbon))) %>% 
  pull(fraction_protected)) # 8.7 %

# additional protection 6.2%
f_to_90 - fraction_in_MPA
```
#  EEZ vs HS

How much of the area neede to reach 90% beneftis is in EEZ vs HS

```{r}
territories_mask <- raster(file.path(emLab_project_dir, "data", "02_processed", "masks", "territories.tif")) # 1 equals high seas

territories_df <- read_csv(file.path(emLab_project_dir, "data", "02_processed", "masks", "territories_lookup.csv"))

eez_mask <- ocean_low_res_moll %>% 
  mask(territories_mask) %>% 
  mask(territories_mask == 1926, maskvalue = 1) # remove Antartica

hs_mask <- ocean_low_res_moll %>% 
  mask(eez_mask, inverse = T) 

eez_mask_df_moll <- eez_mask %>% 
  as.data.frame(xy = TRUE) %>% 
  set_names(c("x", "y", "eez_mask")) %>% 
  mutate(eez_mask = ifelse(eez_mask == 1, 1, NA)) %>% 
  filter(!is.na(eez_mask)) %>% 
  inner_join(ocean_matrix)
```

```{r}
benefit_curve %>% 
  mutate(in_EEZ = cell_id %in% eez_mask_df_moll$cell_id) %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) %>% 
  filter(fraction_protected <= f_to_90) %>% 
  group_by(in_EEZ) %>% 
  summarise(n_cells = n()) %>% 
  mutate(f = n_cells/sum(n_cells))
```

How much of the top 10% is in EEZ?

```{r}
benefit_curve %>% 
  mutate(in_EEZ = cell_id %in% eez_mask_df_moll$cell_id) %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) %>% 
  filter(fraction_protected <= 0.1) %>% 
  group_by(in_EEZ) %>% 
  summarise(n_cells = n()) %>% 
  ungroup() %>% 
  mutate(f_cell = n_cells/sum(n_cells))
```

## rescaled maps

```{r}
ranking_raster <- raster(file.path(variant_dir, "ranking_raster.tif"))
  
HS_MPAs_df <- MPAs %>% 
  mask(hs_mask, mask_value = 0, updatevalue = NA) %>% 
  raster::as.data.frame(xy = T) %>%
  inner_join(ocean_matrix) %>% 
  select(lon = x, lat = y, cell_id, mpa = fully_highly_protected_reviewed_MPAs) %>%
  as_tibble() 

HS_carbon_ranking_df <- ranking_raster %>% 
  mask(hs_mask, mask_value = 0, updatevalue = NA) %>% 
  as.data.frame(xy = TRUE) %>% 
  set_names(c("lon", "lat", "rank")) %>% 
  filter(!is.na(rank)) %>% 
  left_join(HS_MPAs_df) %>% 
  mutate(rank = 100*rank)

fraction_HS_protected <- length(HS_carbon_ranking_df$rank[!is.na(HS_carbon_ranking_df$mpa)])/length(HS_carbon_ranking_df$rank)

min_HS_MPA_rank <- min(HS_carbon_ranking_df$rank[!is.na(HS_carbon_ranking_df$mpa)])

HS_breaks <- sort(unique(c(0, 50, 60, 70, 80, 85, 90, 95, min_HS_MPA_rank, 100)))

HS_labels <- HS_breaks[-1]

HS_colors <- priorities_map_pal$colors
HS_colors <- viridis::viridis(n = 8)

HS_colors[HS_labels > min_HS_MPA_rank] <- "lightblue"

HS_labels <- 100 - HS_labels

HS_labels[HS_labels ==  (100-min_HS_MPA_rank)] <- " "

HS_legend_key_widths <- tibble(r = as.numeric(as.character(HS_breaks[-1]/100)),
                                  r_lag = lag(r)) %>% 
  replace_na(list(r_lag = 0)) %>%
  mutate(l = r - r_lag) %>% 
  mutate(w = ifelse(r <= 0.7, l/0.7*0.2, l/0.3*0.4),
         w1 = l*0.6) %>% pull(w1)

HS_legend_key_widths[length(HS_legend_key_widths)] <- HS_legend_key_widths[length(HS_legend_key_widths)]/4

HS_carbon_map <- HS_carbon_ranking_df %>% 
  mutate(rank_cut = cut(rank,
                        breaks = raster::quantile(HS_carbon_ranking_df$rank,
                                                  probs = c(0.00, 0.20, 0.50 ,0.70, 0.80,0.85, 0.90, 0.95, 1 - fraction_HS_protected, 1.00)),
                        labels = HS_labels,
                        include.lowest = T)) %>% 
  ggplot()+
  geom_raster(aes(x = lon, y = lat, fill = rank_cut))+
  geom_raster(data = HS_MPAs_df %>% 
                filter(!is.na(mpa)), aes(x = lon, y = lat), fill = "lightblue")+
  geom_sf(data = land_50_moll, col = "transparent", fill = "grey", inherit.aes = FALSE)+
  scale_fill_manual(values  = HS_colors,
                    guide = guide_legend(
                      title = "Top % of the High Seas",
                      direction = "horizontal",
                      keyheight = unit(0.01, "npc"),
                      keywidth = unit(rev(HS_legend_key_widths)/1.5, "npc"),
                      title.position = 'top',
                      title.hjust = 0.5,
                      label.hjust = -.02,
                      nrow = 1,
                      byrow = T,
                      reverse = T,
                      label.position = "bottom"))+
  theme(legend.text = element_text(size = 6),
        legend.text.align = 0,
        legend.position = "bottom",
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_line(color = "#ebebe5", size = 0.2),
        panel.grid.minor = element_blank(),
        plot.background = element_rect(fill = NA, color = NA), 
        panel.background = element_rect(fill = NA, color = NA), 
        legend.background = element_rect(fill = NA, color = NA),
        panel.border = element_blank())
```

```{r}
EEZ_MPAs_df <- MPAs %>% 
  mask(hs_mask, mask_value = 0, updatevalue = NA, inverse = T) %>% 
  raster::as.data.frame(xy = T) %>%
  inner_join(ocean_matrix) %>% 
  select(lon = x, lat = y, cell_id, mpa = fully_highly_protected_reviewed_MPAs) %>%
  as_tibble() 

EEZ_carbon_ranking_df <- ranking_raster %>% 
  mask(hs_mask, mask_value = 0, updatevalue = NA, inverse = T) %>% 
  as.data.frame(xy = TRUE) %>% 
  set_names(c("lon", "lat", "rank")) %>% 
  filter(!is.na(rank)) %>% 
  left_join(EEZ_MPAs_df) %>% 
  mutate(rank = 100*rank)

fraction_EEZ_protected <- length(EEZ_carbon_ranking_df$rank[!is.na(EEZ_carbon_ranking_df$mpa)])/length(EEZ_carbon_ranking_df$rank)

min_EEZ_MPA_rank <- min(EEZ_carbon_ranking_df$rank[!is.na(EEZ_carbon_ranking_df$mpa)])

EEZ_breaks <- sort(unique(c(0, 50, 60, 70, 80, 85, 90, 95, min_EEZ_MPA_rank, 100)))

EEZ_labels <- EEZ_breaks[-1]

EEZ_colors <- priorities_map_pal$colors
EEZ_colors <- viridis::viridis(n = 8)

EEZ_colors[EEZ_labels > min_EEZ_MPA_rank] <- "lightblue"

EEZ_labels <- 100 - EEZ_labels

EEZ_labels[EEZ_labels ==  (100-min_EEZ_MPA_rank)] <- " "

EEZ_legend_key_widths <- tibble(r = as.numeric(as.character(EEZ_breaks[-1]/100)),
                                  r_lag = lag(r)) %>% 
  replace_na(list(r_lag = 0)) %>%
  mutate(l = r - r_lag) %>% 
  mutate(w = ifelse(r <= 0.7, l/0.7*0.2, l/0.3*0.4),
         w1 = l*0.6) %>% pull(w1)

EEZ_legend_key_widths[length(EEZ_legend_key_widths)] <- 0.03*0.9
EEZ_legend_key_widths[length(EEZ_legend_key_widths) -1] <- 0.03 - 0.03*0.9

EEZ_carbon_map <- EEZ_carbon_ranking_df %>% 
  mutate(rank_cut = cut(rank,
                        breaks = raster::quantile(EEZ_carbon_ranking_df$rank,
                                                  probs = c(0.00, 0.20, 0.50 ,0.70, 0.80,0.85, 0.90, 0.95, 1 - fraction_EEZ_protected, 1.00)),
                        labels = EEZ_labels,
                        include.lowest = T)) %>% 
  ggplot()+
  geom_raster(aes(x = lon, y = lat, fill = rank_cut))+
  geom_raster(data = EEZ_MPAs_df %>% 
                filter(!is.na(mpa)), aes(x = lon, y = lat), fill = "lightblue")+
  geom_sf(data = land_50_moll, col = "transparent", fill = "grey", inherit.aes = FALSE)+
  scale_fill_manual(values  = EEZ_colors,
                    guide = guide_legend(
                      title = "Top % of the Exclusive Economic Zones",
                      direction = "horizontal",
                      keyheight = unit(0.01, "npc"),
                      keywidth = unit(rev(EEZ_legend_key_widths)/1.5, "npc"),
                      title.position = 'top',
                      title.hjust = 0.5,
                      label.hjust = -.02,
                      nrow = 1,
                      byrow = T,
                      reverse = T,
                      label.position = "bottom"))+
  theme(legend.text = element_text(size = 6),
        legend.text.align = 0,
        legend.position = "bottom",
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_line(color = "#ebebe5", size = 0.2),
        panel.grid.minor = element_blank(),
        plot.background = element_rect(fill = NA, color = NA), 
        panel.background = element_rect(fill = NA, color = NA), 
        legend.background = element_rect(fill = NA, color = NA),
        panel.border = element_blank())
```

```{r}
carbon_eez_vs_hs_map <- EEZ_carbon_map + 
  HS_carbon_map + 
  plot_annotation(tag_levels = "a")+
  plot_layout(ncol = 1) +
  theme(plot.margin = grid::unit(c(0,0,0,0), "mm")) 

ggsave(carbon_eez_vs_hs_map, filename = file.path(emLab_project_dir, "figures", "supplement", "carbon_eez_vs_hs.pdf"), dpi = 300,  height = 9, width = 8)
ggsave(carbon_eez_vs_hs_map, filename = file.path(emLab_project_dir, "figures", "supplement", "carbon_eez_vs_hs.png"), dpi = 300,  height = 9, width = 8)
```




