---
title: "Carbon in Juanation"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
---

```{r}
library(raster)
library(sf)
library(tidyverse)
library(patchwork)

source(here::here("common.R"))
source(here::here("functions", "map_priorities.R"))

variant_name <- "06_carbon"

variant_dir <- file.path(emLab_project_dir,"data", "03_output", variant_name)

dir.create(variant_dir)
```

# Load features data 

```{r}
carbon_features_info <- tibble(filepath = file.path(emLab_project_dir, 
                                                    "data", "02_processed", "carbon", "carbon_lr_mean_moll.tif"),
                               w = 1)
```

```{r, eval = F}
carbon_features_df <- stack(carbon_features_info$filepath) %>%
  raster::as.data.frame(xy = T) %>%
  inner_join(ocean_matrix) %>%
  select(x, y, cell_id, everything()) %>%
  rename(carbon = layer) %>% 
  as_tibble()

data.table::fwrite(carbon_features_df, 
                   file.path(emLab_project_dir, "data", "02_processed", "features_df", "carbon_df.csv"))
```

```{r}
features_df <- data.table::fread(file.path(emLab_project_dir, "data", "02_processed", "features_df", "carbon_df.csv")) %>% 
  as_tibble()
```

## Normalize and trim

```{r}
features_matrix <- features_df %>% 
  select(-x,-y,-cell_id) %>% 
  as.matrix()

stopifnot(
  sum(map_lgl(features_df, is.numeric)) == ncol(features_df) 
  ) # ALL numeric?

Norm_features_matrix <- sweep(features_matrix, 2, colSums(features_matrix, na.rm = T), FUN = "/")

stopifnot(
  sum(colSums(Norm_features_matrix, na.rm = T)) == ncol(features_matrix)
  )  # Normalized OK?

rownames(Norm_features_matrix) <- features_df$cell_id

Norm_features_matrix <- Norm_features_matrix[rowSums(is.na(Norm_features_matrix)) != ncol(Norm_features_matrix), ] 

Norm_features_matrix <- as.matrix(Norm_features_matrix)

colnames(Norm_features_matrix) <- "carbon"

stopifnot(
  identical(colnames(Norm_features_matrix), 
          features_df %>% 
            select(-x,-y,-cell_id) %>% 
            colnames())
  )  # Is the order of the features mantained?
```

# Current MPAs

```{r}
MPAs <- raster(file.path(emLab_project_dir, "data", "02_processed", "masks", "fully_highly_protected_reviewed_MPAs.tif"))

MPAs_df <- MPAs %>% 
  raster::as.data.frame(xy = T) %>%
  inner_join(ocean_matrix) %>% 
  select(x, y, cell_id, mpa = fully_highly_protected_reviewed_MPAs) %>%
  as_tibble() 

fraction_in_MPA <- sum(!is.na(MPAs_df$mpa))/nrow(MPAs_df)
  
MPA_vect <- rownames(Norm_features_matrix) %>% 
  enframe() %>% 
  select(cell_id = value) %>% 
  left_join(MPAs_df) %>% 
  mutate(mpa =  !is.na(mpa)) %>% pull(mpa)
```

# Ranking 

## Scorched Earth

### Initial conditions

```{r}
z <- 1

protected_cells <- matrix(MPA_vect, nrow = 1, ncol = nrow(Norm_features_matrix))

colnames(protected_cells) <- rownames(Norm_features_matrix)

unprotected_matrix <- Norm_features_matrix[!protected_cells, ,drop = F]

baseline_state <- protected_cells%*%Norm_features_matrix

protected_matrix <- Norm_features_matrix[protected_cells, , drop = T]

max_slopes <- z*matrixStats::colMaxs(unprotected_matrix, na.rm = T)^(z - 1)
```

### Run algorithm

```{r}
ranking <- list()

current_pick <- 0

current_state <- baseline_state

start_time <- Sys.time()

while(nrow(unprotected_matrix) > 0){
  
  current_pick <- sum(current_pick, 1)
  
  slopes <- pmin(max_slopes, z*current_state^(z - 1))
  
  delta <- unprotected_matrix%*%as.matrix(slopes)
  
  best_cell_indeces <- doBy::which.maxn(delta, n = 100)
  
  best_cells <- delta[best_cell_indeces, ] %>% 
    enframe() %>% 
    set_names(c('cell_id', "delta")) %>% 
    mutate(pick_order = current_pick)
  
  ranking <- bind_rows(ranking, best_cells)
  
  current_state <- unprotected_matrix[best_cell_indeces, , drop = F] %>% 
    colSums() +
    current_state
  
  unprotected_matrix <- unprotected_matrix[-best_cell_indeces, ,drop = F]
  
}

end_time <- Sys.time()

end_time - start_time

write_rds(ranking, 
          file.path(variant_dir, "ranking_results_se.rds"))
```

## With impacts

### Apply Impacts 

```{r}
# abatable_impacts_df <- raster(file.path(emLab_project_dir, 
#                                             "data", "02_processed", "impacts", "bottom_trawling", "abatable_impact_carbon_svr_log.tif")) %>% 
#   raster::as.data.frame(xy = T) %>% 
#   set_names(c("x", "y", "Ia")) %>% 
#   inner_join(ocean_matrix) %>% 
#   as_tibble() %>% 
#   replace_na(list(Ia = 0)) %>% 
#   filter(cell_id %in% rownames(Norm_features_matrix))
# 
# abatable_impacts_df <- abatable_impacts_df %>% 
#   mutate(Ia = if_else(Ia == 0, 10^-6, Ia))
# 
# Norm_features_matrix_BAU <- Norm_features_matrix %>% 
#   sweep(1, (1 - abatable_impacts_df$Ia), FUN = "*") 
# 
# remains_BAU <- colSums(Norm_features_matrix_BAU, na.rm = T)
# 
# 1- remains_BAU
# 
# Norm_features_matrix_MPAs <- Norm_features_matrix %>% 
#   sweep(1, 1, FUN = "*")
# 
# remains_MPA <- colSums(Norm_features_matrix_MPAs, na.rm = T)
# 
# Norm_features_matrix_diff <- Norm_features_matrix %>% 
#   sweep(1, (abatable_impacts_df$Ia), FUN = "*")
```

```{r}
Ia <- raster(file.path(emLab_project_dir, "data", "02_processed", "impacts", "bottom_trawling", "Ia.tif"))

abatable_impacts_df <- Ia %>% 
  raster::as.data.frame(xy = T) %>% 
  set_names(c("x", "y", "Ia")) %>% 
  inner_join(ocean_matrix) %>% 
  as_tibble() %>% 
  replace_na(list(Ia = 0)) %>% 
  filter(cell_id %in% rownames(Norm_features_matrix))

abatable_impacts_df <- abatable_impacts_df %>% 
  mutate(Ia = if_else(Ia == 0, 10^-30, Ia))

Norm_features_matrix_BAU <- Norm_features_matrix %>% 
  sweep(1, (1 - abatable_impacts_df$Ia), FUN = "*") 

remains_BAU <- colSums(Norm_features_matrix_BAU, na.rm = T)

1 - remains_BAU

Norm_features_matrix_MPAs <- Norm_features_matrix %>% 
  sweep(1, 1, FUN = "*")

remains_MPA <- colSums(Norm_features_matrix_MPAs, na.rm = T)

Norm_features_matrix_diff <- Norm_features_matrix %>% 
  sweep(1, (abatable_impacts_df$Ia), FUN = "*")
```

### Initial conditions

```{r}
z <- 1

protected_cells <- matrix(MPA_vect, nrow = 1, ncol = nrow(Norm_features_matrix_diff))

colnames(protected_cells) <- rownames(Norm_features_matrix_diff)

unprotected_matrix <- Norm_features_matrix_diff[!protected_cells, ,drop = F]

baseline_state <- remains_BAU + protected_cells%*%Norm_features_matrix_diff

protected_matrix <- Norm_features_matrix_diff[protected_cells, ,drop =F]

max_slopes <- z*baseline_state^(z - 1)
```

### Run algorithm

```{r}
ranking <- list()

current_pick <- 0

current_state <- baseline_state

start_time <- Sys.time()

while(nrow(unprotected_matrix) > 0){
  
  current_pick <- sum(current_pick, 1)
  
  slopes <- pmin(max_slopes, z*current_state^(z - 1))
  
  delta <- unprotected_matrix%*%t(as.matrix(slopes))
  
  best_cell_indeces <- doBy::which.maxn(delta, n = 100)
  
  best_cells <- delta[best_cell_indeces, ] %>% 
    enframe() %>% 
    set_names(c('cell_id', "delta")) %>% 
    mutate(pick_order = current_pick)
  
  ranking <- bind_rows(ranking, best_cells)
  
  current_state <- unprotected_matrix[best_cell_indeces, , drop = F] %>% 
    colSums() +
    current_state
  
  unprotected_matrix <- unprotected_matrix[-best_cell_indeces, , drop = F]
  
}

end_time <- Sys.time()

end_time - start_time

write_rds(ranking, 
          file.path(variant_dir, "ranking_results_updated.rds"))
```

# Results

## Benefit Curves

### Scorched Earth 

```{r}
ranking_se <- read_rds(file.path(variant_dir, "ranking_results_se.rds"))

rep_curves_se <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(ranking_se) %>% 
  mutate(is_mpa = is.na(delta)) %>% 
  full_join(ocean_matrix) %>% 
  left_join(Norm_features_matrix %>% 
              as_tibble(rownames = NA) %>% 
              rownames_to_column("cell_id")) %>% 
  replace(is.na(.), 0) %>% 
  select(-x, -y, -cell_id, -delta, -pick_order) %>% 
  mutate_at(vars(-is_mpa), .funs = cumsum)

benefit_curve_se <- rep_curves_se %>% 
  select(-is_mpa) %>% 
  mutate_all(.funs = ~(.)^z) %>% 
  as.matrix() %>% 
  rowSums() %>% 
  as_tibble() %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) 

rep_curves_se %>% 
  select(is_mpa) %>% 
  bind_cols(benefit_curve_se) %>% 
  filter(!is_mpa) %>%
  ggplot()+
  geom_line(aes(x = fraction_protected, y = V/max(V))) +
  geom_vline(xintercept = c(0.1, 0.2, 0.3),
             linetype = c("dashed","dotted","dotdash"))+
  labs(x = "Fraction Ocean Protected",
       y = "Carbon Benefit",
       caption = "Carbon, scorched-earth")+
  theme_classic()+
  scale_y_continuous(limits = c(0, 1), expand = c(0,0))+
  scale_x_continuous(limits = c(0, 1), expand = c(0,0))+
  geom_rect(aes(xmin = 0, xmax = fraction_in_MPA, ymin = 0, ymax = 1), fill = "#69B578", alpha = 0.01)+
  ggsave(file.path(variant_dir, "benefit_curve_se.png"),
         width = 7, height = 4)
```

```{r}
avg_rep_curve_se <- rep_curves_se %>% 
  select(-is_mpa) %>% 
  as.matrix() %>% 
  rowSums() %>% 
  as_tibble() %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) 

rep_curves_se %>% 
  select(is_mpa) %>% 
  bind_cols(avg_rep_curve_se) %>% 
  filter(!is_mpa) %>%
  ggplot()+
  geom_line(aes(x = fraction_protected, y = V)) +
  geom_vline(xintercept = c(0.1, 0.2, 0.3),
             linetype = c("dashed","dotted","dotdash"))+
  labs(x = "Fraction Ocean Protected",
       y = "Average feature representation",
       caption = "Carbon, Scorched Earth")+
  theme_classic()+
  scale_y_continuous(limits = c(0, 1), expand = c(0,0))+
  scale_x_continuous(limits = c(0, 1), expand = c(0,0))+
  geom_rect(aes(xmin = 0, xmax = fraction_in_MPA, ymin = 0, ymax = 1), fill = "#69B578", alpha = 0.01)+
  ggsave(file.path(variant_dir, "rep_curve_se.png"),
         width = 7, height = 5)
```


### With Impacts 

```{r}
ranking <- read_rds(file.path(variant_dir, "ranking_results_updated.rds"))

total_remineralized_co2 <- 1.47*10^9

benefit_curve <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(ranking) %>% 
  mutate(is_mpa = is.na(delta))  %>% 
  full_join(ocean_matrix) %>% 
  left_join(Norm_features_matrix_diff %>% 
              as_tibble(rownames = NA) %>% 
              rownames_to_column("cell_id")) %>% 
  replace(is.na(.), 0) %>% 
  select(-x, -y, -delta, -pick_order, -is_mpa) %>% 
  replace(is.na(.), 0) %>% 
  mutate(cell_id = as.numeric(cell_id),
         carbon = cumsum(carbon),
         carbon = scales::rescale(carbon)*total_remineralized_co2)

benefit_curve %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) %>% 
  filter(fraction_protected > fraction_in_MPA) %>% 
  ggplot()+
  geom_line(aes(x = fraction_protected, y = carbon/10^9))+
  geom_vline(xintercept = c(0.1, 0.2, 0.3),
             linetype = c("dashed","dotted","dotdash"))+
  labs(x = "Fraction Ocean Protected",
       y = "Carbon Benefit (x10^9 Mg CO2)",
       caption = "Carbon, with impacts")+
  theme_classic()+
  scale_x_continuous(limits = c(0,1), expand = c(0,0))+
  geom_rect(aes(xmin = 0, 
                xmax = fraction_in_MPA, 
                ymin = 0, 
                ymax = max(carbon/10^9)),
            fill = "#69B578", alpha = 0.01)+
  ggsave(file.path(variant_dir, "benefit_curve_updated.png"),
         width = 7, height = 4)

write_rds(benefit_curve, 
          file.path(variant_dir, "carbon_benefit_updated.rds"))
```

```{r}
# ranking <- read_rds(file.path(variant_dir, "ranking_results.rds"))
# 
# benefit_curve <- MPAs_df %>%
#   filter(!is.na(mpa)) %>%
#   select(cell_id) %>%
#   bind_rows(ranking) %>%
#   mutate(is_mpa = is.na(delta))  %>%
#   full_join(ocean_matrix) %>%
#   left_join(Norm_features_matrix_diff %>%
#               as_tibble(rownames = NA) %>%
#               rownames_to_column("cell_id")) %>%
#   replace(is.na(.), 0) %>%
#   select(-x, -y, -delta, -pick_order, -is_mpa) %>%
#   mutate(cell_id = as.numeric(cell_id)) %>%
#   mutate_at(vars(-cell_id),.funs = cumsum) %>%
#   sweep(MARGIN = 2, c(0, remains_BAU), `+`) %>%
#   mutate_at(vars(-cell_id), .funs = ~(.)^z)
# 
# write_rds(benefit_curve, 
#           file.path(variant_dir, "carbon_benefit.rds"))
# 
# benefit_curve %>% 
#   select(-cell_id) %>% 
#   rowSums() %>% 
#   as_tibble() %>% 
#   set_names("V") %>% 
#   mutate(fraction_protected = 1/n(),
#          fraction_protected = cumsum(fraction_protected)) %>% 
#   filter(fraction_protected > fraction_in_MPA) %>% 
#   ggplot()+
#   geom_line(aes(x = fraction_protected, y = V)) +
#   geom_vline(xintercept = c(0.1, 0.2, 0.3),
#              linetype = c("dashed","dotted","dotdash"))+
#   labs(x = "Fraction Ocean Protected",
#        y = "Carbon Benefit",
#        caption = "Carbon, with impacts")+
#   theme_classic()+
#   scale_x_continuous(limits = c(0,1), expand = c(0,0))+
#   geom_rect(aes(xmin = 0, 
#                 xmax = fraction_in_MPA, 
#                 ymin = min(benefit_curve$V), 
#                 ymax = max(benefit_curve$V)),
#                 fill = "#69B578", alpha = 0.01)
#   ggsave(file.path(variant_dir, "benefit_curve.png"),
#          width = 7, height = 4)
```

```{r}
rep_curves <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(ranking) %>% 
  mutate(is_mpa = is.na(delta)) %>% 
  full_join(ocean_matrix) %>% 
  left_join(Norm_features_matrix %>% 
              as_tibble(rownames = NA) %>% 
              rownames_to_column("cell_id")) %>% 
  replace(is.na(.), 0) %>% 
  select(-x, -y, -cell_id, -delta, -pick_order, -is_mpa) %>% 
  mutate_all(.funs = cumsum) 
  
rep_curves %>% 
  as.matrix() %>% 
  rowSums() %>% 
  as_tibble() %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) %>% 
  filter(fraction_protected > fraction_in_MPA) %>% 
  ggplot()+
  geom_line(aes(x = fraction_protected, y = V)) +
  geom_vline(xintercept = c(0.1, 0.2, 0.3),
             linetype = c("dashed","dotted","dotdash"))+
  labs(x = "Fraction Ocean Protected",
       y = "Average feature representation",
       caption = "carbon, with impacts")+
  theme_classic()+
  scale_x_continuous(limits = c(0,1), expand = c(0,0))+
  geom_rect(aes(xmin = 0, 
                xmax = fraction_in_MPA, 
                ymin = 0,
                ymax = 1),
            fill = "#69B578", alpha = 0.01)+
  ggsave(file.path(variant_dir, "rep_curve_updated.png"),
         width = 7, height = 4)
```

## Priority Maps

```{r}
ranking_raster_se <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(ranking_se) %>% 
  mutate(is_mpa = is.na(delta)) %>% 
  left_join(ocean_matrix) %>% 
  tibble::rowid_to_column("ID") %>% 
  mutate(rank = 1- scales::rescale(ID)) %>% 
  select(x, y, rank) %>% 
  rasterFromXYZ(crs = crs(ocean_low_res_moll)) %>% 
  projectRaster(ocean_low_res_moll) %>% 
  mask(ocean_low_res_moll) 

writeRaster(ranking_raster_se, 
            file.path(variant_dir, "ranking_raster_se.tif"), 
            overwrite = T)

ranking_raster <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(ranking) %>% 
  mutate(is_mpa = is.na(delta)) %>% 
  left_join(ocean_matrix) %>% 
  tibble::rowid_to_column("ID") %>% 
  mutate(rank = 1- scales::rescale(ID)) %>% 
  select(x, y, rank) %>% 
  rasterFromXYZ(crs = crs(ocean_low_res_moll)) %>% 
  projectRaster(ocean_low_res_moll) %>% 
  mask(ocean_low_res_moll) 

writeRaster(ranking_raster, 
            file.path(variant_dir, "ranking_raster_updated.tif"), 
            overwrite = T)
```

```{r}
dv_raster <- benefit_curve %>%
  select(carbon) %>%
  as_tibble() %>%
  set_names("V") %>%
  bind_cols(benefit_curve %>%
              transmute(cell_id = as.character(cell_id))) %>%
  left_join(ocean_matrix) %>%
  mutate(dV = V - lag(V)) %>%
  replace_na(list(dV = 0)) %>%
  select(x,y,dV) %>%
  rasterFromXYZ(crs = crs(ocean_low_res_moll)) %>%
  projectRaster(ocean_low_res_moll) %>%
  mask(ocean_low_res_moll)

plot(dv_raster)

writeRaster(dv_raster, 
            file.path(variant_dir, "delta_v_raster_updated.tif"), 
            overwrite = T)
```

```{r}
ranking_raster_se <- raster(file.path(variant_dir, "ranking_raster_se.tif"))

map_priorities(ranking_raster_se, discrete = T, mpas_raster = MPAs, caption = "Carbon, scorched Earth", ocean_df = ocean_matrix) +
  ggsave(file.path(variant_dir, "priority_map_se.png"),
         dpi = 300,  height = 4, width = 8)

ranking_raster <- raster(file.path(variant_dir, "ranking_raster.tif"))

plot(ranking_raster)

map_priorities(ranking_raster, discrete = T, mpas_raster = MPAs, caption = "Carbon, with impacts", ocean_df = ocean_matrix) +
  ggsave(file.path(variant_dir, "priority_map_updated.png"),
         dpi = 300,  height = 4, width = 8)
```

```{r}
ranking_diff <- ranking_raster_se - ranking_raster

ranking_diff %>% 
  as.data.frame(xy = TRUE, centroids = T) %>% 
  set_names(c("lon", "lat", "diff_rank")) %>% 
  filter(!is.na(diff_rank)) %>% 
  ggplot()+
  geom_raster(aes(x = lon, y = lat, fill = diff_rank))+
  geom_raster(data = MPAs_df %>% 
                filter(!is.na(mpa)), aes(x = x, y = y), fill = "lightblue")+
  geom_sf(data = land_50_moll, col = "transparent", fill = "grey", inherit.aes = FALSE)+
  colorspace::scale_fill_continuous_diverging(rev = T,
                                  guide = guide_colorbar(title = "ranking difference",
                                                         title.position = 'bottom',
                                                         direction = "horizontal",
                                                         title.hjust = 0.5,
                                                         label.hjust = -.02,
                                                         keywidth = 0.3))+
  theme(legend.text = element_text(size = 6),
        legend.text.align = 0,
        legend.position = "bottom",
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_line(color = "#ebebe5", size = 0.2),
        panel.grid.minor = element_blank(),
        plot.background = element_rect(fill = NA, color = NA), 
        panel.background = element_rect(fill = NA, color = NA), 
        legend.background = element_rect(fill = NA, color = NA),
        panel.border = element_blank())+
  ggsave(filename = file.path(emLab_project_dir, "figures", "supplement", "carbon_priorities_diff_scorched_earth_updated.png"), 
         dpi = 300,  height = 9, width = 8)
```

# What does it take to get 90% of the benefits?

```{r}
benefit_curve <- read_rds(file.path(variant_dir, "carbon_benefit_updated.rds"))

carbon_benefit_bau <- benefit_curve %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected),
         V_rescaled = scales::rescale(carbon)) %>% 
  slice(which.min(abs(fraction_protected - fraction_in_MPA))) %>% 
  pull(V_rescaled)

ninety_percent_delta_carbon <- carbon_benefit_bau + 0.9*(1 - carbon_benefit_bau)

(f_to_90 <- benefit_curve %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected),
         V_rescaled = scales::rescale(carbon)) %>% 
  slice(which.min(abs(V_rescaled - ninety_percent_delta_carbon))) %>% 
  pull(fraction_protected)) 

# additional protection 1.18%
f_to_90 - fraction_in_MPA
```

#  EEZ vs HS

How much of the area needed to reach 90% beneftis is in EEZ vs HS

```{r}
territories_mask <- raster(file.path(emLab_project_dir, "data", "02_processed", "masks", "territories.tif")) # 1 equals high seas

territories_df <- read_csv(file.path(emLab_project_dir, "data", "02_processed", "masks", "territories_lookup.csv"))

eez_mask <- ocean_low_res_moll %>% 
  mask(territories_mask) %>% 
  mask(territories_mask == 1926, maskvalue = 1) # remove Antartica

hs_mask <- ocean_low_res_moll %>% 
  mask(eez_mask, inverse = T) 

eez_mask_df_moll <- eez_mask %>% 
  as.data.frame(xy = TRUE) %>% 
  set_names(c("x", "y", "eez_mask")) %>% 
  mutate(eez_mask = ifelse(eez_mask == 1, 1, NA)) %>% 
  filter(!is.na(eez_mask)) %>% 
  inner_join(ocean_matrix)
```

```{r}
benefit_curve %>% 
  mutate(in_EEZ = cell_id %in% eez_mask_df_moll$cell_id) %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) %>% 
  filter(fraction_protected <= f_to_90) %>% 
  group_by(in_EEZ) %>% 
  summarise(n_cells = n()) %>% 
  mutate(f = n_cells/sum(n_cells))
```

How much of the top 10% is in EEZ?

```{r}
benefit_curve %>% 
  mutate(in_EEZ = cell_id %in% eez_mask_df_moll$cell_id) %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) %>% 
  filter(fraction_protected <= 0.1) %>% 
  group_by(in_EEZ) %>% 
  summarise(n_cells = n()) %>% 
  ungroup() %>% 
  mutate(f_cell = n_cells/sum(n_cells))
```
