---
title: "Biodiversity with Climate Change in Juanation"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
---

```{r}
library(raster)
library(sf)
library(tidyverse)
library(patchwork)

source(here::here("common.R"))

variant_name <- "10_biodiversity_2050"

variant_dir <- file.path(emLab_project_dir,"data", "03_output", variant_name)

dir.create(variant_dir)
```

# Approach 1: Design for the 2050

This approach plans for the future using the most relevant dataset. Thus, we use aquamaps 2050 distribution + seamounts but exclude provinces and threats as we don't know how those will change in the future. 

## Load data 

### species 2050

```{r, eval = F}
spp_files <- tibble(filepath = list.files(file.path(emLab_project_dir, "data", "02_processed", "species_distributions", "aquamaps_2050"), 
                                            full.names = T),
                   valid_sci_name = str_replace_all(str_remove(basename(filepath), "\\.tif"), "_", " "))

bio_features_df <- stack(spp_files$filepath) %>%
  raster::as.data.frame(xy = T) %>%
  inner_join(ocean_matrix) %>%
  select(x, y, cell_id, everything()) %>%
  as_tibble()

data.table::fwrite(bio_features_df, 
                   file.path(emLab_project_dir, "data", "02_processed", "species_distributions", "aquamaps_spp_distributions_2050_df.csv"))
```

```{r}
spp_2050_df <- data.table::fread(file.path(emLab_project_dir, 
                                           "data", "02_processed", "species_distributions", "aquamaps_spp_distributions_2050_df.csv")) %>% 
  as_tibble()

spp_2050_files <- tibble(filepath = list.files(file.path(emLab_project_dir, "data", "02_processed", "species_distributions", "aquamaps_2050"), 
                                            full.names = T),
                         valid_sci_name = str_replace_all(str_remove(basename(filepath), ".tif"), "_", " "))

spp_info <- data.table::fread(file.path(emLab_project_dir, "data", "02_processed", "species_list", "spp_weights.csv")) %>% 
  as_tibble() %>% 
  rename(w = spp_weight) %>% 
  mutate(new_file = basename(filepath)) %>% 
  filter(new_file %in% basename(spp_2050_files$filepath)) %>% 
  select(-new_file)

smts_df <- data.table::fread(file.path(emLab_project_dir, "data", "02_processed", "features_df", "smts_df.csv")) %>% 
  as_tibble() 

smts_info <- tibble(filepath = list.files(file.path(emLab_project_dir, "data", "02_processed","seamounts"), full.names = T)) %>% 
  mutate(w = sum(spp_info$w)/n())

sum(spp_info$w) ==  sum(smts_info$w)

features_df <- spp_2050_df %>% 
  left_join(smts_df) 

n_features <- nrow(spp_info)  + nrow(smts_info) 

feature_names <- colnames(features_df)[c(-1,-2,-3)]
```

### Normalize and trim

```{r}
features_matrix <- features_df %>% 
  select(-x,-y,-cell_id) %>% 
  as.matrix()

stopifnot(
  sum(map_lgl(features_df, is.numeric)) == ncol(features_df) 
  ) # ALL numeric?

Norm_features_matrix <- sweep(features_matrix, 2, colSums(features_matrix, na.rm = T), FUN = "/")

stopifnot(
  sum(colSums(Norm_features_matrix, na.rm = T)) == ncol(features_matrix)
  )  # Normalized OK?

rownames(Norm_features_matrix) <- features_df$cell_id

Norm_features_matrix <- Norm_features_matrix[rowSums(is.na(Norm_features_matrix)) != ncol(Norm_features_matrix), , drop = F]

stopifnot(
  identical(colnames(Norm_features_matrix), 
          features_df %>% 
            select(-x,-y,-cell_id) %>% 
            colnames())
  )  # Is the order of the features mantained?

Norm_features_matrix[is.na(Norm_features_matrix)] <- 0
```

### Feature weights 

```{r}
feature_wts <- c(spp_info$w, smts_info$w)
  
names(feature_wts) <- feature_names
```

### Current MPAs

```{r}
MPAs <- raster(file.path(emLab_project_dir, "data", "02_processed", "masks", "fully_highly_protected_reviewed_MPAs.tif"))

MPAs_df <- MPAs %>% 
  raster::as.data.frame(xy = T) %>%
  inner_join(ocean_matrix) %>% 
  select(x, y, cell_id, mpa = fully_highly_protected_reviewed_MPAs) %>%
  as_tibble() 

fraction_in_MPA <- sum(!is.na(MPAs_df$mpa))/nrow(MPAs_df)
  
MPA_vect <- rownames(Norm_features_matrix) %>% 
  enframe() %>% 
  select(cell_id = value) %>% 
  left_join(MPAs_df) %>% 
  mutate(mpa =  !is.na(mpa)) %>% pull(mpa)
```

## Initial conditions

```{r}
z <- 0.25

protected_cells <- matrix(MPA_vect, nrow = 1, ncol = nrow(Norm_features_matrix))

colnames(protected_cells) <- rownames(Norm_features_matrix)

unprotected_matrix <- Norm_features_matrix[!protected_cells, ,drop = F]

baseline_state <-  protected_cells%*%Norm_features_matrix

protected_matrix <- Norm_features_matrix[protected_cells, ,drop =F]

max_slopes <- z*matrixStats::colMaxs(Norm_features_matrix, na.rm = T)^(z - 1)

max(max_slopes)
```

## Run Algorithm

```{r}
ranking <- list()

current_pick <- 0

current_state <- baseline_state

start_time <- Sys.time()

while(nrow(unprotected_matrix) > 0){
  
  current_pick <- sum(current_pick, 1)
  
  slopes <- feature_wts*pmin(max_slopes, z*current_state^(z - 1))
  
  delta <- unprotected_matrix%*%as.matrix(slopes)
  
  best_cell_indeces <- doBy::which.maxn(delta, n = 100)
  
  best_cells <- delta[best_cell_indeces, ] %>% 
    enframe() %>% 
    set_names(c('cell_id', "delta")) %>% 
    mutate(pick_order = current_pick)
  
  ranking <- bind_rows(ranking, best_cells)
  
  current_state <- unprotected_matrix[best_cell_indeces, , drop = F] %>% 
    colSums() +
    current_state
  
  unprotected_matrix <- unprotected_matrix[-best_cell_indeces, ,drop = F]
  
}

end_time <- Sys.time()

end_time - start_time

write_rds(ranking, 
          file.path(variant_dir, "ranking_results_approach_1.rds"))
```

## Results

### Benefit curve

```{r}
ranking <- read_rds(file.path(variant_dir, "ranking_results_approach_1.rds"))

representation_curves <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(ranking) %>% 
  mutate(is_mpa = is.na(delta)) %>% 
  full_join(ocean_matrix) %>% 
  left_join(Norm_features_matrix %>% 
              as_tibble(rownames = NA) %>% 
              rownames_to_column("cell_id")) %>% 
  replace(is.na(.), 0) %>% 
  select(-x, -y, -cell_id, -delta, -pick_order, -is_mpa) %>% 
  mutate_all(.funs = cumsum) 

benefit_curve <- representation_curves %>% 
  mutate_all(.funs = ~(.)^0.25) %>% 
  as.matrix() %*% as.matrix(feature_wts) %>% 
  as_tibble() %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected))

benefit_curve %>% 
  filter(fraction_protected > fraction_in_MPA) %>% 
  ggplot()+
  geom_line(aes(x = fraction_protected, y = V/max(V))) +
  geom_vline(xintercept = c(0.1, 0.2, 0.3),
             linetype = c("dashed","dotted","dotdash"))+
  labs(x = "Fraction Ocean Protected",
       y = "Biodiversity Benefit",
       caption = "Biodiversity + Climate Change, approach 1")+
  theme_classic()+
  scale_x_continuous(limits = c(0,1), expand = c(0,0))+
  scale_y_continuous(limits = c(0,1), expand = c(0,0))+
  geom_rect(aes(xmin = 0, 
                xmax = fraction_in_MPA, 
                ymin = 0, 
                ymax = 1),
                fill = "#69B578", alpha = 0.01)+
  ggsave(file.path(variant_dir, "benefit_curve_approach_1.png"),
         width = 7, height = 4)
```

### Maps

```{r}
ranking_raster <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(ranking) %>% 
  mutate(is_mpa = is.na(delta)) %>% 
  left_join(ocean_matrix) %>% 
  tibble::rowid_to_column("ID") %>% 
  mutate(rank = 1- scales::rescale(ID)) %>% 
  select(x, y, rank) %>% 
  rasterFromXYZ(crs = crs(ocean_low_res_moll)) %>% 
  projectRaster(ocean_low_res_moll) %>% 
  mask(ocean_low_res_moll) 

writeRaster(ranking_raster, 
            file.path(variant_dir, "ranking_raster_approach_1.tif"), 
            overwrite = T)
```

```{r}
ranking_raster <- raster( file.path(variant_dir, "ranking_raster_approach_1.tif"))

ranking_map <- tmap::tm_shape(ranking_raster)+
  tmap::tm_raster(title = "",
                  palette  = priorities_map_pal$colors,
                  breaks = priorities_map_pal$breaks,
                  labels = priorities_map_pal$labels,
                  legend.is.portrait = T, 
                  legend.reverse = T)+
  tmap::tm_shape(land_50_moll)+
  tmap::tm_fill(col = "black", border.col = "transparent")+
  tmap::tm_credits("Biodiversity + Climate change, approach 1") +
  tmap::tm_shape(MPAs)+
  tmap::tm_raster(col = "fully_highly_protected_reviewed_MPAs", palette = "#00A600FF", legend.show	= F)+
  tmap::tm_layout(title = "Biodiversity Conservation Priorities",
                  title.position = c("center", .95),
                  inner.margins = c(0.12, 0, 0.08, 0.04), 
                  frame = F,
                  legend.position = c(.99, "center")) 

tmap::tmap_save(ranking_map, 
                filename = file.path(variant_dir, "priority_map_approach_1.png"),
                dpi = 300,  height = 4, width = 8 )
```

#### diff map

```{r}
raster_today <- raster(file.path(emLab_project_dir, "data", "03_output", "05_spp_wts_smts_provs_MPAs","ranking_raster.tif"))

comparison_raster <- overlay(ranking_raster, raster_today, fun = function(x, y){if_else(x >= 0.95 & y >= 0.95, 1, 
                                                                                        if_else(x <= 0.5 & y <= 0.5, -1, 
                                                                                                if_else(abs(x-y) <= 0.01, 0, 
                                                                                                        if_else(x > y, 0.5, -0.5))))})
diff_map <- comparison_raster %>%
  tmap::tm_shape()+
  tmap::tm_raster(title = "",
                  palette = "seq",
                  style = "cat",
                  labels = rev(c("irreplaceable (top 5%)", "increasing", "comparable", "decreasing", "least (bottom 50%)")),
                  legend.is.portrait = T, 
                  legend.reverse = T)+
  tmap::tm_shape(land_50_moll)+
  tmap::tm_fill(col = "gray", border.col = "transparent")+
  tmap::tm_shape(MPAs)+
  tmap::tm_raster(col = "fully_highly_protected_reviewed_MPAs", palette = "#00A600FF", legend.show	= F)+
  tmap::tm_layout(title = "Global Ocean Conservation Priorities",
                  aes.palette = list(seq = "-RdBu"),
                  title.position = c("center", .95),
                  inner.margins = c(0.12, 0, 0.08, 0.04), 
                  frame = F,
                  legend.position = c(.88, .1))

tmap::tmap_save(diff_map, 
               filename = file.path(variant_dir, "map_2050_vs_today_approach_1.png"),
                dpi = 300,  height = 4, width = 8 )

```

# Approach 2: Isolate the effect of Climate change

Here we keep all layer and impacts and only swap the species distributions. 

## Load data 

```{r}
provs_info <- tibble(filepath = list.files(file.path(emLab_project_dir, "data", "02_processed", "biogeography"), full.names = T, pattern = "\\.tif")) %>% 
  mutate(w = sum(spp_info$w)/n())

provs_df <- data.table::fread(file.path(emLab_project_dir, "data", "02_processed", "features_df", "provinces_df.csv"))

sum(spp_info$w) ==  sum(provs_info$w)
```

```{r}
features_df <- spp_2050_df %>% 
  left_join(smts_df) %>% 
  left_join(provs_df)

n_features <- nrow(spp_info)  + nrow(smts_info) + nrow(provs_info) 

feature_names <- colnames(features_df)[c(-1,-2,-3)]
```

### Normalize and trim

```{r}
features_matrix <- features_df %>% 
  select(-x,-y,-cell_id) %>% 
  as.matrix()

stopifnot(
  sum(map_lgl(features_df, is.numeric)) == ncol(features_df) 
  ) # ALL numeric?

Norm_features_matrix <- sweep(features_matrix, 2, colSums(features_matrix, na.rm = T), FUN = "/")

stopifnot(
  sum(colSums(Norm_features_matrix, na.rm = T)) == ncol(features_matrix)
  )  # Normalized OK?

rownames(Norm_features_matrix) <- features_df$cell_id

Norm_features_matrix <- Norm_features_matrix[rowSums(is.na(Norm_features_matrix)) != ncol(Norm_features_matrix), , drop = F]

stopifnot(
  identical(colnames(Norm_features_matrix), 
          features_df %>% 
            select(-x,-y,-cell_id) %>% 
            colnames())
  )  # Is the order of the features mantained?

Norm_features_matrix[is.na(Norm_features_matrix)] <- 0
```

### Feature weights 

```{r}
feature_wts <- c(spp_info$w, smts_info$w, provs_info$w)
  
names(feature_wts) <- feature_names
```

### Current MPAs

```{r}
MPAs <- raster(file.path(emLab_project_dir, "data", "02_processed", "masks", "fully_highly_protected_reviewed_MPAs.tif"))

MPAs_df <- MPAs %>% 
  raster::as.data.frame(xy = T) %>%
  inner_join(ocean_matrix) %>% 
  select(x, y, cell_id, mpa = fully_highly_protected_reviewed_MPAs) %>%
  as_tibble() 

fraction_in_MPA <- sum(!is.na(MPAs_df$mpa))/nrow(MPAs_df)
  
MPA_vect <- rownames(Norm_features_matrix) %>% 
  enframe() %>% 
  select(cell_id = value) %>% 
  left_join(MPAs_df) %>% 
  mutate(mpa =  !is.na(mpa)) %>% pull(mpa)
```

### Apply Impacts 


```{r}
bio_abatable_impacts_df <- raster(file.path(emLab_project_dir, 
                                            "data", "02_processed", "impacts", "chi", "abatable_impacts_5_yr_avg_log.tif")) %>% 
  raster::as.data.frame(xy = T) %>% 
  set_names(c("x", "y", "Ia")) %>% 
  inner_join(ocean_matrix) %>% 
  as_tibble() %>% 
  replace_na(list(Ia = 0))

bio_unabatable_impacts_df <- raster(file.path(emLab_project_dir, 
                                              "data", "02_processed", "impacts", "chi", "unabatable_impacts_5_yr_avg_log.tif")) %>% 
  raster::as.data.frame(xy = T) %>% 
  set_names(c("x", "y", "Iu")) %>% 
  inner_join(ocean_matrix) %>% 
  as_tibble()%>% 
  replace_na(list(Iu = 0))

remains_BAU <- Norm_features_matrix %>% 
  sweep(1, (1 - bio_abatable_impacts_df$Ia), FUN = "*") %>% 
  sweep(1, (1 - bio_unabatable_impacts_df$Iu), FUN = "*") %>% 
  colSums(na.rm = T)

remains_MPA <- Norm_features_matrix %>% 
  sweep(1, (1 - bio_unabatable_impacts_df$Iu), FUN = "*") %>% 
  colSums(na.rm = T)

Norm_features_matrix_diff <- Norm_features_matrix %>% 
  sweep(1, (bio_abatable_impacts_df$Ia - bio_abatable_impacts_df$Ia*bio_unabatable_impacts_df$Iu), FUN = "*")
```

## Initial conditions

```{r}
z <- 0.25

protected_cells <- matrix(MPA_vect, nrow = 1, ncol = nrow(Norm_features_matrix_diff))

colnames(protected_cells) <- rownames(Norm_features_matrix_diff)

unprotected_matrix <- Norm_features_matrix_diff[!protected_cells, ,drop = F]

baseline_state <- remains_BAU + protected_cells%*%Norm_features_matrix_diff

protected_matrix <- Norm_features_matrix_diff[protected_cells, ,drop =F]

max_slopes <- z*baseline_state^(z - 1)
```

## Run algorithm 

```{r}
ranking <- list()

current_pick <- 0

current_state <- baseline_state

start_time <- Sys.time()

while(nrow(unprotected_matrix) > 0){
  
  current_pick <- sum(current_pick, 1)
  
  slopes <- feature_wts*pmin(max_slopes, z*current_state^(z - 1))
  
  delta <- unprotected_matrix%*%t(as.matrix(slopes))
  
  best_cell_indeces <- doBy::which.maxn(delta, n = 100)
  
  best_cells <- delta[best_cell_indeces, ] %>% 
    enframe() %>% 
    set_names(c('cell_id', "delta")) %>% 
    mutate(pick_order = current_pick)
  
  ranking <- bind_rows(ranking, best_cells)
  
  current_state <- unprotected_matrix[best_cell_indeces, , drop = F] %>% 
    colSums() +
    current_state
  
  unprotected_matrix <- unprotected_matrix[-best_cell_indeces, , drop = F]
  
}

end_time <- Sys.time()

end_time - start_time

write_rds(ranking, 
          file.path(variant_dir, "ranking_results_approach_2.rds"))
```

## Results

### Benefit curve

```{r}
ranking_approach_2 <- read_rds(file.path(variant_dir, "ranking_results_approach_2.rds"))

benefit_curve_approach_2 <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(ranking_approach_2) %>% 
  mutate(is_mpa = is.na(delta))  %>% 
  full_join(ocean_matrix) %>% 
  left_join(Norm_features_matrix_diff %>% 
              as_tibble(rownames = NA) %>% 
              rownames_to_column("cell_id")) %>% 
  replace(is.na(.), 0) %>% 
  select(-x, -y, -cell_id, -delta, -pick_order, -is_mpa) %>% 
  mutate_all(.funs = cumsum) %>% 
  sweep(MARGIN = 2, remains_BAU, `+`) %>%
  mutate_all(.funs = ~(.)^0.25) %>% 
  as.matrix() %*% as.matrix(feature_wts) %>% 
  rowSums() %>% 
  as_tibble() %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) 

benefit_curve_approach_2 %>% 
  filter(fraction_protected > fraction_in_MPA) %>% 
  ggplot()+
  geom_line(aes(x = fraction_protected, y = V/sum(feature_wts))) +
  geom_vline(xintercept = c(0.1, 0.2, 0.3),
             linetype = c("dashed","dotted","dotdash"))+
  labs(x = "Fraction Ocean Protected",
       y = "Biodiversity Benefit",
       caption = "Biodiversity + Climate Change, approach 2, Juanation")+
  theme_classic()+
  scale_x_continuous(limits = c(0,1), expand = c(0,0))+
  geom_rect(aes(xmin = 0, 
                xmax = fraction_in_MPA, 
                ymin = min(benefit_curve_approach_2$V)/sum(feature_wts), 
                ymax = max(benefit_curve_approach_2$V)/sum(feature_wts)),
                fill = "#69B578", alpha = 0.01)+
  ggsave(file.path(variant_dir, "benefit_curve_approach_2.png"),
         width = 7, height = 4)
```

### Maps

```{r}
ranking_raster_approach_2 <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(ranking_approach_2) %>% 
  mutate(is_mpa = is.na(delta)) %>% 
  left_join(ocean_matrix) %>% 
  tibble::rowid_to_column("ID") %>% 
  mutate(rank = 1- scales::rescale(ID)) %>% 
  select(x, y, rank) %>% 
  rasterFromXYZ(crs = crs(ocean_low_res_moll)) %>% 
  projectRaster(ocean_low_res_moll) %>% 
  mask(ocean_low_res_moll) 

writeRaster(ranking_raster_approach_2, 
            file.path(variant_dir, "ranking_raster_approach_2.tif"), 
            overwrite = T)
```

```{r}
ranking_map_approach_2 <- tmap::tm_shape(ranking_raster_approach_2)+
  tmap::tm_raster(title = "",
                  palette  = priorities_map_pal$colors,
                  breaks = priorities_map_pal$breaks,
                  labels = priorities_map_pal$labels,
                  legend.is.portrait = T, 
                  legend.reverse = T)+
  tmap::tm_shape(land_50_moll)+
  tmap::tm_fill(col = "black", border.col = "transparent")+
  tmap::tm_credits("Biodiversity + Climate change, approach 2") +
  tmap::tm_shape(MPAs)+
  tmap::tm_raster(col = "fully_highly_protected_reviewed_MPAs", palette = "#00A600FF", legend.show	= F)+
  tmap::tm_layout(title = "Biodiversity Conservation Priorities",
                  title.position = c("center", .95),
                  inner.margins = c(0.12, 0, 0.08, 0.04), 
                  frame = F,
                  legend.position = c(.99, "center")) 

tmap::tmap_save(ranking_map_approach_2, 
                filename = file.path(variant_dir, "priority_map_approach_2.png"),
                dpi = 300,  height = 4, width = 8 )
```

```{r}
comparison_raster_approach_2 <- overlay(ranking_raster_approach_2, 
                                        raster_today,
                                        fun = function(x, y){if_else(x >= 0.95 & y >= 0.95, 1, 
                                                                     if_else(x <= 0.5 & y <= 0.5, -1, 
                                                                             if_else(abs(x-y) <= 0.01, 0, 
                                                                                     if_else(x > y, 0.5, -0.5))))})

diff_map_approach_2 <- comparison_raster_approach_2 %>%
  tmap::tm_shape()+
  tmap::tm_raster(title = "",
                  palette = "seq",
                  style = "cat",
                  labels = rev(c("irreplaceable (top 5%)", "increasing", "comparable", "decreasing", "least (bottom 50%)")),
                  legend.is.portrait = T, 
                  legend.reverse = T)+
  tmap::tm_shape(land_50_moll)+
  tmap::tm_fill(col = "gray", border.col = "transparent")+
  tmap::tm_shape(MPAs)+
  tmap::tm_raster(col = "fully_highly_protected_reviewed_MPAs", palette = "#00A600FF", legend.show	= F)+
  tmap::tm_layout(title = "Global Ocean Conservation Priorities",
                  aes.palette = list(seq = "-RdBu"),
                  title.position = c("center", .95),
                  inner.margins = c(0.12, 0, 0.08, 0.04), 
                  frame = F,
                  legend.position = c(.88, .1))

tmap::tmap_save(diff_map_approach_2, 
               filename = file.path(variant_dir, "map_2050_vs_today_approach_2.png"),
                dpi = 300,  height = 4, width = 8 )
```
