---
title: "Multi-objective prioritization in Juanation"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
---

```{r}
library(raster)
library(sf)
library(tidyverse)
library(patchwork)
library(parallel)
library(foreach)
library(doParallel)

source(here::here("common.R"))

variant_name <- "11_multiple_objectives"

variant_dir <- file.path(emLab_project_dir,"data", "03_output", variant_name)

dir.create(variant_dir)
```

# Load features data 

## Biodiversity 

```{r}
bio_df <- data.table::fread(file.path(emLab_project_dir, "data", "02_processed", "features_df", "all_bio_features_df.csv"))

bio_feature_names <- colnames(bio_df)[c(-1,-2,-3)]

n_bio_features <- length(bio_feature_names)

spp_wts <- data.table::fread(file.path(emLab_project_dir, "data", "02_processed", "species_list", "spp_weights.csv")) %>% 
  as_tibble() %>% 
  rename(w = spp_weight)

smts_wts <- tibble(filepath = list.files(file.path(emLab_project_dir, "data", "02_processed","seamounts"), full.names = T)) %>% 
  mutate(w = sum(spp_wts$w)/n())

provs_wts <- tibble(filepath = list.files(file.path(emLab_project_dir, "data", "02_processed", "biogeography"), full.names = T, pattern = "\\.tif")) %>% 
  mutate(w = sum(spp_wts$w)/n())

bio_features_info <- bind_rows(spp_wts %>% 
                                 mutate(sub_goal = "species"), 
                               smts_wts %>% 
                                 mutate(sub_goal = "seamounts"), 
                               provs_wts %>% 
                                 mutate(sub_goal = "provinces"))

z_bio <- 0.25
```

## Food

```{r}
stocks_df <- data.table::fread(file.path(emLab_project_dir, "data", "02_processed", "features_df", "stocks_df.csv")) %>% 
  as_tibble()

stocks <- tibble(filename = list.files(file.path(emLab_project_dir, "data", "02_processed", "stock_distributions"),
                                       full.names = T),
                 stockid = str_remove(basename(filename), "\\.tif"))

stocks_info <- read_csv(file.path(emLab_project_dir,"data", "02_processed", "food_provision", "MegaData_UncertaintyAnalysis.csv")) %>%
  select(stockid, k = Kfin, ex_rate = ExploitationRate_BAU1, m = m_fin, r = r_fin)

stocks_info <- stocks %>%
  left_join(stocks_info)

stocks_info$ex_rate[stocks_info$ex_rate >= 1] <- 0.99999

n_stocks <- nrow(stocks_info)

stock_names <- colnames(stocks_df)[c(-1,-2,-3)]
```

## Carbon 

```{r}
carbon_df <- data.table::fread(file.path(emLab_project_dir, "data", "02_processed", "features_df", "carbon_df.csv")) %>% 
  as_tibble()

n_carbon <- 1

z_carbon <- 1
```

```{r}
features_df <- bio_df %>% 
  left_join(carbon_df) %>% 
  left_join(stocks_df)

n_features <- n_stocks + n_carbon + n_bio_features
```

# Within-goal weights

```{r}
bio_weights <- bio_features_info$w

names(bio_weights) <- bio_feature_names

carbon_weights <- c("carbon" = 1)

food_weights <- rep(1, n_stocks)

names(food_weights) <- stock_names
```

# Normalize and trim the data

```{r}
features_matrix <- features_df %>% 
  select(-x,-y,-cell_id) %>% 
  as.matrix()

stopifnot(
  sum(map_lgl(features_df %>%
                select(-x,-y,-cell_id), is.numeric)) == ncol(features_matrix)
  ) # ALL numeric?

Norm_features_matrix <- sweep(features_matrix, 2, colSums(features_matrix, na.rm = T), FUN = "/")

rownames(Norm_features_matrix) <- features_df$cell_id

stopifnot(
  sum(colSums(Norm_features_matrix, na.rm = T)) == ncol(features_matrix)
  )  # Normalized OK?

dim(Norm_features_matrix)

Norm_features_matrix <- Norm_features_matrix[rowSums(is.na(Norm_features_matrix)) != n_features, ]

stopifnot(
  identical(colnames(Norm_features_matrix),
          features_df %>%
            select(-x,-y,-cell_id) %>%
            colnames())
  )  # Is the order of the features mantained?
```

# Current MPAs

```{r}
MPAs <- raster(file.path(emLab_project_dir, "data", "02_processed", "masks", "fully_highly_protected_reviewed_MPAs.tif"))

MPAs_df <- MPAs %>% 
  raster::as.data.frame(xy = T) %>%
  inner_join(ocean_matrix) %>% 
  select(x, y, cell_id, mpa = fully_highly_protected_reviewed_MPAs) %>%
  as_tibble() 

fraction_in_MPA <- sum(!is.na(MPAs_df$mpa))/nrow(MPAs_df)
  
MPA_vect <- rownames(Norm_features_matrix) %>% 
  enframe() %>% 
  select(cell_id = value) %>% 
  left_join(MPAs_df) %>% 
  mutate(mpa =  !is.na(mpa)) %>% pull(mpa)
```

# Apply impacts

## Biodiversity

```{r}
bio_abatable_impacts_df <- raster(file.path(emLab_project_dir, 
                                            "data", "02_processed", "impacts", "chi", "abatable_impacts_5_yr_avg_log.tif")) %>% 
  raster::as.data.frame(xy = T) %>% 
  set_names(c("x", "y", "Ia")) %>% 
  inner_join(ocean_matrix) %>% 
  as_tibble() %>% 
  replace_na(list(Ia = 0))

bio_unabatable_impacts_df <- raster(file.path(emLab_project_dir, 
                                              "data", "02_processed", "impacts", "chi", "unabatable_impacts_5_yr_avg_log.tif")) %>% 
  raster::as.data.frame(xy = T) %>% 
  set_names(c("x", "y", "Iu")) %>% 
  inner_join(ocean_matrix) %>% 
  as_tibble()%>% 
  replace_na(list(Iu = 0))

bio_remains_BAU <- Norm_features_matrix[, bio_feature_names] %>% 
  sweep(1, (1 - bio_abatable_impacts_df$Ia), FUN = "*") %>% 
  sweep(1, (1 - bio_unabatable_impacts_df$Iu), FUN = "*") %>% 
  colSums(na.rm = T)

sum(bio_remains_BAU^z_bio*bio_weights)/sum(1^z_bio*bio_weights)

bio_remains_MPA <- Norm_features_matrix[, bio_feature_names] %>% 
  sweep(1, (1 - bio_unabatable_impacts_df$Iu), FUN = "*") %>% 
  colSums(na.rm = T)

#(sum(bio_remains_MPA^0.25*bio_weights) - sum(bio_remains_BAU^0.25*bio_weights))/ sum(bio_remains_BAU^0.25*bio_weights)

bio_diff_matrix <- Norm_features_matrix[, bio_feature_names] %>% 
  sweep(1, (bio_abatable_impacts_df$Ia - bio_abatable_impacts_df$Ia*bio_unabatable_impacts_df$Iu), FUN = "*")

bio_diff_made <- colSums(bio_diff_matrix, na.rm = T)

Norm_features_matrix[, bio_feature_names] <- bio_diff_matrix
```

## Carbon

```{r}
carbon_abatable_impacts_df <- raster(file.path(emLab_project_dir, 
                                            "data", "02_processed", "impacts", "bottom_trawling", "abatable_impact_carbon_svr.tif")) %>% 
  raster::as.data.frame(xy = T) %>% 
  set_names(c("x", "y", "Ia")) %>% 
  inner_join(ocean_matrix) %>% 
  as_tibble() %>% 
  replace_na(list(Ia = 0)) %>% 
  filter(cell_id %in% rownames(Norm_features_matrix))

carbon_abatable_impacts_df <- carbon_abatable_impacts_df %>% 
  mutate(Ia = if_else(Ia == 0, 10^-6, Ia))

carbon_remains_BAU <- Norm_features_matrix[, "carbon", drop = F] %>% 
  sweep(1, (1 - carbon_abatable_impacts_df$Ia), FUN = "*") %>% 
  colSums(na.rm = T)

carbon_remains_MPA <- Norm_features_matrix[, "carbon", drop = F] %>% 
  sweep(1, (1 - 0), FUN = "*") %>% 
  colSums(na.rm = T)

carbon_diff_matrix <- Norm_features_matrix[, "carbon", drop = F] %>% 
  sweep(1, (carbon_abatable_impacts_df$Ia), FUN = "*")

carbon_diff_made <- colSums(carbon_diff_matrix, na.rm = T)

Norm_features_matrix[, "carbon"] <- carbon_diff_matrix

Norm_features_matrix[is.na(Norm_features_matrix)] <- 0

# write_rds(Norm_features_matrix, 
#           file.path(variant_dir, "multi_objective_norm_features_diff_matrix.rds"))
```

# Feature weights

```{r}
max_food_delta_a1 <- read_rds(file.path(emLab_project_dir, "data", "03_output", "07_food", "results_per_pixel.rds")) %>% 
  summarize(max_deltaH = max(delta_catch_a1),
            max_deltaH_bau =  max(delta_catch_a1[is_mpa])) %>% 
  transmute(max_deltaH - max_deltaH_bau) %>% 
  pull()

max_food_delta_a2 <- read_rds(file.path(emLab_project_dir, "data", "03_output", "07_food", "results_per_pixel.rds")) %>% 
  summarize(max_deltaH = max(delta_catch_a2),
            max_deltaH_bau =  max(delta_catch_a2[is_mpa])) %>% 
  transmute(max_deltaH - max_deltaH_bau) %>% 
  pull()

max_bio_delta <- sum(bio_remains_MPA^z_bio*bio_weights) - sum(bio_remains_BAU^z_bio*bio_weights)

max_carbon_delta <- sum(carbon_remains_MPA*carbon_weights) - sum(carbon_remains_BAU*carbon_weights)

calculate_weights <- function(w_bio, w_carbon, w_food, food_model){
  
  if(w_bio + w_carbon + w_food != 1) stop("weights don't add up to 1")
  
  if(food_model == "a1"){
    
    final_bio_w <- w_bio*bio_weights*max_food_delta_a1/max_bio_delta
    
    final_carbon_w <- w_carbon*carbon_weights*max_food_delta_a1/max_carbon_delta
    
    final_food_w <- w_food*food_weights
    
  }
  else if (food_model == "a2"){
    
    final_bio_w <- w_bio*bio_weights*max_food_delta_a2/max_bio_delta
    
    final_carbon_w <- w_carbon*carbon_weights*max_food_delta_a2/max_carbon_delta
    
    final_food_w <- w_food*food_weights
    
  }
  
  feature_weigths <- c(final_bio_w, final_carbon_w, final_food_w)
  
  return(feature_weigths)
  
}
```

# Food benefit functions

```{r, food_slopes_a1}
estimate_food_slopes_a1 <- function(e = stocks_info$ex_rate, 
                               m = stocks_info$m,
                               r = stocks_info$r, 
                               k = stocks_info$k,
                               R){
  
  R[R > 1] <- 1 
  
  slope <- -e*k*m*(m+e)*((e*r + 2*e*m)*R + m*r - 2*e*m)/(r*(e*R + m)^3)
  
  slope[R == 1] <- -9*10^9
  
  return(slope)
  
}

estimate_delta_catch_a1 <- function(x, 
                                    e = stocks_info$ex_rate,
                                    m = stocks_info$m, 
                                    r = stocks_info$r, 
                                    k = stocks_info$k){
  
  R_bau <- 0
  
  x[x > 1] <- 1 
  
  exp_rate <- e
  
  catch_mpa <- exp_rate*(m*k)*(1 - x)/(exp_rate*x + m)*(1 - exp_rate*(1 - x)*m/((exp_rate*x + m)*r))

  catch_bau <- e*(m*k*(1 - R_bau)/(e*R_bau + m)*(1 - e*(1 - R_bau)*m/((e*R_bau + m)*r)))
  
  delta_catch <- catch_mpa - catch_bau
  
  return(delta_catch)
  
}

estimate_food_slopes_a2 <- function(e = stocks_info$ex_rate, 
                                    m = stocks_info$m,
                                    r = stocks_info$r, 
                                    k = stocks_info$k, 
                                    R){
  
  R[R > 1] <- 1 
  
  A <- (1 - e)^(1/(1 - R))
  
  part_a <- A*log(1 - e)*k*m*(1 - (1 - A)*m*(1 - R)/(r*((1 - A)*R + m)))/((1 - R)*((1 - A)*R + m))
  
  part_b <- (1 - A)*k*m*(1 - (1 - A)*m*(1 - R)/(r*((1 - A)*R + m)))/((1 - A)*R + m)
  
  part_c <- (1 - A)*k*m*(1 - R)*(- A*log(1 - e)*R/(1 - R)^2 - A + 1)*(1 - (1 - A)*m*(1 - R)/(r*((1 - A)*R + m)))/((1 - A)*R + m)^2
  
  part_d <- (1 - A)*k*m*(1 - R)*(A*log(1 - e)*m/(r*(1 - R)*((1 - A)*R + m)) + (1 - A)*m/(r*((1 - A)*R + m)) + (1 - A)*m*(1 - R)*(-A*log(1 - e)*R/(1 - R)^2 - A + 1)/(r*((1 - A)*R + m)^2))/((1 - A)*R + m)
  
  slope <- - part_a - part_b - part_c + part_d
  
  slope[R == 1] <- -9*10^9
  
  return(slope)
  
}

estimate_delta_catch_a2 <- function(x, 
                                    e = stocks_info$ex_rate, 
                                    m = stocks_info$m,
                                    r = stocks_info$r, 
                                    k = stocks_info$k) {
  
  R_bau <- 0
  
  x[x > 1] <- 1 
  
  exp_rate <- 1 - (1 - e)^(1/(1 - x))
  
  catch_mpa <- exp_rate*(m*k)*(1 - x)/(exp_rate*x + m)*(1 - exp_rate*(1 - x)*m/((exp_rate*x + m)*r))
  
  catch_bau <- e*(m*k*(1 - R_bau)/(e*R_bau + m)*(1 - e*(1 - R_bau)*m/((e*R_bau + m)*r)))
  
  delta_catch <- catch_mpa - catch_bau
  
  return(delta_catch)
  
}
```

# Get slopes foo

```{r}
estimate_current_slopes <- function(x, g, wts, food_model){
  
  slopes <- vector(mode = "numeric", length = n_features)
  
  slopes[g == 1] <- pmin(max_slopes_bio, z_bio*x[g == 1]^(z_bio - 1))*wts[g == 1]
  
  slopes[g == 2] <- pmin(max_slope_carbon, z_carbon*x[g == 2]^(z_carbon - 1))*wts[g == 2]
  
  if(food_model == "a1") {
    
    slopes[g == 3] <- estimate_food_slopes_a1(R = x[g == 3])*wts[g == 3]}
  
  else if (food_model == "a2") {
    
    slopes[g == 3] <- estimate_food_slopes_a2(R = x[g == 3])*wts[g == 3]
  }
  
  return(slopes)
  
}

# where g is a vector that identifies each feature as part of each goal: bio (1), carbon (2), food (3)
group_ids <- c(rep(1, n_bio_features), rep(2, n_carbon), rep(3, n_stocks))
```

# Ranking 

## Food Assumption A1

```{r}
multiobjective_feature_weigths <- calculate_weights(w_bio = 1/3, w_carbon = 1/3, w_food = 1/3, food_model = "a1")

# sum(bio_remains_MPA^z_bio*multiobjective_feature_weigths[bio_feature_names]) - sum(bio_remains_BAU^z_bio*multiobjective_feature_weigths[bio_feature_names])
# multiobjective_feature_weigths["carbon"]*max_carbon_delta
# max_food_delta_a1/3

protected_cells <- matrix(MPA_vect, nrow = 1, ncol = nrow(Norm_features_matrix))

unprotected_matrix <- Norm_features_matrix[!protected_cells, ]

protected_matrix <- Norm_features_matrix[protected_cells, ]

baseline_state <- protected_cells%*%Norm_features_matrix

baseline_state[,"carbon"] <- baseline_state[, "carbon"] + carbon_remains_BAU

baseline_state[, bio_feature_names] <- baseline_state[, bio_feature_names] + bio_remains_BAU

max_slopes_bio <- z_bio*baseline_state[, bio_feature_names]^(z_bio - 1)

max_slope_carbon <- 1

names(max_slope_carbon) <- "carbon"
```

```{r}
ranking <- list()

current_pick <- 0

current_state <- baseline_state 

start_time <- Sys.time()

while(nrow(unprotected_matrix) > 0){
    
    current_pick <- sum(current_pick, 1)
    
    slopes <- estimate_current_slopes(x = current_state, g = group_ids, wts = multiobjective_feature_weigths, food_model = "a1")
  
    delta <- unprotected_matrix%*%as.matrix(slopes)
    
    best_cell_indeces <- doBy::which.maxn(delta, n = 100)
    
    best_cells <- delta[best_cell_indeces,] %>% 
      enframe() %>% 
      set_names(c('cell_id', "delta")) %>% 
      mutate(pick_order = current_pick)
    
    ranking <- bind_rows(ranking, best_cells)
    
    current_state <- unprotected_matrix[best_cell_indeces, ] %>% 
      colSums() +
      current_state
    
    unprotected_matrix <- unprotected_matrix[-best_cell_indeces, ]
    
  }

end_time <- Sys.time()

end_time - start_time

write_rds(ranking, 
          file.path(variant_dir, "ranking_results_a1.csv"))
```

## Food Assumption A2

```{r}
multiobjective_feature_weigths <- calculate_weights(w_bio = 1/3, w_carbon = 1/3, w_food = 1/3, food_model = "a2")

protected_cells <- matrix(MPA_vect, nrow = 1, ncol = nrow(Norm_features_matrix))

unprotected_matrix <- Norm_features_matrix[!protected_cells, ]

protected_matrix <- Norm_features_matrix[protected_cells, ]

baseline_state <- protected_cells%*%Norm_features_matrix

baseline_state[,"carbon"] <- baseline_state[, "carbon"] + carbon_remains_BAU

baseline_state[, bio_feature_names] <- baseline_state[, bio_feature_names] + bio_remains_BAU

max_slopes_bio <- z_bio*baseline_state[, bio_feature_names]^(z_bio - 1)

max_slope_carbon <- 1

names(max_slope_carbon) <- "carbon"
```

```{r}
ranking <- list()

current_pick <- 0

current_state <- baseline_state 

start_time <- Sys.time()

while(nrow(unprotected_matrix) > 0){
    
    current_pick <- sum(current_pick, 1)
    
    slopes <- estimate_current_slopes(x = current_state, g = group_ids, wts = multiobjective_feature_weigths, food_model = "a2")
  
    delta <- unprotected_matrix%*%as.matrix(slopes)
    
    best_cell_indeces <- doBy::which.maxn(delta, n = 100)
    
    best_cells <- delta[best_cell_indeces,] %>% 
      enframe() %>% 
      set_names(c('cell_id', "delta")) %>% 
      mutate(pick_order = current_pick)
    
    ranking <- bind_rows(ranking, best_cells)
    
    current_state <- unprotected_matrix[best_cell_indeces, ] %>% 
      colSums() +
      current_state
    
    unprotected_matrix <- unprotected_matrix[-best_cell_indeces, ]
    
  }

end_time <- Sys.time()

end_time - start_time

write_rds(ranking, 
          file.path(variant_dir, "ranking_results_a2.csv"))
```

# Results

## Food assumption A1

```{r}
ranking_a1 <- read_rds(file.path(variant_dir, "ranking_results_a1.csv"))

results_a1 <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(ranking_a1) %>% 
  mutate(is_mpa = is.na(delta))  %>% 
  full_join(ocean_matrix) %>% 
  left_join(Norm_features_matrix %>% 
              as_tibble(rownames = NA) %>% 
              rownames_to_column("cell_id")) %>% 
  replace(is.na(.), 0) %>% 
  select(-x, -y, -delta, -pick_order, -is_mpa) %>% 
  mutate(cell_id = as.numeric(cell_id))
```

### benefit curves

```{r}
carbon_benefit_curve_a1 <- results_a1 %>% 
  select(V = carbon) %>% 
  mutate(V = cumsum(V)*multiobjective_feature_weigths["carbon"],
         V_norm = V/max(V),
         fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) %>% 
  filter(fraction_protected > fraction_in_MPA)

bio_benefit_curve_a1 <- results_a1 %>% 
  select(all_of(bio_feature_names)) %>% 
  mutate_all(.funs = cumsum) %>% 
  sweep(MARGIN = 2, bio_remains_BAU, `+`) %>%
  mutate_all(.funs = ~(.)^z_bio) %>% 
  as.matrix() %*% as.matrix(multiobjective_feature_weigths[bio_feature_names]) %>% 
  as_tibble() %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) %>% 
  filter(fraction_protected > fraction_in_MPA) %>% 
  mutate(V = V - sum(bio_remains_BAU^z_bio*multiobjective_feature_weigths[bio_feature_names]),
         V_norm = V/max(V))

food_benefit_curve_a1 <- results_a1 %>% 
  select(all_of(stock_names)) %>% 
  mutate_all(.funs = cumsum) %>% 
  apply(MARGIN = 1, estimate_delta_catch_a1) %>% 
  t() %>% 
  as.matrix() %*% as.matrix(multiobjective_feature_weigths[stock_names]) %>% 
  as_tibble() %>% 
  rowSums() %>% 
  enframe(name = NULL) %>% 
  set_names("V") %>% 
  mutate(V_norm = V/max(V),
         fraction_protected = 1/n(),
         fraction_protected =  cumsum(fraction_protected)) %>% 
  filter(fraction_protected > fraction_in_MPA)

max(food_benefit_curve_a1$V)
max(bio_benefit_curve_a1$V)
max(carbon_benefit_curve_a1$V)

bio_benefit_curve_a1 %>% 
  select(fraction_protected, biodiversity = V_norm) %>% 
  left_join(carbon_benefit_curve_a1 %>% 
              select(fraction_protected, carbon = V_norm)) %>% 
  left_join(food_benefit_curve_a1 %>% 
              select(fraction_protected, food = V_norm)) %>% 
  gather(var, value, -fraction_protected) %>% 
  ggplot()+
  geom_line(aes(fraction_protected, value, col = var))+ 
  labs(x = "Fraction of the ocean protected", 
       y = "Benefit",
       col = "Objective")+
  theme_classic()+
  ggsave(file.path(variant_dir, "benefit_curves_equal_weights_a1.png"),
         width = 7, height = 4)
```

### priority map

```{r}
ranking_raster_a1 <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(ranking_a1) %>% 
  mutate(is_mpa = is.na(delta)) %>% 
  left_join(ocean_matrix) %>% 
  tibble::rowid_to_column("ID") %>% 
  mutate(rank = 1- scales::rescale(ID)) %>% 
  select(x, y, rank) %>% 
  rasterFromXYZ(crs = crs(ocean_low_res_moll)) %>% 
  projectRaster(ocean_low_res_moll) %>% 
  mask(ocean_low_res_moll) 

writeRaster(ranking_raster_a1, 
            file.path(variant_dir, "ranking_raster_equal_weights_a1.tif"), 
            overwrite = T)

ranking_map_a1 <- tmap::tm_shape(ranking_raster_a1)+
  tmap::tm_raster(title = "",
                  palette  = priorities_map_pal$colors,
                  breaks = priorities_map_pal$breaks,
                  labels = priorities_map_pal$labels,
                  legend.is.portrait = T, 
                  legend.reverse = T)+
  tmap::tm_shape(land_50_moll)+
  tmap::tm_fill(col = "black", border.col = "transparent")+
  tmap::tm_credits("Multi-objective, equal weights, no effort displacement") +
  tmap::tm_shape(MPAs)+
  tmap::tm_raster(col = "fully_highly_protected_reviewed_MPAs", palette = "#00A600FF", legend.show	= F)+
  tmap::tm_layout(title = "Global Conservation Priorities",
                  title.position = c("center", .95),
                  inner.margins = c(0.12, 0, 0.08, 0.04), 
                  frame = F,
                  legend.position = c(.99, "center")) 

tmap::tmap_save(ranking_map_a1, 
                filename = file.path(variant_dir, "priority_map_equal_weights_a1.png"),
                dpi = 300,  height = 4, width = 8 )
```

## Food assumption A2

```{r}
ranking_a2 <- read_rds(file.path(variant_dir, "ranking_results_a2.csv"))

results_a2 <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(ranking_a2) %>% 
  mutate(is_mpa = is.na(delta))  %>% 
  full_join(ocean_matrix) %>% 
  left_join(Norm_features_matrix %>% 
              as_tibble(rownames = NA) %>% 
              rownames_to_column("cell_id")) %>% 
  replace(is.na(.), 0) %>% 
  select(-x, -y, -delta, -pick_order, -is_mpa) %>% 
  mutate(cell_id = as.numeric(cell_id))
```

### benefit curves

```{r}
carbon_benefit_curve_a2 <- results_a2 %>% 
  select(V = carbon) %>% 
  mutate(V = cumsum(V)*multiobjective_feature_weigths["carbon"],
         V_norm = V/max(V),
         fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) %>% 
  filter(fraction_protected > fraction_in_MPA)

bio_benefit_curve_a2 <- results_a2 %>% 
  select(all_of(bio_feature_names)) %>% 
  mutate_all(.funs = cumsum) %>% 
  sweep(MARGIN = 2, bio_remains_BAU, `+`) %>%
  mutate_all(.funs = ~(.)^z_bio) %>% 
  as.matrix() %*% as.matrix(multiobjective_feature_weigths[bio_feature_names]) %>% 
  as_tibble() %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) %>% 
  filter(fraction_protected > fraction_in_MPA) %>% 
  mutate(V = V - sum(bio_remains_BAU^z_bio*multiobjective_feature_weigths[bio_feature_names]),
         V_norm = V/max(V))

food_benefit_curve_a2 <- results_a2 %>% 
  select(all_of(stock_names)) %>% 
  mutate_all(.funs = cumsum) %>% 
  apply(MARGIN = 1, estimate_delta_catch_a2) %>% 
  t() %>% 
  as.matrix() %*% as.matrix(multiobjective_feature_weigths[stock_names]) %>% 
  as_tibble() %>% 
  rowSums() %>% 
  enframe(name = NULL) %>% 
  set_names("V") %>% 
  mutate(V_norm = V/max(V),
         fraction_protected = 1/n(),
         fraction_protected =  cumsum(fraction_protected)) %>% 
  filter(fraction_protected > fraction_in_MPA)

max(food_benefit_curve_a2$V)/10^6
max(bio_benefit_curve_a2$V)/10^6
max(carbon_benefit_curve_a2$V)/10^6

bio_benefit_curve_a2 %>% 
  select(fraction_protected, biodiversity = V_norm) %>% 
  left_join(carbon_benefit_curve_a2 %>% 
              select(fraction_protected, carbon = V_norm)) %>% 
  left_join(food_benefit_curve_a2 %>% 
              select(fraction_protected, food = V_norm)) %>% 
  gather(var, value, -fraction_protected) %>% 
  ggplot()+
  geom_line(aes(fraction_protected, value, col = var))+ 
  labs(x = "Fraction of the ocean protected", 
       y = "Benefit",
       col = "Objective")+
  theme_classic()+
  ggsave(file.path(variant_dir, "benefit_curves_equal_weights_a2.png"),
         width = 7, height = 4)
```

### priority map

```{r}
ranking_raster_a2 <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(ranking_a2) %>% 
  mutate(is_mpa = is.na(delta)) %>% 
  left_join(ocean_matrix) %>% 
  tibble::rowid_to_column("ID") %>% 
  mutate(rank = 1- scales::rescale(ID)) %>% 
  select(x, y, rank) %>% 
  rasterFromXYZ(crs = crs(ocean_low_res_moll)) %>% 
  projectRaster(ocean_low_res_moll) %>% 
  mask(ocean_low_res_moll) 

writeRaster(ranking_raster_a2, 
            file.path(variant_dir, "ranking_raster_equal_weights_a2.tif"), 
            overwrite = T)

ranking_map_a2 <- tmap::tm_shape(ranking_raster_a2)+
  tmap::tm_raster(title = "",
                  palette  = priorities_map_pal$colors,
                  breaks = priorities_map_pal$breaks,
                  labels = priorities_map_pal$labels,
                  legend.is.portrait = T, 
                  legend.reverse = T)+
  tmap::tm_shape(land_50_moll)+
  tmap::tm_fill(col = "black", border.col = "transparent")+
  tmap::tm_credits("Multi-objective, equal weights, Full effort displacement") +
  tmap::tm_shape(MPAs)+
  tmap::tm_raster(col = "fully_highly_protected_reviewed_MPAs", palette = "#00A600FF", legend.show	= F)+
  tmap::tm_layout(title = "Global Conservation Priorities",
                  title.position = c("center", .95),
                  inner.margins = c(0.12, 0, 0.08, 0.04), 
                  frame = F,
                  legend.position = c(.99, "center")) 

tmap::tmap_save(ranking_map_a2, 
                filename = file.path(variant_dir, "priority_map_equal_weights_a2.png"),
                dpi = 300,  height = 4, width = 8 )
```

# Tradeoffs

Here we ask, if we prioritize one goal only, how much do we get of the others as co-benefit? For this we use the main fisheries model A2

```{r}
multi_objective_matrix <- read_rds(file.path(variant_dir, "multi_objective_norm_features_diff_matrix.rds"))

bio_ranking <- read_rds(file.path(emLab_project_dir, "data", "03_output", "05_spp_wts_smts_provs_MPAs", "ranking_results.rds"))

carbon_ranking <- read_rds(file.path(emLab_project_dir, "data", "03_output", "06_carbon", "ranking_results.rds"))

food_ranking <- read_rds(file.path(emLab_project_dir, "data", "03_output", "07_food", "ranking_results_a2.rds"))

max_food <- read_rds(file.path(emLab_project_dir, "data", "03_output", "07_food", "results_per_pixel.rds")) %>% 
  summarize(max_deltaH = max(delta_catch_a2),
            max_deltaH_bau =  max(delta_catch_a2[is_mpa])) %>% 
  pull(max_deltaH)
```

## Biodiversity priority

```{r}
bio_priority_results <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(bio_ranking) %>% 
  mutate(is_mpa = is.na(delta))  %>% 
  full_join(ocean_matrix) %>% 
  left_join(multi_objective_matrix %>% 
              as_tibble(rownames = NA) %>% 
              rownames_to_column("cell_id")) %>% 
  replace(is.na(.), 0) %>% 
  select(-x, -y, -delta, -pick_order, -is_mpa) %>% 
  mutate(cell_id = as.numeric(cell_id)) 

bio_benefit_curve_bio_priority <- bio_priority_results %>% 
  select(bio_feature_names) %>% 
  mutate_all(.funs = cumsum) %>% 
  sweep(MARGIN = 2, bio_remains_BAU, `+`) %>%
  mutate_all(.funs = ~(.)^z_bio) %>% 
  as.matrix() %*% as.matrix(bio_weights) %>% 
  as_tibble() %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) 

carbon_benefit_curve_bio_priority <- bio_priority_results %>% 
  select(carbon) %>% 
  mutate_all(.funs = cumsum) %>% 
  sweep(MARGIN = 2, carbon_remains_BAU, `+`) %>%
  mutate_all(.funs = ~(.)) %>% 
  rowSums() %>% 
  enframe(name = NULL) %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) 

food_benefit_curve_bio_priority <- bio_priority_results %>% 
  select(stock_names) %>% 
  mutate_all(.funs = cumsum) %>% 
  apply(MARGIN = 1, estimate_delta_catch_a2) %>% 
  t() %>% 
  as_tibble() %>% 
  rowSums() %>% 
  enframe(name = NULL) %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected =  cumsum(fraction_protected)) 

co_benefits_biodiversity <- food_benefit_curve_bio_priority %>% 
  rename(food = value) %>% 
  left_join(bio_benefit_curve_bio_priority %>% 
              rename(biodiversity = V) %>% 
              mutate(biodiversity = scales::rescale(biodiversity))) %>% 
  left_join(carbon_benefit_curve_bio_priority %>% 
              rename(carbon = V) %>% 
              mutate(carbon = scales::rescale(carbon))) %>% 
  select(fraction_protected, everything())

write_csv(co_benefits_biodiversity, 
          file.path(variant_dir, "co_benefits_biodiversity_a2.csv"))

multi_objective_curves_biodiversity_priority <- co_benefits_biodiversity %>% 
  mutate(food = food/max_food,
         biodiversity = scales::rescale(biodiversity),
         carbon = scales::rescale(carbon)) %>% 
  gather(var, value, -fraction_protected) %>% 
  ggplot()+
  geom_line(aes(fraction_protected, value, col = var))+
  theme_classic()+
  labs(y = "Benefit", x = "Fraction of the Ocean protected", col = "Objective")
```

```{r}
co_benefits_biodiversity <- read_csv(file.path(output_dir, "co_benefits_biodiversity.csv"))
```

How much catch would we lose with the top 20% of biodiversity? 

```{r}
co_benefits_biodiversity %>% 
  slice(which.min(abs(fraction_protected - 0.2075)))
```

## Carbon priority

```{r}
carbon_priority_results <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(carbon_ranking) %>% 
  mutate(is_mpa = is.na(delta))  %>% 
  full_join(ocean_matrix) %>% 
  left_join(multi_objective_matrix %>% 
              as_tibble(rownames = NA) %>% 
              rownames_to_column("cell_id")) %>% 
  replace(is.na(.), 0) %>% 
  select(-x, -y, -delta, -pick_order, -is_mpa) %>% 
  mutate(cell_id = as.numeric(cell_id)) 

bio_benefit_curve_carbon_priority <- carbon_priority_results %>% 
  select(bio_feature_names) %>% 
  mutate_all(.funs = cumsum) %>% 
  sweep(MARGIN = 2, bio_remains_BAU, `+`) %>%
  mutate_all(.funs = ~(.)^0.25) %>% 
  as.matrix() %*% as.matrix(bio_weights) %>% 
  as_tibble() %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) 

carbon_benefit_curve_carbon_priority <- carbon_priority_results %>% 
  select(carbon) %>% 
  mutate_all(.funs = cumsum) %>% 
  sweep(MARGIN = 2, carbon_remains_BAU, `+`) %>%
  mutate_all(.funs = ~(.)) %>% 
  rowSums() %>% 
  enframe(name = NULL) %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) 

food_benefit_curve_carbon_priority <- carbon_priority_results %>% 
  select(stock_names) %>% 
  mutate_all(.funs = cumsum) %>% 
  apply(MARGIN = 1, estimate_delta_catch_a2) %>% 
  t() %>% 
  as_tibble() %>% 
  rowSums() %>% 
  enframe(name = NULL) %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected =  cumsum(fraction_protected)) 

co_benefits_carbon <- food_benefit_curve_carbon_priority %>% 
  rename(food = value) %>% 
  left_join(bio_benefit_curve_carbon_priority %>% 
              rename(biodiversity = V) %>% 
              mutate(biodiversity = scales::rescale(biodiversity))) %>% 
  left_join(carbon_benefit_curve_carbon_priority %>% 
              rename(carbon = V) %>% 
              mutate(carbon = scales::rescale(carbon))) %>%
  select(fraction_protected, everything())

write_csv(co_benefits_carbon, 
          file.path(variant_dir, "co_benefits_carbon_a2.csv"))

multi_objective_curves_carbon_priority <- co_benefits_carbon %>% 
  mutate(food = food/max_food,
         biodiversity = scales::rescale(biodiversity),
         carbon = scales::rescale(carbon)) %>% 
  gather(var, value, -fraction_protected) %>% 
  ggplot()+
  geom_line(aes(fraction_protected, value, col = var))+
  theme_classic()+
  labs(y = "Benefit", x = "Fraction of the Ocean protected", col = "Objective")
```

## Food priority

```{r}
food_priority_results <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(food_ranking) %>% 
  mutate(is_mpa = is.na(delta))  %>% 
  full_join(ocean_matrix) %>% 
  left_join(multi_objective_matrix %>% 
              as_tibble(rownames = NA) %>% 
              rownames_to_column("cell_id")) %>% 
  replace(is.na(.), 0) %>% 
  select(-x, -y, -delta, -pick_order, -is_mpa) %>% 
  mutate(cell_id = as.numeric(cell_id)) 

bio_benefit_curve_food_priority <- food_priority_results %>% 
  select(bio_feature_names) %>% 
  mutate_all(.funs = cumsum) %>% 
  sweep(MARGIN = 2, bio_remains_BAU, `+`) %>%
  mutate_all(.funs = ~(.)^0.25) %>% 
  as.matrix() %*% as.matrix(bio_weights) %>% 
  as_tibble() %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) 

carbon_benefit_curve_food_priority <- food_priority_results %>% 
  select(carbon) %>% 
  mutate_all(.funs = cumsum) %>% 
  sweep(MARGIN = 2, carbon_remains_BAU, `+`) %>%
  mutate_all(.funs = ~(.)) %>% 
  rowSums() %>% 
  enframe(name = NULL) %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) 

food_benefit_curve_food_priority <- food_priority_results %>% 
  select(stock_names) %>% 
  mutate_all(.funs = cumsum) %>% 
  apply(MARGIN = 1, estimate_delta_catch_a2) %>% 
  t() %>% 
  as_tibble() %>% 
  rowSums() %>% 
  enframe(name = NULL) %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected =  cumsum(fraction_protected)) 

max(food_benefit_curve_food_priority$value)

co_benefits_food <- food_benefit_curve_food_priority %>% 
  rename(food = value) %>% 
  left_join(bio_benefit_curve_food_priority %>% 
              rename(biodiversity = V) %>% 
              mutate(biodiversity = scales::rescale(biodiversity))) %>% 
  left_join(carbon_benefit_curve_food_priority %>% 
              rename(carbon = V) %>% 
              mutate(carbon = scales::rescale(carbon))) 

write_csv(co_benefits_food, 
          file.path(variant_dir, "co_benefits_food_a2.csv"))

multi_objective_curves_food_priority <- co_benefits_food %>% 
  mutate(food = food/max_food,
         biodiversity = scales::rescale(biodiversity),
         carbon = scales::rescale(carbon)) %>% 
  gather(var, value, -fraction_protected) %>% 
  ggplot()+
  geom_line(aes(fraction_protected, value, col = var))+
  theme_classic()+
  labs(y = "Benefit", x = "Fraction of the Ocean protected", col = "Objective")
```







