---
title: "Multi-objective prioritization in Juanation"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
---

```{r}
library(raster)
library(sf)
library(tidyverse)

DIR_GCS <- "~/gcs/git-annex/Carta-Marina"

DIR_GCS_data <- "~/gcs/spatial-datasets"

ocean_low_res_moll <- raster("~/gcs/spatial-datasets/ocean/ocean-low-res-moll.tif")

ocean_matrix <- ocean_low_res_moll %>% 
  raster::as.data.frame(xy = T) %>% 
  filter(ocean.low.res.moll == 1) %>% 
  mutate(cell_id = as.character(group_indices(., x, y))) %>% 
  select(-ocean.low.res.moll) %>% 
  as_tibble()

land_shp <- sf::read_sf("~/gcs/spatial-datasets/land/land_50.shp")

land_shp_moll <- land_shp %>% 
  st_transform(crs = projection(ocean_low_res_moll))

z_pal <- list(breaks = c(0, 0.2, 0.5, 0.7, 0.8, 0.85, 0.9, 0.95, 1),
              labels = c("0-20", "20-50", "50-70", "70-80", "80-85", "85-90", "90-95", "95-100"),
              colors = c("#4575b4", "#74add1", "#abd9e9", "#e0f3f8" ,"#fee090" ,"#fdae61" ,"#f46d43", "#d73027"))

variant_name <- "11_multiple_objectives"

output_dir <- file.path(variant_name, "results")

figs_output_dir <- file.path(variant_name, "figures")

dir.create(variant_name)

dir.create(figs_output_dir)

dir.create(output_dir)
```

# Load features data 

## Biodiversity 

```{r}
spp_df <- data.table::fread("data/spp_df.csv") %>% 
  as_tibble()

spp_info <- data.table::fread("data/spp_weights.csv") %>% 
  as_tibble() %>% 
  rename(w = spp_weight)

smts_df <- data.table::fread("data/smts_df.csv") %>% 
  as_tibble()

smts_info <- tibble(filepath = list.files("../Zonation/data/seamounts/all", full.names = T)) %>% 
  mutate(w = sum(spp_info$w)/n())

sum(spp_info$w) ==  sum(smts_info$w)

provs_df <- data.table::fread("data/provs_df.csv") %>% 
  as_tibble()

provs_info <- tibble(filepath = list.files(c("../Zonation/data/biogeography/Spalding",
                                             "../Zonation/data/biogeography/GOODS"), 
                                           full.names = T)) %>% 
  mutate(w = sum(spp_info$w)/n())

sum(spp_info$w) ==  sum(provs_info$w)

bio_df <- spp_df %>% 
  left_join(smts_df) %>% 
  left_join(provs_df)

n_bio_features <- nrow(spp_info)  + nrow(smts_info) + nrow(provs_info) 

bio_feature_names <- colnames(bio_df)[c(-1,-2,-3)]

bio_features_info <- bind_rows(spp_info %>% 
                                 mutate(sub_goal = "species"), 
                               smts_info %>% 
                                 mutate(sub_goal = "seamounts"), 
                               provs_info %>% 
                                 mutate(sub_goal = "provinces"))
```

## Food

```{r}
stocks_info <- read_csv("data/stocks_info_bau_1.csv")

stocks_df <- data.table::fread("data/stocks_df_bau_1.csv") %>% 
  as_tibble()

n_stocks <- nrow(stocks_info)

stock_names <- colnames(stocks_df)[c(-1,-2,-3)]
```

## Carbon 

```{r}
carbon_df <- data.table::fread("data/carbon_df.csv") %>% 
  as_tibble()

n_carbon <- 1
```

```{r}
features_df <- bio_df %>% 
  left_join(carbon_df) %>% 
  left_join(stocks_df)

n_features <- n_stocks + n_carbon + n_bio_features
```

# Within-goal weights

```{r}
bio_weights <- bio_features_info$w

names(bio_weights) <- bio_feature_names

carbon_weights <- c("carbon" = 1)

food_weights <- rep(1, n_stocks)

names(food_weights) <- stock_names
```

# Normalize and trim the data

```{r}
features_matrix <- features_df %>% 
  select(-x,-y,-cell_id) %>% 
  as.matrix()

# stopifnot(
#   sum(map_lgl(features_df %>%
#                 select(-x,-y,-cell_id), is.numeric)) == ncol(features_matrix)
#   ) # ALL numeric?

Norm_features_matrix <- sweep(features_matrix, 2, colSums(features_matrix, na.rm = T), FUN = "/")

rownames(Norm_features_matrix) <- features_df$cell_id

# stopifnot(
#   sum(colSums(Norm_features_matrix, na.rm = T)) == ncol(features_matrix)
#   )  # Normalized OK?

dim(Norm_features_matrix)

Norm_features_matrix <- Norm_features_matrix[rowSums(is.na(Norm_features_matrix)) != n_features, ]

# stopifnot(
#   identical(colnames(Norm_features_matrix), 
#           features_df %>% 
#             select(-x,-y,-cell_id) %>% 
#             colnames())
#   )  # Is the order of the features mantained?
```

# Current MPAs

```{r}
MPAs <- raster("../Zonation/data/masks/fully_highly_protected_reviewed_MPAs.tif")

MPAs_df <- MPAs %>% 
  raster::as.data.frame(xy = T) %>%
  inner_join(ocean_matrix) %>% 
  select(x, y, cell_id, mpa = fully_highly_protected_reviewed_MPAs) %>%
  as_tibble() 

fraction_in_MPA <- sum(!is.na(MPAs_df$mpa))/nrow(MPAs_df)
  
MPA_vect <- rownames(Norm_features_matrix) %>% 
  enframe() %>% 
  select(cell_id = value) %>% 
  left_join(MPAs_df) %>% 
  mutate(mpa =  !is.na(mpa)) %>% pull(mpa)
```

# Apply retention

## Biodiversity

```{r}
bio_abatable_impacts_df <- raster("../../../gcs/git-annex/Carta-Marina/Impacts/data/processed/Abatable_impacts_5yr_avg_log.tif") %>% 
  raster::as.data.frame(xy = T) %>% 
  set_names(c("x", "y", "Ia")) %>% 
  inner_join(ocean_matrix) %>% 
  as_tibble() %>% 
  replace_na(list(Ia = 0))

bio_unabatable_impacts_df <- raster("../../../gcs/git-annex/Carta-Marina/Impacts/data/processed/Unbatable_impacts_5yr_avg_log.tif") %>% 
  raster::as.data.frame(xy = T) %>% 
  set_names(c("x", "y", "Iu")) %>% 
  inner_join(ocean_matrix) %>% 
  as_tibble()%>% 
  replace_na(list(Iu = 0))

Norm_features_matrix_BAU_bio <- Norm_features_matrix[, bio_feature_names] %>% 
  sweep(1, (1 - bio_abatable_impacts_df$Ia), FUN = "*") %>% 
  sweep(1, (1 - bio_unabatable_impacts_df$Iu), FUN = "*")

bio_remains_BAU <- colSums(Norm_features_matrix_BAU_bio, na.rm = T)

sum(bio_remains_BAU^0.25*bio_weights)/sum(1^0.25*bio_weights)

Norm_features_matrix_MPAs_bio <- Norm_features_matrix[, bio_feature_names] %>% 
  sweep(1, (1 - bio_unabatable_impacts_df$Iu), FUN = "*")

bio_remains_MPA <- colSums(Norm_features_matrix_MPAs_bio, na.rm = T)

#(sum(bio_remains_MPA^0.25*bio_weights) - sum(bio_remains_BAU^0.25*bio_weights))/ sum(bio_remains_BAU^0.25*bio_weights)

bio_diff_matrix <- Norm_features_matrix[, bio_feature_names] %>% 
  sweep(1, (bio_abatable_impacts_df$Ia - bio_abatable_impacts_df$Ia*bio_unabatable_impacts_df$Iu), FUN = "*")

bio_diff_made <- colSums(bio_diff_matrix, na.rm = T)

Norm_features_matrix[, bio_feature_names] <- bio_diff_matrix
```

## Carbon

```{r}
carbon_abatable_impacts_df <- raster("../../../gcs/git-annex/Carta-Marina/Impacts/data/processed/Abatable_impacts_carbon_GFW_SAR.tif") %>% 
  raster::as.data.frame(xy = T) %>% 
  set_names(c("x", "y", "Ia")) %>% 
  inner_join(ocean_matrix) %>% 
  as_tibble() %>% 
  replace_na(list(Ia = 0)) %>% 
  filter(cell_id %in% rownames(Norm_features_matrix))

carbon_abatable_impacts_df <- carbon_abatable_impacts_df %>% 
  mutate(Ia = if_else(Ia == 0, 10^-6, Ia))

Norm_features_matrix_BAU_carbon <- Norm_features_matrix[, "carbon", drop = F] %>% 
  sweep(1, (1 - carbon_abatable_impacts_df$Ia), FUN = "*")

carbon_remains_BAU <- colSums(Norm_features_matrix_BAU_carbon, na.rm = T)

1- carbon_remains_BAU

Norm_features_matrix_MPAs_carbon <- Norm_features_matrix[, "carbon", drop = F] %>% 
  sweep(1, (1 - 0), FUN = "*")

carbon_remains_MPA <- colSums(Norm_features_matrix_MPAs_carbon, na.rm = T)

carbon_diff_matrix <- Norm_features_matrix[, "carbon", drop = F] %>% 
  sweep(1, (carbon_abatable_impacts_df$Ia), FUN = "*")

carbon_diff_made <- colSums(carbon_diff_matrix, na.rm = T)

Norm_features_matrix[, "carbon"] <- carbon_diff_matrix

Norm_features_matrix[is.na(Norm_features_matrix)] <- 0

write_rds(Norm_features_matrix, 
          file.path(output_dir, "multi_objective_norm_features_diff_matrix.rds"))
```

# Initial conditions

```{r}
protected_cells <- matrix(MPA_vect, nrow = 1, ncol = nrow(Norm_features_matrix))

OG_unprotected_matrix <- Norm_features_matrix[!protected_cells, ]

OG_protected_matrix <- Norm_features_matrix[protected_cells, ]
```

```{r}
baseline_state <- protected_cells%*%Norm_features_matrix

baseline_state[,"carbon"] <- baseline_state[, "carbon"] + carbon_remains_BAU

baseline_state[, bio_feature_names] <- baseline_state[, bio_feature_names] + bio_remains_BAU
```

```{r}
max_slopes_bio <- 0.25*baseline_state[, bio_feature_names]^(0.25 - 1)

max_slopes_bio[which(!is.finite(max_slopes_bio))] <- 0

max_slope_carbon <- 1

names(max_slope_carbon) <- "carbon"

food_foo <- function(e, m, r, k, R){
  
  -e*k*m*(m+e)*((e*r + 2*e*m)*R + m*r - 2*e*m)/(r*(e*R + m)^3)
  
}

stocks_info$max_frac_k <- matrixStats::colMaxs(OG_unprotected_matrix[, stock_names], na.rm = T)

stocks_info <- stocks_info %>% 
  mutate(max_slope = food_foo(ex_rate, m, r, k, R = max_frac_k))

max_slopes_food <- pull(stocks_info, max_slope)

names(max_slopes_food) <- stock_names

max_slopes <- c(max_slopes_bio, max_slope_carbon, max_slopes_food)
```

```{r}
estimate_current_slope <- function(e, m, r, k, R, p, z, wts){
  
  wts*(p*food_foo(e, m, r, k, R) + (1 - p)*pmin(max_slopes, z*R^(z - 1)))
  
}
```

```{r}
param_e <- c(rep(1, n_bio_features + n_carbon), stocks_info$ex_rate)
param_m <- c(rep(1, n_bio_features + n_carbon), stocks_info$m)
param_r <- c(rep(1, n_bio_features + n_carbon), stocks_info$r)
param_k <- c(rep(1, n_bio_features + n_carbon), stocks_info$k)
z_vector <- c(rep(0.25, n_bio_features), rep(1, n_carbon), rep(1, n_stocks))
p_vector <- c(rep(0, n_bio_features), rep(0, n_carbon), rep(1, n_stocks))
```

# Run ranking:

## equal weights

```{r}
max_food_delta <- 10.7*10^6

max_bio_delta <- sum(bio_remains_MPA^0.25*bio_weights) - sum(bio_remains_BAU^0.25*bio_weights)

max_carbon_delta <- sum(carbon_remains_MPA*carbon_weights) - sum(carbon_remains_BAU*carbon_weights)

feature_weigths <- c(bio_weights*max_food_delta/max_bio_delta, 
                     carbon_weights*max_food_delta/max_carbon_delta, 
                     food_weights)

sum(feature_weigths[bio_feature_names])/10^6

sum(feature_weigths["carbon"])/10^6

sum(feature_weigths[stock_names])/10^6
```

## run algorithm

```{r}
ranking <- list()

current_pick <- 0

current_state <- baseline_state 

unprotected_matrix <- OG_unprotected_matrix

start_time <- Sys.time()

while(nrow(unprotected_matrix) > 0){
    
    current_pick <- sum(current_pick, 1)
    
    slopes <- estimate_current_slope(param_e, param_m, param_r, param_k, current_state, p_vector, z_vector, wts = feature_weigths)
    
    delta <- unprotected_matrix%*%as.matrix(t(slopes))
    
    best_cell_indeces <- doBy::which.maxn(delta, n = 100)
    
    best_cells <- delta[best_cell_indeces,] %>% 
      enframe() %>% 
      set_names(c('cell_id', "delta")) %>% 
      mutate(pick_order = current_pick)
    
    ranking <- bind_rows(ranking, best_cells)
    
    current_state <- unprotected_matrix[best_cell_indeces, ] %>% 
      colSums() +
      current_state
    
    unprotected_matrix <- unprotected_matrix[-best_cell_indeces, ]
    
  }

end_time <- Sys.time()

end_time - start_time

write_rds(ranking, 
          file.path(output_dir, "ranking_results.csv"))
```

# Results

```{r}
results <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(ranking) %>% 
  mutate(is_mpa = is.na(delta))  %>% 
  full_join(ocean_matrix) %>% 
  left_join(Norm_features_matrix %>% 
              as_tibble(rownames = NA) %>% 
              rownames_to_column("cell_id")) %>% 
  replace(is.na(.), 0) %>% 
  select(-x, -y, -delta, -pick_order, -is_mpa) %>% 
  mutate(cell_id = as.numeric(cell_id))
```

```{r}
carbon_benefit_curve <- results %>% 
  select(V = carbon) %>% 
  mutate(V = cumsum(V)*feature_weigths["carbon"],
         fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) %>% 
  filter(fraction_protected > fraction_in_MPA)

param_e <- stocks_info$ex_rate
param_m <- stocks_info$m
param_r <- stocks_info$r
param_k <- stocks_info$k

food_benefit_curve <- results %>% 
  select(stock_names) %>% 
  mutate_all(.funs = cumsum) %>% 
  apply(MARGIN = 1, 
        function(x) param_e*(param_m*param_k*(1 - x)/(param_e*x+param_m))*(1 - (param_e*(1-x)*param_m)/((param_e*x + param_m)*param_r)) - param_e*((param_r - param_e)/param_r)*param_k) %>% 
  t() %>% 
  as_tibble() %>% 
  rowSums() %>% 
  enframe(name = NULL) %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected =  cumsum(fraction_protected)) %>% 
  filter(fraction_protected > fraction_in_MPA)

bio_benefit_curve <- results %>% 
  select(bio_feature_names) %>% 
  mutate_all(.funs = cumsum) %>% 
  sweep(MARGIN = 2, bio_remains_BAU, `+`) %>%
  mutate_all(.funs = ~(.)^0.25) %>% 
  as.matrix() %*% as.matrix(feature_weigths[bio_feature_names]) %>% 
  as_tibble() %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) %>% 
  filter(fraction_protected > fraction_in_MPA) %>% 
  mutate(V = V - sum(bio_remains_BAU^0.25*feature_weigths[bio_feature_names]))

bio_benefit_curve %>% 
  rename(biodiversity = V) %>% 
left_join(carbon_benefit_curve %>% 
            rename(carbon = V)) %>% 
  left_join(food_benefit_curve %>% 
            rename(food = value)) %>% 
  gather(var, value, -fraction_protected) %>% 
  ggplot()+
  geom_line(aes(fraction_protected, value, col = var))+ 
  labs(x = "Fraction of the ocean protected", 
       y = "Benefit",
       col = "Objective")+
  theme_classic()+
  ggsave(file.path(figs_output_dir, "benefit_curves_equal_weights.png"),
         width = 7, height = 4)
```

```{r}
ranking_raster <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(ranking) %>% 
  mutate(is_mpa = is.na(delta)) %>% 
  left_join(ocean_matrix) %>% 
  tibble::rowid_to_column("ID") %>% 
  mutate(rank = 1- scales::rescale(ID)) %>% 
  select(x, y, rank) %>% 
  rasterFromXYZ(crs = crs(ocean_low_res_moll)) %>% 
  projectRaster(ocean_low_res_moll) %>% 
  mask(ocean_low_res_moll) 

writeRaster(ranking_raster, 
            file.path(output_dir, "ranking_raster_equal_weights.tif"), 
            overwrite = T)

ranking_map <- tmap::tm_shape(ranking_raster)+
  tmap::tm_raster(title = "",
                  palette  = z_pal$colors,
                  breaks = z_pal$breaks,
                  labels = z_pal$labels,
                  legend.is.portrait = T, 
                  legend.reverse = T)+
  tmap::tm_shape(land_shp_moll)+
  tmap::tm_fill(col = "black", border.col = "transparent")+
  tmap::tm_credits("Multi-objective, equal weights") +
  tmap::tm_shape(MPAs)+
  tmap::tm_raster(col = "fully_highly_protected_reviewed_MPAs", palette = "#00A600FF", legend.show	= F)+
  tmap::tm_layout(title = "Global Conservation Priorities",
                  title.position = c("center", .95),
                  inner.margins = c(0.12, 0, 0.08, 0.04), 
                  frame = F,
                  legend.position = c(.99, "center")) 

tmap::tmap_save(ranking_map, 
                filename = file.path(figs_output_dir, "priority_map_equal_weights.png"),
                dpi = 300,  height = 4, width = 8 )
```

# Tradeoffs

Here we ask, if we prioritize one goal only, how much do we get of the others as co-benefit?

```{r}
multi_objective_matrix <- read_rds(file.path(output_dir, "multi_objective_norm_features_diff_matrix.rds"))

bio_ranking <- read_rds("05_spp_wts_smts_provs_MPAs/results/ranking_results_with_impacts.rds")
carbon_ranking <- read_rds("06_carbon/results/ranking_results_with_impacts.rds")
food_ranking <- read_rds("07_food/results/ranking_results_bau_1.rds")

max_food <- 10.7*10^6
```

## Biodiversity priority

```{r}
bio_priority_results <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(bio_ranking) %>% 
  mutate(is_mpa = is.na(delta))  %>% 
  full_join(ocean_matrix) %>% 
  left_join(multi_objective_matrix %>% 
              as_tibble(rownames = NA) %>% 
              rownames_to_column("cell_id")) %>% 
  replace(is.na(.), 0) %>% 
  select(-x, -y, -delta, -pick_order, -is_mpa) %>% 
  mutate(cell_id = as.numeric(cell_id)) 

bio_benefit_curve_bio_priority <- bio_priority_results %>% 
  select(bio_feature_names) %>% 
  mutate_all(.funs = cumsum) %>% 
  sweep(MARGIN = 2, bio_remains_BAU, `+`) %>%
  mutate_all(.funs = ~(.)^0.25) %>% 
  as.matrix() %*% as.matrix(bio_weights) %>% 
  as_tibble() %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) 

carbon_benefit_curve_bio_priority <- bio_priority_results %>% 
  select(carbon) %>% 
  mutate_all(.funs = cumsum) %>% 
  sweep(MARGIN = 2, carbon_remains_BAU, `+`) %>%
  mutate_all(.funs = ~(.)) %>% 
  rowSums() %>% 
  enframe(name = NULL) %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) 

param_e <- stocks_info$ex_rate
param_m <- stocks_info$m
param_r <- stocks_info$r
param_k <- stocks_info$k

food_benefit_curve_bio_priority <- bio_priority_results %>% 
  select(stock_names) %>% 
  mutate_all(.funs = cumsum) %>% 
  apply(MARGIN = 1, 
        function(x) param_e*(param_m*param_k*(1 - x)/(param_e*x+param_m))*(1 - (param_e*(1-x)*param_m)/((param_e*x + param_m)*param_r)) - param_e*((param_r - param_e)/param_r)*param_k) %>% 
  t() %>% 
  as_tibble() %>% 
  rowSums() %>% 
  enframe(name = NULL) %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected =  cumsum(fraction_protected)) 

co_benefits_biodiversity <- food_benefit_curve_bio_priority %>% 
  rename(food = value) %>% 
  left_join(bio_benefit_curve_bio_priority %>% 
              rename(biodiversity = V) %>% 
              mutate(biodiversity = scales::rescale(biodiversity))) %>% 
  left_join(carbon_benefit_curve_bio_priority %>% 
              rename(carbon = V) %>% 
              mutate(carbon = scales::rescale(carbon))) %>% 
  select(fraction_protected, everything())

write_csv(co_benefits_biodiversity, 
          file.path(output_dir, "co_benefits_biodiversity.csv"))

multi_objective_curves_biodiversity_priority <- co_benefits_biodiversity %>% 
  mutate(food = food/max_food,
         biodiversity = scales::rescale(biodiversity),
         carbon = scales::rescale(carbon)) %>% 
  gather(var, value, -fraction_protected) %>% 
  ggplot()+
  geom_line(aes(fraction_protected, value, col = var))+
  theme_classic()+
  labs(y = "Benefit", x = "Fraction of the Ocean protected", col = "Objective")
```

```{r}
co_benefits_biodiversity <- read_csv(file.path(output_dir, "co_benefits_biodiversity.csv"))

multi_objective_curves_biodiversity_priority <- co_benefits_biodiversity %>% 
  mutate(food = food/max_food,
         biodiversity = scales::rescale(biodiversity),
         carbon = scales::rescale(carbon)) %>% 
  gather(var, value, -fraction_protected) %>% 
  ggplot()+
  geom_line(aes(fraction_protected, value, col = var))+
  theme_classic()+
  labs(y = "Benefit", x = "Fraction of the Ocean protected", col = "Objective")
```

How much catch would we lose with the top 20% of biodiversity? 

```{r}
co_benefits_biodiversity %>% 
  slice(which.min(abs(fraction_protected - 0.2075)))
```

## Carbon priority

```{r}
carbon_priority_results <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(carbon_ranking) %>% 
  mutate(is_mpa = is.na(delta))  %>% 
  full_join(ocean_matrix) %>% 
  left_join(multi_objective_matrix %>% 
              as_tibble(rownames = NA) %>% 
              rownames_to_column("cell_id")) %>% 
  replace(is.na(.), 0) %>% 
  select(-x, -y, -delta, -pick_order, -is_mpa) %>% 
  mutate(cell_id = as.numeric(cell_id)) 

bio_benefit_curve_carbon_priority <- carbon_priority_results %>% 
  select(bio_feature_names) %>% 
  mutate_all(.funs = cumsum) %>% 
  sweep(MARGIN = 2, bio_remains_BAU, `+`) %>%
  mutate_all(.funs = ~(.)^0.25) %>% 
  as.matrix() %*% as.matrix(bio_weights) %>% 
  as_tibble() %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) 

carbon_benefit_curve_carbon_priority <- carbon_priority_results %>% 
  select(carbon) %>% 
  mutate_all(.funs = cumsum) %>% 
  sweep(MARGIN = 2, carbon_remains_BAU, `+`) %>%
  mutate_all(.funs = ~(.)) %>% 
  rowSums() %>% 
  enframe(name = NULL) %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) 


food_benefit_curve_carbon_priority <- carbon_priority_results %>% 
  select(stock_names) %>% 
  mutate_all(.funs = cumsum) %>% 
  apply(MARGIN = 1, 
        function(x) param_e*(param_m*param_k*(1 - x)/(param_e*x+param_m))*(1 - (param_e*(1-x)*param_m)/((param_e*x + param_m)*param_r)) - param_e*((param_r - param_e)/param_r)*param_k) %>% 
  t() %>% 
  as_tibble() %>% 
  rowSums() %>% 
  enframe(name = NULL) %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected =  cumsum(fraction_protected)) 

co_benefits_carbon <- food_benefit_curve_carbon_priority %>% 
  rename(food = value) %>% 
  left_join(bio_benefit_curve_carbon_priority %>% 
              rename(biodiversity = V) %>% 
              mutate(biodiversity = scales::rescale(biodiversity))) %>% 
  left_join(carbon_benefit_curve_carbon_priority %>% 
              rename(carbon = V) %>% 
              mutate(carbon = scales::rescale(carbon))) %>%
  select(fraction_protected, everything())

write_csv(co_benefits_carbon, 
          file.path(output_dir, "co_benefits_carbon.csv"))

multi_objective_curves_carbon_priority <- co_benefits_carbon %>% 
  mutate(food = food/max_food,
         biodiversity = scales::rescale(biodiversity),
         carbon = scales::rescale(carbon)) %>% 
  gather(var, value, -fraction_protected) %>% 
  ggplot()+
  geom_line(aes(fraction_protected, value, col = var))+
  theme_classic()+
  labs(y = "Benefit", x = "Fraction of the Ocean protected", col = "Objective")
```

## Food priority

```{r}
food_priority_results <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(food_ranking) %>% 
  mutate(is_mpa = is.na(delta))  %>% 
  full_join(ocean_matrix) %>% 
  left_join(multi_objective_matrix %>% 
              as_tibble(rownames = NA) %>% 
              rownames_to_column("cell_id")) %>% 
  replace(is.na(.), 0) %>% 
  select(-x, -y, -delta, -pick_order, -is_mpa) %>% 
  mutate(cell_id = as.numeric(cell_id)) 

bio_benefit_curve_food_priority <- food_priority_results %>% 
  select(bio_feature_names) %>% 
  mutate_all(.funs = cumsum) %>% 
  sweep(MARGIN = 2, bio_remains_BAU, `+`) %>%
  mutate_all(.funs = ~(.)^0.25) %>% 
  as.matrix() %*% as.matrix(bio_weights) %>% 
  as_tibble() %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) 

carbon_benefit_curve_food_priority <- food_priority_results %>% 
  select(carbon) %>% 
  mutate_all(.funs = cumsum) %>% 
  sweep(MARGIN = 2, carbon_remains_BAU, `+`) %>%
  mutate_all(.funs = ~(.)) %>% 
  rowSums() %>% 
  enframe(name = NULL) %>% 
  set_names("V") %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected = cumsum(fraction_protected)) 

food_benefit_curve_food_priority <- food_priority_results %>% 
  select(stock_names) %>% 
  mutate_all(.funs = cumsum) %>% 
  apply(MARGIN = 1, 
        function(x) param_e*(param_m*param_k*(1 - x)/(param_e*x+param_m))*(1 - (param_e*(1-x)*param_m)/((param_e*x + param_m)*param_r)) - param_e*((param_r - param_e)/param_r)*param_k) %>% 
  t() %>% 
  as_tibble() %>% 
  rowSums() %>% 
  enframe(name = NULL) %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected =  cumsum(fraction_protected)) 

co_benefits_food <- food_benefit_curve_food_priority %>% 
  rename(food = value) %>% 
  left_join(bio_benefit_curve_food_priority %>% 
              rename(biodiversity = V) %>% 
              mutate(biodiversity = scales::rescale(biodiversity))) %>% 
  left_join(carbon_benefit_curve_food_priority %>% 
              rename(carbon = V) %>% 
              mutate(carbon = scales::rescale(carbon))) 

write_csv(co_benefits_food, 
          file.path(output_dir, "co_benefits_food.csv"))

multi_objective_curves_food_priority <- co_benefits_food %>% 
  mutate(food = food/max_food,
         biodiversity = scales::rescale(biodiversity),
         carbon = scales::rescale(carbon)) %>% 
  gather(var, value, -fraction_protected) %>% 
  ggplot()+
  geom_line(aes(fraction_protected, value, col = var))+
  theme_classic()+
  labs(y = "Benefit", x = "Fraction of the Ocean protected", col = "Objective")
```







