---
title: "13_efficiency_curve"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
---

```{r}
library(raster)
library(sf)
library(tidyverse)
library(patchwork)
library(parallel)
library(foreach)
library(doParallel)

source(here::here("common.R"))

variant_name <- "13_efficiency_curve"

variant_dir <- file.path(emLab_project_dir, "data", "03_output", variant_name)

dir.create(variant_dir)

options(scipen = 999)
```

# Define preferences

This represent how important is biodiversity relative to fisheries. A value of 2 means biodiversity is 2x more important than food. A value of 0.1 means food is 10 times as valuable as biodiversity. A negative value of -1 means that biodiversity is a valuable as provision but is negatively weighted. 

```{r}
x_bios <- c(seq(-100, -10, by = 10),
            seq(-9, -1, by = 1),
            seq(-0.9, 0, by = .1),
            seq(0.01, 0.1, by = .01),
            seq(0.11, 1, by = 0.1),
            seq(2, 10, by = 1), seq(20, 100, by = 10), seq(200, 1000, by = 100), 10000)
```

# Load features data 

```{r}
# source(here::here("functions", "food_provision_foos.R"))
# 
# load(file.path(emLab_project_dir, "data", "03_output", "11_multiple_objectives", "ranking_inputs.RData"))
# 
# save(bio_feature_names, bio_weights, carbon_weights, food_weights, max_bio_delta, max_carbon_delta,
#      Norm_features_matrix, protected_cells, stocks_info, stock_names, estimate_multiobjective_slopes, 
#      calculate_multiobjective_weights, max_food_delta_a1, max_food_delta_a2, MPAs_df,
#      carbon_remains_BAU, bio_remains_BAU, group_ids, fraction_in_MPA, MPAs,
#      file = file.path("13_efficiency_curve/ranking_inputs.RData"))
```

# Ranking 

```{r}
registerDoParallel(38)

run_efficiency_curve <- function(effort_assumption, bio_preferences){
  
  foreach(i = bio_preferences)%dopar%{
    
    source(here::here("functions", "food_provision_foos.R"))
    
    load(file.path(emLab_project_dir, "data", "03_output", "11_multiple_objectives", "ranking_inputs.RData"))
    
    unprotected_matrix <- Norm_features_matrix[!protected_cells, ]
    
    protected_matrix <- Norm_features_matrix[protected_cells, ]
    
    baseline_state <- protected_cells%*%Norm_features_matrix
    
    baseline_state[,"carbon"] <- baseline_state[, "carbon"] + carbon_remains_BAU
    
    baseline_state[, bio_feature_names] <- baseline_state[, bio_feature_names] + bio_remains_BAU
    
    max_slopes_bio <- 0.25*baseline_state[, bio_feature_names]^(0.25 - 1)
    
    max_slope_carbon <- 1
    
    names(max_slope_carbon) <- "carbon"
    
    if(i > 0){
      
      w_bio = 1/(1/i + 1)
      w_carbon = 0
      w_food =  1 - w_bio
      
    } else if (i <= 0){
      
      w_food = 1
      w_carbon = 0
      w_bio = i
      
    }
    
    multiobjective_feature_weigths <- calculate_multiobjective_weights(alpha_b = w_bio, alpha_c = w_carbon, alpha_f = w_food,
                                                                       b_wts = bio_weights, c_wts = carbon_weights, f_wts = food_weights, 
                                                                       max_b = max_bio_delta, max_c = max_carbon_delta, max_f_a1  = max_food_delta_a1,
                                                                       max_f_a2  = max_food_delta_a2, effort_assumption = effort_assumption)
    
    ranking <- list()
    
    current_pick <- 0
    
    current_state <- baseline_state 
    
    while(nrow(unprotected_matrix) > 0){
      
      current_pick <- sum(current_pick, 1)
      
      slopes <- estimate_multiobjective_slopes(x = current_state, feature_group = group_ids, feature_wts  = multiobjective_feature_weigths,
                                               max_slopes_b = max_slopes_bio, max_slopes_c = max_slope_carbon, e = stocks_info$ex_rate,
                                               m = stocks_info$m, k = stocks_info$k, r = stocks_info$r, effort_assumption = effort_assumption)
      
      delta <- unprotected_matrix%*%as.matrix(slopes)
      
      best_cell_indeces <- doBy::which.maxn(delta, n = 100)
      
      best_cells <- delta[best_cell_indeces,] %>% 
        enframe() %>% 
        set_names(c('cell_id', "delta")) %>% 
        mutate(pick_order = current_pick)
      
      ranking <- bind_rows(ranking, best_cells)
      
      current_state <- unprotected_matrix[best_cell_indeces, ] %>% 
        colSums() +
        current_state
      
      unprotected_matrix <- unprotected_matrix[-best_cell_indeces, ]
      
    }
    
    write_rds(ranking, 
              file.path(variant_dir, "runs",  paste0("ranking_a", as.character(effort_assumption), "_", as.character(i), ".rds")))
  }
}

efficiency_rankings_a1 <- tryCatch(run_efficiency_curve(effort_assumption = 1, bio_preferences = x_bios), error = function(e) print(e))
efficiency_rankings_a2 <- tryCatch(run_efficiency_curve(effort_assumption = 2, bio_preferences = x_bios), error = function(e) print(e))
```

# Summarize runs

```{r}
source(here::here("functions", "food_provision_foos.R"))
```

```{r}
get_full_ranking <- function(ranking_df, mpas_df, ocean_df){
  
  full_ranking <- mpas_df %>% 
    filter(!is.na(mpa)) %>% 
    select(cell_id) %>% 
    bind_rows(ranking_df) %>% 
    mutate(is_mpa = is.na(delta))  %>% 
    full_join(ocean_matrix) %>% 
    replace(is.na(.), 0) 
  
  return(full_ranking)
}

get_ranked_features_df <- function(ranking_df, features_df){
  
  ranked_features_df <- ranking_df %>% 
    left_join(features_df %>% 
                as_tibble(rownames = NA) %>% 
                rownames_to_column("cell_id")) %>% 
    replace(is.na(.), 0) 
  
    return(ranked_features_df)

}
```

```{r}
files <- list.files(file.path(variant_dir, "runs"), pattern = "ranking_a", full.names = T)

registerDoParallel(7)

get_efficiency_curve <- function(files){
  
  foreach(i = files, .combine = rbind)%dopar%{
  
  file <- i
  
  load(file.path(emLab_project_dir, "data", "03_output", "11_multiple_objectives", "ranking_inputs.RData"))
  
  effort_assumption <- as.numeric(str_extract(str_split(tools::file_path_sans_ext(basename(file)), pattern = "_", simplify = T)[,2], '[0-9]+'))
  
  run_bio_factor <- as.numeric(str_split(tools::file_path_sans_ext(basename(file)), pattern = "_", simplify = T)[,3])

  ranking <- read_rds(file)
  
  ranked_features_df <- ranking %>% 
    get_full_ranking(mpas_df = MPAs_df, ocean_df = ocean_matrix) %>% 
    get_ranked_features_df(features_df = Norm_features_matrix)
  
  optimal_protection <-  ranked_features_df %>% 
    select(delta) %>% 
    mutate(fraction_protected = 1/n(),
           fraction_protected = cumsum(fraction_protected),
           V = cumsum(delta)) %>% 
    slice(which.max(V)) %>%  
    pull(fraction_protected)
  
  food_benefit_curve <- ranked_features_df %>% 
    select(all_of(stock_names)) %>% 
    mutate_all(.funs = cumsum) %>% 
    apply(MARGIN = 1, FUN = estimate_delta_catch, e = stocks_info$ex_rate, m = stocks_info$m,
          r = stocks_info$r, k = stocks_info$k, effort_assumption = effort_assumption) %>% 
    t() %>% 
    rowSums() %>% 
    enframe(name = NULL, value = "delta_catch") %>% 
    mutate(fraction_protected = 1/n(),
           fraction_protected =  cumsum(fraction_protected))
  
  carbon_benefit_curve <- ranked_features_df %>% 
    select(carbon) %>% 
    mutate_all(.funs = cumsum) %>% 
    sweep(MARGIN = 2, carbon_remains_BAU, `+`) %>%
    mutate_all(.funs = ~(.)) %>% 
    rowSums() %>% 
    enframe(name = NULL, value = "carbon_benefit") %>% 
    mutate(carbon_benefit = scales::rescale(carbon_benefit, c(0,1)),
           fraction_protected = 1/n(),
           fraction_protected = cumsum(fraction_protected))
  
  bio_benefit_curve <-  ranked_features_df %>% 
    select(all_of(bio_feature_names)) %>% 
    mutate_all(.funs = cumsum) %>% 
    sweep(MARGIN = 2, bio_remains_BAU, `+`) %>%
    mutate_all(.funs = ~(.)^0.25) %>% 
    as.matrix() %*% as.matrix(bio_weights) %>% 
    as_tibble() %>% 
    set_names("bio_benefit") %>% 
    mutate(bio_benefit = scales::rescale(bio_benefit, c(0,1)),
           fraction_protected = 1/n(),
           fraction_protected = cumsum(fraction_protected)) 
  
  summary <- bio_benefit_curve %>% 
    left_join(carbon_benefit_curve) %>% 
    left_join(food_benefit_curve) %>% 
    filter(fraction_protected == optimal_protection) %>% 
    mutate(bio_x_factor = run_bio_factor,
           effort_assumption = effort_assumption)
  
  return(summary)
  }
  }

start <- Sys.time()
tryCatch(efficiency_curve <- get_efficiency_curve(files = files), error = function(e) print(e))
end <- Sys.time()
end-start

write_rds(efficiency_curve, file.path(variant_dir, "efficiency_curves.rds"))
```

# Results

```{r}
load(file.path(emLab_project_dir, "data", "03_output", "11_multiple_objectives", "ranking_inputs.RData"))

efficiency_curve <- read_rds(file.path(variant_dir, "efficiency_curves.rds"))
```

```{r}
efficiency_curve %>% 
  filter(bio_x_factor <= 100) %>% 
  ggplot()+
  geom_point(aes(x = bio_x_factor, y = bio_benefit, col = factor(effort_assumption)))+
  paletteer::scale_color_paletteer_d("wesanderson::Darjeeling2")+
  theme_minimal()+
  labs(x = "Biodiversity Weight", y = "Biodiversity Benefit", col = "Effort Assumption")+
  ggsave(file.path(emLab_project_dir, "figures", "exploratory", "efficiency_curve_x_bio_vs_bio_benefit.png"), width = 7, height = 5)
```

```{r}
efficiency_curve %>% 
  mutate(delta_catch = delta_catch/10^6,
         effort_assumption = factor(effort_assumption)) %>% 
  select(delta_catch, effort_assumption, fraction_protected, bio_benefit) %>% 
  pivot_longer(-c(delta_catch, effort_assumption), names_to = "x_var", values_to = "value") %>% 
  ggplot()+
  geom_point(aes(value, delta_catch, col = effort_assumption))+
  facet_wrap("x_var", ncol = 1, strip.position = "bottom", labeller = labeller(x_var = c("bio_benefit" = "Biodiversity Benefits", 
                                                                            "fraction_protected" = "Fraction Ocean Protected" )))+
  theme(strip.background = element_blank(), strip.placement = "outside")+
  labs(x = "", y = "Change in catch (MMT)") + 
  paletteer::scale_color_paletteer_d("wesanderson::Darjeeling2")+
  theme_minimal()+
  ggsave(file.path(emLab_project_dir, "figures", "exploratory", "efficiency_curves_comparison.png"), width = 7, height = 5)
```

```{r}
read_ranking_file <- function(run_pattern){
  
  ranking_df <- read_rds(list.files(file.path(variant_dir, "runs"), pattern = paste0(run_pattern, ".rds"), full.names = "T"))
  
  return(ranking_df)
}
  
create_ranking_raster <- function(ranking_df, MPAs_df, ocean_df, ocean_raster){
  
  ranking_raster <- MPAs_df %>% 
    filter(!is.na(mpa)) %>% 
    select(cell_id) %>% 
    bind_rows(ranking_df) %>% 
    mutate(is_mpa = is.na(delta)) %>% 
    left_join(ocean_matrix) %>% 
    tibble::rowid_to_column("ID") %>% 
    mutate(rank = 1- scales::rescale(ID)) %>% 
    select(x, y, rank) %>% 
    rasterFromXYZ(crs = crs(ocean_raster)) %>% 
    projectRaster(ocean_raster) %>% 
    mask(ocean_raster) 
  
  return(ranking_raster)
  
}

get_optimal_raster <- function(ranking_raster, optimal_protection){
  
  optimal_raster <- raster::mask(ranking_raster, mask = ranking_raster >= (1 - optimal_protection), maskvalue = 0)

  return(optimal_raster)
}

map_optimal_raster <- function(optimal_raster, caption, land_raster = land_50_moll){
  
  optimal_map <- optimal_raster %>% 
    tmap::tm_shape()+
    tmap::tm_raster(title = "",
                    palette = rev("-Spectral"), 
                    breaks = quantile(optimal_raster),
                    labels = c("lowest","","","", "highest"),
                    n = 20,
                    style = "cont",
                    legend.is.portrait = T, 
                    legend.reverse = T)+
    tmap::tm_shape(land_50_moll)+
    tmap::tm_fill(col = "grey", border.col = "transparent")+
    tmap::tm_shape(MPAs)+
    tmap::tm_raster(col = "fully_highly_protected_reviewed_MPAs", palette = "#00A600FF", legend.show	= F)+
    tmap::tm_layout(title = "Multiobjective Conservation Priorities",
                    title.position = c("left", .95),
                    inner.margins = c(0.12, 0, 0.08, 0.04), 
                    frame = F,
                    legend.position = c(.99, "center"))
  
  tmap::tmap_save(optimal_map, 
                  filename = file.path(emLab_project_dir, "figures", "supplement",  paste0("multiobjective_optim_map_",caption, ".png")),
                  dpi = 300,  height = 4, width = 8 )
  
  
}
```

## Cheapest way to 90% of biodviersity benefits

```{r}
efficiency_curve %>% 
  group_by(effort_assumption) %>% 
  slice(which.min(abs(bio_benefit - 0.9))) %>% 
  mutate(run_pattern = paste(effort_assumption, bio_x_factor, sep = "_"),
         caption =  paste0("cheapest_90_percent_bio_a", run_pattern, "x_bio_", round(fraction_protected, 2), "protected"),
         ranking_df = map(run_pattern, read_ranking_file),
         ranking_raster = map(ranking_df, create_ranking_raster,MPAs_df = MPAs_df, ocean_df = ocean_matrix, ocean_raster = ocean_low_res_moll),
         masked_raster = map2(ranking_raster, fraction_protected, .f = get_optimal_raster),
         optim_map = walk2(masked_raster, caption, .f = map_optimal_raster))
```

## Max bio benefit at "zero costs"

```{r}
efficiency_curve %>% 
  group_by(effort_assumption) %>% 
  slice(which.min(abs(delta_catch - 0))) %>% 
  mutate(run_pattern = paste(effort_assumption, bio_x_factor, sep = "_"),
         caption =  paste0("max_bio_zero_delta_catch_a", run_pattern, "x_bio_", round(fraction_protected, 2), "protected"),
         ranking_df = map(run_pattern, read_ranking_file),
         ranking_raster = map(ranking_df, create_ranking_raster,MPAs_df = MPAs_df, ocean_df = ocean_matrix, ocean_raster = ocean_low_res_moll),
         masked_raster = map2(ranking_raster, fraction_protected, .f = get_optimal_raster),
         optim_map = walk2(masked_raster, caption, .f = map_optimal_raster))
```

## Cheapest 30% of the ocean

```{r}
efficiency_curve %>% 
  group_by(effort_assumption) %>% 
  slice(which.min(abs(fraction_protected - 0.3))) %>% 
    mutate(run_pattern = paste(effort_assumption, bio_x_factor, sep = "_"),
         caption =  paste0("cheapest_30_percent_ocean_protected_a", run_pattern, "x_bio_", round(fraction_protected, 2), "protected"),
         ranking_df = map(run_pattern, read_ranking_file),
         ranking_raster = map(ranking_df, create_ranking_raster,MPAs_df = MPAs_df, ocean_df = ocean_matrix, ocean_raster = ocean_low_res_moll),
         masked_raster = map2(ranking_raster, fraction_protected, .f = get_optimal_raster),
         optim_map = walk2(masked_raster, caption, .f = map_optimal_raster))
```

## Indiferent to biodiversity

```{r}
efficiency_curve %>% 
  group_by(effort_assumption) %>% 
  slice(which.min(abs(bio_x_factor - 0))) %>% 
      mutate(run_pattern = paste(effort_assumption, bio_x_factor, sep = "_"),
         caption =  paste0("biodiversity_indifferent_a", run_pattern, "x_bio_", round(fraction_protected, 2), "protected"),
         ranking_df = map(run_pattern, read_ranking_file),
         ranking_raster = map(ranking_df, create_ranking_raster,MPAs_df = MPAs_df, ocean_df = ocean_matrix, ocean_raster = ocean_low_res_moll),
         masked_raster = map2(ranking_raster, fraction_protected, .f = get_optimal_raster),
         optim_map = walk2(masked_raster, caption, .f = map_optimal_raster))
```

## Two thirds biodiversity

```{r}
efficiency_curve %>% 
  group_by(effort_assumption) %>% 
  slice(which.min(abs(bio_benefit - 2/3))) %>% 
  mutate(run_pattern = paste(effort_assumption, bio_x_factor, sep = "_"),
         caption =  paste0("two_thirds_bio_a", run_pattern, "x_bio_", round(fraction_protected, 2), "protected"),
         ranking_df = map(run_pattern, read_ranking_file),
         ranking_raster = map(ranking_df, create_ranking_raster,MPAs_df = MPAs_df, ocean_df = ocean_matrix, ocean_raster = ocean_low_res_moll),
         masked_raster = map2(ranking_raster, fraction_protected, .f = get_optimal_raster),
         optim_map = walk2(masked_raster, caption, .f = map_optimal_raster))
         
```








