---
title: "Food provision in Juanation"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
---

```{r, message = FALSE}
library(raster)
library(sf)
library(tidyverse)
library(patchwork)

source(here::here("common.R"))

variant_name <- "16_food_alternative_effort_models"

variant_dir <- file.path(emLab_project_dir,"data", "03_output", variant_name)

dir.create(variant_dir)
```

# Load  data 

```{r, eval = F}
stocks <- tibble(filename = list.files(file.path(emLab_project_dir, "data", "02_processed", "stock_distributions"),
                                       full.names = T),
                 stockid = str_remove(basename(filename), "\\.tif"))

stocks_df <- stack(stocks$filename) %>%
  raster::as.data.frame(xy = T) %>%
  inner_join(ocean_matrix) %>% 
  select(x, y, cell_id, everything()) %>%
  as_tibble() 

data.table::fwrite(stocks_df, 
                   file.path(emLab_project_dir, "data", "02_processed", "features_df", "stocks_df.csv"))
```

```{r}
stocks <- tibble(filename = list.files(file.path(emLab_project_dir, "data", "02_processed", "stock_distributions"),
                                       full.names = T),
                 stockid = str_remove(basename(filename), "\\.tif"))

stocks_info <- read_csv(file.path(emLab_project_dir,"data", "02_processed", "food_provision", "MegaData_UncertaintyAnalysis.csv")) %>%
  select(stockid, k = Kfin, ex_rate = ExploitationRate_BAU1, m = m_fin, r = r_fin)

stocks_info <- stocks %>%
  left_join(stocks_info)

stocks_df <- data.table::fread(file.path(emLab_project_dir, "data", "02_processed", "features_df", "stocks_df.csv")) %>% 
  as_tibble()

stocks_info %>% filter(ex_rate >= 1)
```

```{r}
identical(stocks_info$stockid , str_replace_all(names(stocks_df)[-c(1,2,3)], "\\.", "-"))
```

## Normalize and trim

```{r}
features_matrix <- stocks_df %>%
  select(-x,-y,-cell_id) %>% 
  as.matrix()

stopifnot(
  sum(map_lgl(stocks_df %>%
                select(-x,-y,-cell_id), is.numeric)) == ncol(features_matrix)
  ) # ALL numeric?

Norm_features_matrix <- sweep(features_matrix, 2, colSums(features_matrix, na.rm = T), FUN = "/")

rownames(Norm_features_matrix) <- stocks_df$cell_id

stopifnot(
  sum(colSums(Norm_features_matrix, na.rm = T)) == ncol(features_matrix)
  )  # Normalized OK?

Norm_features_matrix <- Norm_features_matrix[rowSums(is.na(Norm_features_matrix)) != ncol(Norm_features_matrix), ]

stopifnot(
  identical(colnames(Norm_features_matrix), 
          stocks_df %>% 
            select(-x,-y,-cell_id) %>% 
            colnames())
  )  # Is the order of the features mantained?
```

# Current MPAs

```{r}
MPAs <- raster(file.path(emLab_project_dir, "data", "02_processed", "masks", "fully_highly_protected_reviewed_MPAs.tif"))

MPAs_df <- MPAs %>% 
  raster::as.data.frame(xy = T) %>%
  inner_join(ocean_matrix) %>% 
  select(x, y, cell_id, mpa = fully_highly_protected_reviewed_MPAs) %>%
  as_tibble() 

fraction_in_MPA <- sum(!is.na(MPAs_df$mpa))/nrow(MPAs_df)
  
MPA_vect <- rownames(Norm_features_matrix) %>% 
  enframe() %>% 
  select(cell_id = value) %>% 
  left_join(MPAs_df) %>% 
  mutate(mpa =  !is.na(mpa)) %>% pull(mpa)
```

# Ranking 

```{r}
Norm_features_matrix[is.na(Norm_features_matrix)] <- 0

protected_cells <- matrix(MPA_vect, nrow = 1, ncol = nrow(Norm_features_matrix))

colnames(protected_cells) <- rownames(Norm_features_matrix)

baseline_state <- protected_cells%*%Norm_features_matrix

stocks_info$ex_rate[stocks_info$ex_rate >= 1] <- 0.99999
```

## Effort goes outside

### functions

```{r}
estimate_slopes_a1 <- function(e = stocks_info$ex_rate, 
                               m = stocks_info$m,
                               r = stocks_info$r, 
                               k = stocks_info$k, 
                               R){
  
  R[R > 1] <- 1 
  
  A <- (1 - e)^(1/(1 - R))
  
  part_a <- A*log(1 - e)*k*m*(1 - (1 - A)*m*(1 - R)/(r*((1 - A)*R + m)))/((1 - R)*((1 - A)*R + m))
  
  part_b <- (1 - A)*k*m*(1 - (1 - A)*m*(1 - R)/(r*((1 - A)*R + m)))/((1 - A)*R + m)
  
  part_c <- (1 - A)*k*m*(1 - R)*(- A*log(1 - e)*R/(1 - R)^2 - A + 1)*(1 - (1 - A)*m*(1 - R)/(r*((1 - A)*R + m)))/((1 - A)*R + m)^2
  
  part_d <- (1 - A)*k*m*(1 - R)*(A*log(1 - e)*m/(r*(1 - R)*((1 - A)*R + m)) + (1 - A)*m/(r*((1 - A)*R + m)) + (1 - A)*m*(1 - R)*(-A*log(1 - e)*R/(1 - R)^2 - A + 1)/(r*((1 - A)*R + m)^2))/((1 - A)*R + m)
  
  slope <- - part_a - part_b - part_c + part_d
  
  slope[R == 1] <- -9*10^9
  
  return(slope)
  
}

estimate_delta_catch_a1 <- function(x, 
                                    e = stocks_info$ex_rate, 
                                    m = stocks_info$m,
                                    r = stocks_info$r, 
                                    k = stocks_info$k) {
  
  R_bau <- 0
  
  x[x > 1] <- 1 
  
  exp_rate <- 1 - (1 - e)^(1/(1 - x))
  
  part_a <- exp_rate*(m*k)*(1 - x)/(exp_rate*x + m)*(1 - exp_rate*(1 - x)*m/((exp_rate*x + m)*r))
  
  part_b <- e*(m*k*(1 - R_bau)/(e*R_bau + m)*(1 - e*(1 - R_bau)*m/((e*R_bau + m)*r)))
  
  return(part_a - part_b)
  
}
```

### run algorithm

```{r}
unprotected_matrix <- Norm_features_matrix[!protected_cells, ]

protected_matrix <- Norm_features_matrix[protected_cells, ]

ranking <- list()

current_pick <- 0

current_state <- baseline_state 

start_time <- Sys.time()

while(nrow(unprotected_matrix) > 0){
  
  current_pick <- sum(current_pick, 1)
  
  slopes <- estimate_slopes_a1(R = current_state)
  
  delta <- unprotected_matrix%*%as.matrix(t(slopes))
  
  best_cell_indeces <- doBy::which.maxn(delta, n = 100)
  
  best_cells <- delta[best_cell_indeces,] %>% 
    enframe() %>% 
    set_names(c('cell_id', "delta")) %>% 
    mutate(pick_order = current_pick)
  
  ranking <- bind_rows(ranking, best_cells)
  
  current_state <- unprotected_matrix[best_cell_indeces, ] %>% 
    colSums() +
    current_state
  
  unprotected_matrix <- unprotected_matrix[-best_cell_indeces, ]
  
}

end_time <- Sys.time()

end_time - start_time

write_rds(ranking,
          file.path(variant_dir, "ranking_results_a1.rds"))
```

# Results

## benefit curve

```{r}
ranking <- read_rds(file.path(variant_dir, "ranking_results_a1.rds"))

ranked_rep_levels <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(ranking) %>% 
  mutate(is_mpa = is.na(delta)) %>% 
  full_join(ocean_matrix) %>% 
  left_join(Norm_features_matrix %>% 
              as_tibble(rownames = NA) %>% 
              rownames_to_column("cell_id")) %>% 
  replace(is.na(.), 0)

network_delta_catch_per_stock <- ranked_rep_levels %>% 
  select(-x, -y, -cell_id, -delta, -pick_order, -is_mpa) %>% 
  mutate_all(.funs = cumsum) %>% 
  apply(MARGIN = 1, estimate_delta_catch_a1)  %>% 
  t() 

network_delta_catch <- network_delta_catch_per_stock %>% 
  rowSums() %>% 
  enframe() 
  
pixel_delta_catch <- ranked_rep_levels %>%
  select(-x, -y, -cell_id, -delta, -pick_order, -is_mpa) %>%
  apply(MARGIN = 1, estimate_delta_catch_a1)  %>% 
  t() %>%
  rowSums() %>%
  enframe()

results_per_pixel <- ranked_rep_levels %>% 
  select(x, y, is_mpa) %>% 
  cbind(pixel_delta_catch = pixel_delta_catch$value,
        network_cummulative_delta_catch = network_delta_catch$value) %>% 
  mutate(pixel_contribution_to_network = network_cummulative_delta_catch - lag(network_cummulative_delta_catch),
         pixel_contribution_to_network = if_else(is.na(pixel_contribution_to_network), network_cummulative_delta_catch, pixel_contribution_to_network)) 

write_csv(results_per_pixel, 
          file.path(variant_dir, "results_per_pixel_a1.csv"))

network_delta_catch_per_stock %>% 
  as_tibble() %>% 
  mutate(cell_id = ranked_rep_levels$cell_id) %>%
  select(cell_id, everything()) %>% 
  write_rds(file.path(variant_dir, "network_delta_catch_per_stock_a1.rds"))
```

```{r}
network_delta_catch %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected =  cumsum(fraction_protected)) %>% 
  filter(fraction_protected > fraction_in_MPA) %>% 
  ggplot()+
  geom_line(aes(fraction_protected, value))+
  labs(x = "Fraction Ocean Protected",
       y = "Delta Catch",
       caption = "a1, all effort is displaced") +
  theme_classic()+
  geom_rect(aes(xmin = 0, 
                xmax = fraction_in_MPA, 
                ymin = min(network_delta_catch$value), 
                ymax = max(network_delta_catch$value)),
                fill = "#69B578", alpha = 0.01)+
  scale_x_continuous(limits = c(0,1), expand = c(0, 0))+
  scale_y_continuous(expand = c(0, 100))+
  ggsave(file.path(variant_dir, "delta_catch_curve_a1.png"),
         width = 7, 
         height = 4)

results_per_pixel %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected =  cumsum(fraction_protected)) %>% 
  ggplot()+
  geom_line(aes(fraction_protected, pixel_contribution_to_network))+
  labs(x = "Fraction Ocean Protected",
       y = "Delta Catch",
       caption = "Juanation")+
  theme_minimal()
```

## max delta catch

```{r}
max(network_delta_catch$value, na.rm = T)/10^6

bau_delta_catch <- network_delta_catch %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected =  cumsum(fraction_protected)) %>% 
  filter(fraction_protected <= fraction_in_MPA) %>% 
  summarise(current_delta_catch = max(value)) %>% pull()

(max_delta_catch <- max(network_delta_catch$value, na.rm = T) - bau_delta_catch)/10^6

optimal_protection_level <- network_delta_catch %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected =  cumsum(fraction_protected)) %>% 
  filter(value == max(network_delta_catch$value)) %>% 
  pull(fraction_protected)
```

```{r}
diff_90 <- bau_delta_catch + 0.9*(max_delta_catch)

results_per_pixel %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected =  cumsum(fraction_protected)) %>% 
  filter(fraction_protected < optimal_protection_level,
         fraction_protected > fraction_in_MPA) %>% 
  slice(which.min(abs(network_cummulative_delta_catch - diff_90))) %>% 
  pull(fraction_protected)
```

## maps

### global ranking

```{r}
ranking_raster <- MPAs_df %>% 
  filter(!is.na(mpa)) %>% 
  select(cell_id) %>% 
  bind_rows(ranking) %>% 
  mutate(is_mpa = is.na(delta)) %>% 
  left_join(ocean_matrix) %>% 
  tibble::rowid_to_column("ID") %>% 
  mutate(rank = 1- scales::rescale(ID)) %>% 
  select(x, y, rank) %>% 
  rasterFromXYZ(crs = crs(ocean_low_res_moll)) %>% 
  projectRaster(ocean_low_res_moll) %>% 
  mask(ocean_low_res_moll) 

writeRaster(ranking_raster, 
            file.path(variant_dir, "ranking_raster.tif"), 
            overwrite = T)

ranking_map <- tmap::tm_shape(ranking_raster)+
  tmap::tm_raster(title = "",
                  palette  = priorities_map_pal$colors,
                  breaks = priorities_map_pal$breaks,
                  labels = priorities_map_pal$labels,
                  legend.is.portrait = T, 
                  legend.reverse = T)+
  tmap::tm_shape(land_50_moll)+
  tmap::tm_fill(col = "black", border.col = "transparent")+
  tmap::tm_credits("Food") +
  tmap::tm_shape(MPAs)+
  tmap::tm_raster(col = "fully_highly_protected_reviewed_MPAs", palette = "#00A600FF", legend.show	= F)+
  tmap::tm_layout(title = "Food Provision Conservation Priorities",
                  title.position = c("center", .95),
                  inner.margins = c(0.12, 0, 0.08, 0.04), 
                  frame = F,
                  legend.position = c(.99, "center")) 

tmap::tmap_save(ranking_map, 
                filename = file.path(variant_dir, "priority_map.png"),
                dpi = 300,  height = 4, width = 8 )
```

### optimal food protection map

```{r}
optimal_protection_map <- ranking_raster %>% 
  mask(ranking_raster >= (1 - optimal_protection_level), maskvalue = 0) %>% 
  tmap::tm_shape()+
  tmap::tm_raster(title = "",
                  palette = rev("-Spectral"), 
                  n = 20,
                  
                  style = "cont",
                  legend.is.portrait = T, 
                  legend.reverse = T)+
  tmap::tm_shape(land_50_moll)+
  tmap::tm_fill(col = "grey", border.col = "transparent")+
  tmap::tm_credits("Food") +
  tmap::tm_shape(MPAs)+
  tmap::tm_raster(col = "fully_highly_protected_reviewed_MPAs", palette = "#00A600FF", legend.show	= F)+
  tmap::tm_layout(title = "Food Provision Conservation Priorities",
                  title.position = c("center", .95),
                  inner.margins = c(0.12, 0, 0.08, 0.04), 
                  frame = F,
                  legend.position = c(.99, "center"))

tmap::tmap_save(optimal_protection_map, 
                filename = file.path(variant_dir, "optimal_protection_map.png"),
                dpi = 300,  height = 4, width = 8 )
```

### pixel contribution to network

```{r}
results_per_pixel %>% 
  mutate(tmp = abs(pixel_contribution_to_network)^(1/3),
         tmp_2 = ifelse(pixel_contribution_to_network < 0, -tmp, tmp)) %>% 
  ggplot()+
  geom_raster(aes(x, y, fill = tmp_2))+
  scale_fill_gradient2(labels = function(x){x^3},
                       low = "darkred", mid = "white",
                            high = "navy", space = "Lab")+
  labs(title = "Change in global catch (Tonnes)", fill = "", y = "", x = "")+
  theme_minimal()+
  ggsave(file.path(variant_dir, "pixel_contribution_map.png"),dpi = 300,  height = 4, width = 8)
```

## Delta H per stock

Here we examine the delta H for each stock at 10% of optimal ocean protection

```{r}
network_delta_catch_per_stock_A1 <- read_rds(file.path(variant_dir, "network_delta_catch_per_stock_a1.rds"))

protection_levels_per_stock <- ranked_rep_levels %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected =  cumsum(fraction_protected)) %>% 
  filter(fraction_protected < 0.1) %>% 
  select(-x, -y, -fraction_protected, -delta, -pick_order, -is_mpa) %>% 
  mutate_at(vars(-cell_id), .funs = cumsum) %>% 
  pivot_longer(-c(cell_id), names_to = "stockid", values_to = "f_stock_protected")

delta_catch_per_stock_at_10_percent_protection_A1 <- network_delta_catch_per_stock_A1 %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected =  cumsum(fraction_protected)) %>% 
  filter(fraction_protected < 0.1) %>% 
  select(-fraction_protected) %>% 
  pivot_longer(-c(cell_id), names_to = "stockid", values_to = "delta_catch_A1") %>% 
  group_by(stockid) %>% 
  summarise(dH_A1 = last(delta_catch_A1),
            cell_id = cell_id[which.max(delta_catch_A1)]) %>% 
  left_join(protection_levels_per_stock, by = c("cell_id", "stockid")) %>% 
  select(-cell_id) %>% 
  mutate(stockid = str_replace_all(stockid, "\\.", "-")) %>% 
  left_join(stocks_info %>% 
              select(-filename), by = "stockid") %>% 
  mutate(E_msy = 0.5*r,
         msy = r*k/4,
         m = as.factor(m),
         e_over_msy = ex_rate/E_msy,
         dH_over_k_A1 = dH_A1/k,
         dH_over_msy_A1 = dH_A1/msy)

stock_range_sizes <- ranked_rep_levels %>% 
  select(-x, -y, -cell_id, -delta, -pick_order, -is_mpa) %>% 
  summarize_all(~sum(.x > 0)) %>% 
  pivot_longer(everything(), names_to = "stockid", values_to = "n_cells") %>% 
  mutate(stockid = str_replace_all(stockid, "\\.", "-")) 

delta_catch_per_stock_at_10_percent_protection_A1 <- delta_catch_per_stock_at_10_percent_protection_A1 %>% 
  left_join(stock_range_sizes)

write_csv(delta_catch_per_stock_at_10_percent_protection_A1, 
          file.path(variant_dir, "delta_catch_per_stock_at_10_percent_protection_A1.csv"))
```


```{r}
delta_catch_per_stock_at_10_percent_protection_A1 %>% 
  mutate(m = as.factor(m)) %>% 
  select(stockid, m, r, ex_rate, e_over_msy,  n_cells) %>% 
  pivot_longer(-c(stockid, m, r), names_to = c("x_var"), values_to = "x_val") %>% 
  left_join(delta_catch_per_stock_at_10_percent_protection_A1 %>% 
              mutate(dH_A1 = dH_A1/10^6) %>% 
              select(stockid, dH_A1 , dH_over_k_A1, dH_over_msy_A1) %>% 
              pivot_longer(-c(stockid), names_to = c("y_var"), values_to = "y_val")) %>% 
  ggplot()+
  geom_point(aes(x = x_val, y = y_val, col = r, shape = m))+
  facet_grid(rows = vars(y_var), cols = vars(x_var), scales = 'free', switch="both", 
             labeller = labeller(x_var = c(ex_rate = "Exploitation rate", e_over_msy = "E/E_msy", n_cells = "Stock range size"),
                                 y_var = c(dH_A1 = "Delta H", dH_over_k_A1 = "Delta H/K", dH_over_msy_A1 = "Delta H/MSY")))+
  labs(x = "", y = "")+
  theme(
    aspect.ratio = 1,
    strip.background = element_blank(),
    strip.placement = "outside",
    plot.title.position = "plot"
  )+
  labs(title = "Delta catch per stock at 10% optimal ocean protection", subtitle = "Full displacement model")+
  scale_color_viridis_c()+
  ggsave(filename = file.path(variant_dir, "deltaH_diagnostic_plots.png"), height = 10, width = 12)
```

```{r}
delta_catch_per_stock_at_10_percent_protection_A1 %>% 
  ggplot()+
  geom_point(aes(x = e_over_msy, y = f_stock_protected, col = dH_over_k_A1, size = dH_A1/10^6))+
  scale_color_gradient2()+
  theme_dark()+
  labs(x = "E/E_msy", y = "fraction of stock protected", title = "Delta catch (dH) and protection levels per stock", subtitle = "At 10% optimal ocean protection - full displacement model", color = "dH/K", size = "dH")+
  theme(plot.title.position = "plot")+
  ggsave(filename = file.path(variant_dir, "deltaH_and_protection_levels_at_10_percent.png"))
```

# Comparing effort assumptions

## at 10% protection

```{r}
delta_catch_per_stock_at_10_percent_protection_A0 <- read_csv(file.path(emLab_project_dir, "data", "03_output", "07_food", 
                                                                         "delta_catch_per_stock_at_10_percent_protection.csv"))

comparing_asssumptions_df <- delta_catch_per_stock_at_10_percent_protection_A1 %>% 
  left_join(delta_catch_per_stock_at_10_percent_protection_A0 %>% 
              select(stockid, f_stock_protected_A0 = f_stock_protected, dH_A0 = dH, dH_over_k_A0 = dH_over_k, dH_over_msy_A0 = dH_over_msy))
  
comparing_asssumptions_df %>% 
  ggplot()+
  geom_point(aes(x = e_over_msy, y =  f_stock_protected - f_stock_protected_A0, 
                 shape = m, col =  dH_over_k_A1 - dH_over_k_A0),
             alpha = 0.5)+
  scale_color_gradient2()+
  labs(x = "E / E_msy", y = "Difference in fraction protected", col = "Difference in dH")+
  theme_dark()
```

## benefit curves

```{r}
network_delta_catch_per_stock_A1 <- read_rds(file.path(variant_dir, "network_delta_catch_per_stock_a1.rds"))
network_delta_catch_per_stock_A0 <- read_rds(file.path(emLab_project_dir, "data", "03_output", "07_food","network_delta_catch_per_stock.rds"))

benefit_curve_A0 <-  network_delta_catch_per_stock_A0 %>% 
  select(-cell_id) %>% 
  rowSums() %>% 
  enframe(name = "pick", value = "delta_catch_A0") 

benefit_curve_A1 <-  network_delta_catch_per_stock_A1 %>% 
  select(-cell_id) %>% 
  rowSums() %>% 
  enframe(name = "pick", value = "delta_catch_A1")     

left_join(benefit_curve_A0, benefit_curve_A1) %>% 
  mutate(fraction_protected = 1/n(),
         fraction_protected =  cumsum(fraction_protected)) %>% 
  pivot_longer(cols = c("delta_catch_A1", "delta_catch_A0"), names_to = "effort_model", values_to = "delta_catch") %>% 
  ggplot()+
  geom_line(aes(x = fraction_protected, y = delta_catch/10^6, col = effort_model))
  ggsave(filename = file.path(variant_dir, "benefit_curve_comparison_a0_vs_a1.png"), width = 6, height = 5)
```

## curves by ...

```{r}
network_delta_catch_per_stock_A1 <- read_rds(file.path(variant_dir, "network_delta_catch_per_stock_a1.rds"))

delta_H_per_stock_per_step <- network_delta_catch_per_stock_A1 %>% 
  pivot_longer(-c(cell_id), names_to = "stockid", values_to = "delta_H")

protection_per_stock_per_step <- ranked_rep_levels %>% 
  select(-x, -y, -delta, -pick_order, -is_mpa) %>% 
  mutate_at(vars(-cell_id), .funs = cumsum) %>% 
  pivot_longer(-c(cell_id), names_to = "stockid", values_to = "f_stock_protected")

protection_and_delta_H_per_step <- delta_H_per_stock_per_step %>% 
  left_join(protection_per_stock_per_step)

protection_and_delta_H_per_step <- protection_and_delta_H_per_step %>% 
  mutate(stockid = str_replace_all(stockid, "\\.", "-")) %>% 
  left_join(stocks_info %>% 
              select(-filename), by = "stockid") %>% 
  mutate(E_msy = 0.5*r,
         msy = r*k/4,
         m = as.factor(m),
         e_over_msy = ex_rate/E_msy,
         delta_H_over_k = delta_H/k,
         delta_H_over_msy = delta_H/msy)

protection_and_delta_H_per_step %>% 
  write_rds(file.path(variant_dir, "protection_and_delta_H_per_step.rds"))
```


```{r}
protection_and_delta_H_per_step <- read_rds(file.path(variant_dir, "protection_and_delta_H_per_step.rds"))

protection_and_delta_H_per_m <- protection_and_delta_H_per_step %>% 
  group_by(cell_id, m) %>% 
  summarize(delta_H = sum(delta_H),
            avg_protection = mean(f_stock_protected)) %>% 
  left_join(ranked_rep_levels %>% 
              mutate(fraction_ocean_protected = 1/n(),
                     fraction_ocean_protected =  cumsum(fraction_ocean_protected)) %>% 
              select(cell_id, fraction_ocean_protected))

protection_and_delta_H_per_m %>% 
  ungroup() %>% 
  select(-cell_id) %>% 
  pivot_longer(cols = -c(m,fraction_ocean_protected), names_to = "variable") %>% 
  ggplot()+
  geom_line(aes(x = fraction_ocean_protected, y = value, col = m))+
  facet_wrap(vars(variable), scales = 'free', ncol = 1, strip.position = "left")+
  scale_color_viridis_d()+
  labs(y = "")+
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        plot.title.position = "plot")+
  ggsave(filename = file.path(variant_dir, "curves_by_m.png"))
```

```{r}
protection_and_delta_H_per_E <- protection_and_delta_H_per_step %>% 
  mutate(e_over_msy = cut(e_over_msy, 
                          breaks = c(0, .4, .8, 1.2, 1.6, 2.1),
                          include.lowest = TRUE, 
                          right = FALSE)) %>% 
  group_by(cell_id, e_over_msy) %>% 
  summarize(delta_H = sum(delta_H),
            avg_protection = mean(f_stock_protected)) %>% 
  left_join(ranked_rep_levels %>% 
              mutate(fraction_ocean_protected = 1/n(),
                     fraction_ocean_protected =  cumsum(fraction_ocean_protected)) %>% 
              select(cell_id, fraction_ocean_protected))

protection_and_delta_H_per_E %>% 
  ungroup() %>% 
  select(-cell_id) %>% 
  pivot_longer(cols = -c(e_over_msy, fraction_ocean_protected), names_to = "variable") %>% 
  ggplot()+
  geom_line(aes(x = fraction_ocean_protected, y = value, col = e_over_msy))+
  facet_wrap(vars(variable), scales = 'free', ncol = 1, strip.position = "left")+
  scale_color_viridis_d()+
  labs(y = "")+
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        plot.title.position = "plot")+
  ggsave(filename = file.path(variant_dir, "curves_by_e_over_e_msy.png"))
```

```{r}
protection_and_delta_H_per_step %>% 
  mutate(e_over_msy = cut(e_over_msy, 
                          breaks = c(0, .4, .8, 1.2, 1.6, 2.1),
                          include.lowest = TRUE, 
                          right = FALSE),
         delta_H = delta_H/10^6) %>% 
  ggplot()+
  geom_line(aes(x = f_stock_protected, y = delta_H, group = stockid, col = r))+
  facet_grid(rows = vars(e_over_msy), cols = vars(m), scales = 'free')+
  ggsave(filename = file.path(variant_dir, "curves_per_stock.png"), width = 10, height = 10)
```

