---
title: "Get the list of species to use for the biodiversity component of the analysis"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
---

```{r setup, message=F,warning=F}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',message = FALSE, warning = FALSE)

library(raster)
library(tidyverse)
library(sf)
library(rredlist)
library(future)
library(furrr)
library(parallel)

Sys.setenv(IUCN_REDLIST_KEY = "0b64bc7473e4e6dd698b42d5d239ff12786d3007a30aa4aa30064ae139d79bbe")

emLab_project_dir <- "~/Google Drive File Stream/Shared drives/emlab/projects/current-projects/ocean-conservation-priorities/data/"

emLab_shrd_data_dir <- "~/Google Drive File Stream/Shared drives/emlab/data/"
```

The list of species to be included in our analysis includes exploited species (RAM + FAO + Watson's), and those that have been assessed by IUCN having fishing and habitat destruction as a threat. 

# Get individual lists

## IUCN 

Get all IUCN species and their habitats
```{r}
plan(multiprocess)

boot_df <- function(x) x[sample(nrow(x), replace = T), ]
rsquared <- function(mod) summary(mod)$r.squared
boot_lm <- function(i) {
  rsquared(lm(mpg ~ wt + disp, data = boot_df(mtcars)))
}

system.time(future_map(1:500, boot_lm))

system.time(parallel::mclapply(1:500, boot_lm, mc.cores = 4))
```

```{r}
all_sp <- rl_sp(all = TRUE)

all_sp <- do.call(rbind, lapply(all_sp, "[[", "result"))

get_habitat_codes <- function(sci_name){paste(rl_habitats(sci_name)$result$code, collapse = ",")}

habitats <- parallel::mclapply(all_sp$scientific_name, get_habitat_codes, mc.cores = 10)
```



```{r, eval = F}
all_sp <- rl_sp(all = TRUE)

all_sp <- do.call(rbind, lapply(all_sp, "[[", "result"))

get_habitat_codes <- function(sci_name){paste(rl_habitats(sci_name)$result$code, collapse = ",")}

plan(multiprocess)

all_sp <- all_sp %>% 
  mutate(habitats = furrr::future_map_chr(scientific_name, 
                                          get_habitat_codes))

saveRDS(all_sp, 
        file.path(emLab_project_dir, "01_raw", "IUCN_spp_list_with_habitats.rds"))
```


```{r, eval = F}
all_sp <- read_rds(file.path(DIR_GCS, "Biodiversity/data/preprocessed/IUCN_spp_list_with_habitats.rds"))
```

### classify as marine

Based on their habitat classification, which of these is marine?

```{r, eval = F}
iucn_habs <- read_csv(file.path(DIR_GCS, "Biodiversity/data/raw/species-list/iucn_habitat_codes.csv"))
```

The full list of iucn habitats was manually assigned as marine or not marine, as follows:

```{r, eval = F}
iucn_habs
```

```{r, eval = F}
(marine_hab_codes <- iucn_habs$code[iucn_habs$marine == T])
```

```{r, eval = F}
all_sp <- all_sp %>% 
  mutate(habitats = if_else(habitats == "", NA_character_, habitats)) %>% 
  rowwise() %>% 
  mutate(marine_habs = if_else(is.na(habitats), NA, any(unlist(strsplit(habitats, ",")) %in% marine_hab_codes)))
```

For how many species we have no habitat info for? `r nrow(filter(all_sp, is.na(habitats)))` 

```{r, eval = F}
all_sp %>% 
  filter(is.na(habitats)) %>% 
  group_by(class_name) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

For how many species we have Unknown habitat info for? `r nrow(filter(all_sp, habitats %in% c("17","18")))` 

```{r, eval = F}
all_sp %>% 
  filter(category != "EX") %>% 
  filter(habitats ==  "17" | habitats == "18") %>% 
  group_by(class_name) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

Most of these species with missing or unknown habitat are definitely not marine. So we need to apply an additional filter. For this, we download a list from IUCN of species cataloged as Marine systems. This list contains 13414 records and was downloaded from IUCN on December 9th 2018. 

```{r, eval = F}
iucn_marine_system <- read_csv(file.path(DIR_GCS, "Biodiversity/data/raw/species-list/iucn_website_marine_spp.csv"))

all_sp <- all_sp %>% 
  rowwise() %>% 
  mutate(marine_system = taxonid  %in% iucn_marine_system$internalTaxonId) 
```

What species are marine system but not marine habitat?

```{r, eval = F}
all_sp %>% 
  filter(marine_system & marine_habs == F & !is.na(habitats))
```

Now what species have marine, missing or unknow habitats AND are classified as marine systems?

```{r, eval = F}
all_sp %>% 
  filter((is.na(habitats) | marine_habs) & marine_system)
```

We include species that are listed as "Marine System" AND their habitat is of a marine type or unknown. 

```{r, eval = F}
marine_spp <- all_sp %>% 
  filter((is.na(habitats) | marine_habs) & marine_system)
```

In total our sample of marine species is `r nrow(marine_spp)`

```{r, eval = F}
marine_spp %>% 
  group_by(phylum_name,class_name) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

Now, how many of these have fishing and/or habitat destruction as a threat?

### get threats

```{r, eval = F}
get_threats_codes <- function(sci_name){paste(rl_threats(sci_name)$result$code, collapse = ",")}
```

```{r eval = F}
plan(multiprocess)

tictoc::tic()
marine_spp <- marine_spp %>% 
  mutate(threats = furrr::future_map(scientific_name, 
                                get_threats_codes))
tictoc::toc()

saveRDS(marine_spp, file.path(DIR_GCS, "Biodiversity/data/preprocessed/marine_spp_with_habitats_and_threats.rds"))
```

#### what threats are there?

```{r}
marine_spp <- read_rds(file.path(DIR_GCS, "Biodiversity/data/preprocessed/marine_spp_with_habitats_and_threats.rds"))
```

```{r, eval = F}
iucn_threats <- read_csv(file.path(DIR_GCS, "Biodiversity/data/raw/species-list/iucn_threat_codes.csv"))

iucn_threats$code <- str_trim(iucn_threats$code)

iucn_threats
```

##### Abatable threats 

```{r}
detect_threat <- function(threat_list, threat_code){
  
  threat_vector <- unlist(threat_list)
  
  code <- paste0("^", threat_code, collapse = "|")
  
  split_v <- str_split(threat_vector, ",")
  
  detected_vector <- split_v %>% 
    map(str_detect, code) %>% 
    unlist()
  
  detected <- sum(detected_vector) > 0
  
  detected
}
```

```{r}
marine_spp_abatable_threats <- marine_spp %>% 
  mutate(threat_group = case_when(
    detect_threat(threats, 5.4) ~ "fishing",
    detect_threat(threats, 2.4) ~ "aquaculture",
    detect_threat(threats, 3.1) ~ "oil_and_gas",
    detect_threat(threats, 3.2) ~ "mining",
    detect_threat(threats, 5.3) ~ "logging",
    detect_threat(threats, c(1.1, 1.2)) ~ "development"
  )) %>% 
  filter(!is.na(threat_group))
```

```{r}
marine_spp_abatable_threats %>% 
  group_by(threat_group) %>% 
  summarize(n_spp = n()) %>% 
  arrange(desc(n_spp))
```

```{r}
IUCN_spp_list <- marine_spp_abatable_threats %>% 
  rename(kingdom = kingdom_name, phylum = phylum_name, class = class_name, order = order_name, family = family_name, genus = genus_name) %>% 
  mutate_at(vars(kingdom, phylum, class, order, family), stringr::str_to_title)
```

### Resolve scientific names

```{r}
sources <- c(worms = 9, fishbase = 155, eol = 12, col = 1, gbif = 11)

IUCN_resolved_names <- sources %>% 
  map(~ taxize::gnr_resolve(data_source_ids = .x, 
                            names = IUCN_spp_list$scientific_name,
                            best_match_only = T, 
                            fields = c("all"), 
                            canonical = T ))

IUCN_resolved_names_df <- IUCN_resolved_names %>% 
  map(~ .x %>% 
        filter(!match_value %in% c("Could only match genus") & str_count(matched_name2, "\\w+") >= 2) %>% 
        select(user_supplied_name, matched_name2, taxon_id)) %>% 
  reduce(full_join, by = "user_supplied_name") %>% 
  set_names(c("user_supplied_name", "worms_sci_name", "worms_id", "fishbase_sci_name", "fishbase_id", "eol_sci_name", "eol_id", "col_sci_name", "col_id","gbif_sci_name", "gbif_id"))

IUCN_resolved_names <- IUCN_resolved_names_df %>% 
  mutate(resolved_scientific_name = case_when(
    !is.na(worms_sci_name) ~ worms_sci_name,
    !is.na(fishbase_sci_name) ~ fishbase_sci_name,
    !is.na(eol_sci_name) ~ eol_sci_name,
    !is.na(col_sci_name) ~ col_sci_name,
    TRUE ~ user_supplied_name
  )) %>% 
  select(user_supplied_name, resolved_scientific_name, everything()) 
```

```{r}
IUCN_spp_list <- IUCN_spp_list %>% 
  select(iucn_sci_name = scientific_name, iucn_id = taxonid, iucn_category = category, habitats, threats, threat_group) %>% 
  left_join(IUCN_resolved_names %>% 
              select(user_supplied_name, resolved_sci_name = resolved_scientific_name), 
            by = c("iucn_sci_name" = "user_supplied_name")) %>% 
  select(iucn_sci_name, resolved_sci_name, everything())
```

```{r}
IUCN_spp_list %>% 
  filter(iucn_sci_name != resolved_sci_name)

IUCN_spp_list %>% map_df(~sum(is.na(.)))
```

```{r}
IUCN_spp_list <- IUCN_spp_list %>% 
  mutate(resolved_sci_name = coalesce(resolved_sci_name, iucn_sci_name)) 

IUCN_spp_list$resolved_sci_name[IUCN_spp_list$iucn_sci_name == "Pagrus auratus"] <- 'Chrysophrys auratus'
```

### validate names and get taxonomic info from WoRMS

```{r}
get_worms_info <- function(sci_name){
  
  worrms::wm_records_name(name = sci_name) %>% 
    mutate(provided_sci_name = sci_name) %>% 
    select(provided_sci_name, scientificname, status, valid_name, kingdom, phylum, class, order, family, genus)
  
}
```

```{r}
plan(multiprocess)

IUCN_worms_info <- IUCN_spp_list %>%
  pull(resolved_sci_name) %>% 
  furrr::future_map(safely(get_worms_info)) %>% 
  transpose()

IUCN_worms_info <- IUCN_worms_info$result %>% 
  compact() %>% 
  bind_rows()
```

Lets take those entries that have no NAs, where the name queried matches the name supplied, and when there are only one valid name per entry

```{r}
valid_IUCN_worms_info <- IUCN_worms_info %>% 
  filter(!is.na(scientificname), !is.na(valid_name)) %>% # delete entries with NAs
  group_by(provided_sci_name) %>% 
  filter(provided_sci_name == scientificname) %>% # make sure to keep only rows that match the sci_name we supplied
  filter(n_distinct(valid_name, na.rm = T) == 1) %>%  # get those for which there is only one valid name
  ungroup() %>% 
  group_by(provided_sci_name) %>% 
  summarise_at(vars(valid_name, kingdom, phylum, class, order, family, genus), .funs = unique)
```

Now, from those with multiple valid names per row, lets get only the row for which status == "accepted", which means that the supplied name was already valid

```{r}
accepted_IUCN_worms_info <- IUCN_worms_info %>% 
  filter(!is.na(scientificname), !is.na(valid_name)) %>% # delete entries with NAs
  group_by(provided_sci_name) %>% 
  filter(provided_sci_name == scientificname) %>% # make sure to keep only rows that match the sci_name we supplied
  filter(n_distinct(valid_name, na.rm = T) > 1) %>% 
  filter(status == "accepted") %>% 
  group_by(provided_sci_name) %>% 
  summarise_at(vars(valid_name, kingdom, phylum, class, order, family, genus), .funs = unique)
```

Lets now add the WoRMS valid name to the dataset

```{r}
IUCN_spp_list <- IUCN_spp_list %>% 
  left_join(bind_rows(accepted_IUCN_worms_info,
                      valid_IUCN_worms_info),
            by = c("resolved_sci_name" = "provided_sci_name")
)

IUCN_spp_list %>% 
  filter(iucn_sci_name != valid_name)
```

```{r}
IUCN_spp_list <- IUCN_spp_list %>% 
  mutate(valid_sci_name = coalesce(valid_name, resolved_sci_name)) %>% 
  select(iucn_sci_name, resolved_sci_name,  valid_sci_name, everything(), -valid_name)
```

### fill taxonomic gaps

```{r}
get_taxonomy <- function (sci_name, db) {
  
  taxize::classification(sci_name, db, accepted = T, ask = F)[[1]] %>% 
    select(-id) %>% 
    mutate(rank = stringr::str_to_lower(rank)) %>% 
    filter(rank %in% c("genus","family",  "order", "class","phylum","kingdom")) %>% 
    spread(rank, name) %>% 
    mutate(sci_name = sci_name) %>% 
    janitor::clean_names() %>% 
    select(sci_name, genus, family, order, class, phylum, kingdom)
  
}
```

```{r}
taxonomy_gaps_IUCN <- IUCN_spp_list %>% 
  filter(is.na(kingdom)) %>% 
  select(valid_sci_name)
```

#### COL

```{r}
plan(multiprocess)

IUCN_taxonomy_col <- taxonomy_gaps_IUCN %>% 
  pull(valid_sci_name) %>% 
  furrr::future_map(safely(get_taxonomy), db = "col") %>% 
  transpose()

IUCN_taxonomy_col <- bind_rows(IUCN_taxonomy_col$result)
```

#### ITIS

```{r}
plan(multiprocess)

IUCN_taxonomy_itis <- taxonomy_gaps_IUCN %>% 
  filter(!valid_sci_name %in% IUCN_taxonomy_col$sci_name) %>% 
  pull(valid_sci_name) %>% 
  furrr::future_map(safely(get_taxonomy), db = "itis") %>% 
  transpose()

IUCN_taxonomy_itis <- bind_rows(IUCN_taxonomy_itis$result)
```

#### GBIF

```{r}
plan(multiprocess)

IUCN_taxonomy_gbif <- taxonomy_gaps_IUCN %>% 
  filter(!valid_sci_name %in% IUCN_taxonomy_col$sci_name) %>% 
  filter(!valid_sci_name %in% IUCN_taxonomy_itis$sci_name) %>% 
  pull(valid_sci_name) %>% 
  furrr::future_map(safely(get_taxonomy), db = "gbif") %>% 
  transpose()

IUCN_taxonomy_gbif <- bind_rows(IUCN_taxonomy_gbif$result)
```

```{r}
filled_taxonomy_gaps_IUCN <- bind_rows(IUCN_taxonomy_col, IUCN_taxonomy_itis, IUCN_taxonomy_gbif)

filled_taxonomy_gaps_IUCN <- filled_taxonomy_gaps_IUCN %>% 
  mutate(valid_sci_name = sci_name) %>% 
  group_by(valid_sci_name) %>% 
  summarise_at(vars(kingdom, phylum, class, order, family, genus), .funs = unique)
```

```{r}
IUCN_spp_list <- IUCN_spp_list %>% 
  left_join(filled_taxonomy_gaps_IUCN, by = c("valid_sci_name" = "valid_sci_name")) %>% 
  mutate(genus = coalesce(genus.x, genus.y),
         family = coalesce(family.x, family.y),
         order = coalesce(order.x, order.y),
         class = coalesce(class.x, class.y),
         phylum = coalesce(phylum.x, phylum.y),
         kingdom = coalesce(kingdom.x, kingdom.y)
         ) %>% 
  select(-contains(".x"), -contains(".y"))

IUCN_spp_list %>% filter(is.na(kingdom))
```

#### IUCN

```{r}
iucn_last_missing_taxonomy_spp_ids <- IUCN_spp_list %>% 
  filter(is.na(kingdom)) %>% 
  pull(iucn_id)

iucn_last_missing_taxonomy <- marine_spp %>% 
  filter(taxonid %in% iucn_last_missing_taxonomy_spp_ids) %>% 
  select(kingdom = kingdom_name, phylum = phylum_name, class = class_name, order = order_name, family = family_name, genus = genus_name, taxonid) %>% 
  mutate_at(vars(kingdom, phylum, class, order, family), stringr::str_to_title)

IUCN_spp_list <- IUCN_spp_list %>% 
  left_join(iucn_last_missing_taxonomy, by = c("iucn_id" = "taxonid")) %>% 
  mutate(genus = coalesce(genus.x, genus.y),
         family = coalesce(family.x, family.y),
         order = coalesce(order.x, order.y),
         class = coalesce(class.x, class.y),
         phylum = coalesce(phylum.x, phylum.y),
         kingdom = coalesce(kingdom.x, kingdom.y)
         ) %>% 
  select(-contains(".x"), -contains(".y")) 
```

## FAO

FAO makes available complete timeseries data of exploited species via FishStatJ. This is the most comprehensive list of exploited species but it is a pain to obtain. Accessed via Tyler Clavelle on December 11th 2018. 

```{r, eval = F}
load(file.path(DIR_GCS, "Biodiversity/data/raw/species-list/fao_fish.rda")) # this loads a dataframe called fish

FAO_capture_list <- fish %>% 
  filter(waters == 'Marine areas') %>% 
  distinct(common_name, scientific_name, fao_code = taxo_code) %>% 
  arrange(scientific_name)
```

This list contains records at multiple taxonomic levels: some are species, other are genus, other families. Lets deal with spp first

```{r, eval = F}
# Get species records
FAO_spp_list <- FAO_capture_list %>% 
  filter(!str_detect(fao_code, "X")) 
```

### Resolve scientific names

```{r}
FAO_resolved_names <- sources %>% 
  map(~ taxize::gnr_resolve(data_source_ids = .x, 
                            names = FAO_spp_list$scientific_name,
                            best_match_only = T, 
                            fields = c("all"), 
                            canonical = T ))

FAO_resolved_names_df <- FAO_resolved_names %>% 
  map(~ .x %>% 
        filter(!match_value %in% c("Could only match genus") & str_count(matched_name2, "\\w+") >= 2) %>% 
        select(user_supplied_name, matched_name2, taxon_id)) %>% 
  reduce(full_join, by = "user_supplied_name") %>% 
  set_names(c("user_supplied_name", "worms_sci_name", "worms_id", "fishbase_sci_name", "fishbase_id", "eol_sci_name", "eol_id", "col_sci_name", "col_id","gbif_sci_name", "gbif_id"))

FAO_resolved_names <- FAO_resolved_names_df %>% 
  mutate(resolved_scientific_name = case_when(
    !is.na(worms_sci_name) ~ worms_sci_name,
    !is.na(fishbase_sci_name) ~ fishbase_sci_name,
    !is.na(eol_sci_name) ~ eol_sci_name,
    !is.na(col_sci_name) ~ col_sci_name,
    TRUE ~ user_supplied_name
  )) %>% 
  select(user_supplied_name, resolved_scientific_name, everything()) 
```

```{r}
FAO_spp_list <- FAO_spp_list %>% 
  select(fao_sci_name = scientific_name, fao_id = fao_code)

FAO_spp_list <- FAO_spp_list %>%
  left_join(FAO_resolved_names %>% 
              select(user_supplied_name, resolved_sci_name = resolved_scientific_name), 
            by = c("fao_sci_name" = "user_supplied_name")) %>% 
  select(fao_sci_name, resolved_sci_name, everything())

FAO_spp_list <- FAO_spp_list %>% 
  mutate(resolved_sci_name = coalesce(resolved_sci_name, fao_sci_name)) 
```

```{r}
FAO_spp_list$resolved_sci_name[FAO_spp_list$fao_sci_name == "Pagrus auratus"] <- 'Chrysophrys auratus'
```

### validate names and get taxonomic info from WoRMS

```{r}
plan(multiprocess)

FAO_worms_info <- FAO_spp_list %>%
  pull(resolved_sci_name) %>% 
  furrr::future_map(safely(get_worms_info)) %>% 
  transpose()

FAO_worms_info <- FAO_worms_info$result %>% 
  compact() %>% 
  bind_rows()
```
Lets take those entries that have no NAs, where the name queried matches the name supplied, and when there are only one valid name per entry

```{r}
valid_FAO_worms_info <- FAO_worms_info %>% 
  filter(!is.na(scientificname), !is.na(valid_name)) %>% # delete entries with NAs
  group_by(provided_sci_name) %>% 
  filter(provided_sci_name == scientificname) %>% # make sure to keep only rows that match the sci_name we supplied
  filter(n_distinct(valid_name, na.rm = T) == 1) %>%  # get those for which there is only one valid name
  ungroup() %>% 
  group_by(provided_sci_name) %>% 
  summarise_at(vars(valid_name, kingdom, phylum, class, order, family, genus), .funs = unique)
```

Now, from those with multiple valid names per row, lets get only the row for which status == "accepted", which means that the supplied name was already valid

```{r}
accepted_FAO_worms_info <- FAO_worms_info %>% 
  filter(!is.na(scientificname), !is.na(valid_name)) %>% # delete entries with NAs
  group_by(provided_sci_name) %>% 
  filter(provided_sci_name == scientificname) %>% # make sure to keep only rows that match the sci_name we supplied
  filter(n_distinct(valid_name, na.rm = T) > 1) %>% 
  filter(status == "accepted") %>% 
  group_by(provided_sci_name) %>% 
  summarise_at(vars(valid_name, kingdom, phylum, class, order, family, genus), .funs = unique)
```

```{r}
FAO_spp_list <- FAO_spp_list %>% 
  left_join(bind_rows(accepted_FAO_worms_info,
                      valid_FAO_worms_info),
            by = c("resolved_sci_name" = "provided_sci_name")
)

FAO_spp_list <- FAO_spp_list %>% 
  mutate(valid_sci_name = coalesce(valid_name, resolved_sci_name)) %>% 
  select(fao_sci_name, resolved_sci_name,  valid_sci_name, everything(), -valid_name)

FAO_spp_list %>% filter(is.na(kingdom))
```

### fill taxonomic  gaps

```{r}
taxonomy_gaps_FAO <- FAO_spp_list %>% 
  filter(is.na(kingdom)) %>% 
  select(valid_sci_name)
```

#### COL

```{r}
plan(multiprocess)

FAO_taxonomy_col <- taxonomy_gaps_FAO %>% 
  pull(valid_sci_name) %>% 
  furrr::future_map(safely(get_taxonomy), db = "col") %>% 
  transpose()

FAO_taxonomy_col <- bind_rows(FAO_taxonomy_col$result)
```

#### ITIS

```{r}
plan(multiprocess)

FAO_taxonomy_itis <- taxonomy_gaps_FAO %>% 
  filter(!valid_sci_name %in% FAO_taxonomy_col$sci_name) %>% 
  pull(valid_sci_name) %>% 
  furrr::future_map(safely(get_taxonomy), db = "itis") %>% 
  transpose()

FAO_taxonomy_itis <- bind_rows(FAO_taxonomy_itis$result)
```

#### GBIF

```{r}
plan(multiprocess)

FAO_taxonomy_gbif <- taxonomy_gaps_FAO %>% 
  filter(!valid_sci_name %in% FAO_taxonomy_col$sci_name) %>% 
  filter(!valid_sci_name %in% FAO_taxonomy_itis$sci_name) %>% 
  pull(valid_sci_name) %>% 
  furrr::future_map(safely(get_taxonomy), db = "gbif") %>% 
  transpose()

FAO_taxonomy_gbif <- bind_rows(FAO_taxonomy_gbif$result)
```

```{r}
filled_taxonomy_gaps_FAO <- bind_rows(FAO_taxonomy_col, FAO_taxonomy_itis, FAO_taxonomy_gbif)

filled_taxonomy_gaps_FAO <- filled_taxonomy_gaps_FAO %>% 
  mutate(valid_sci_name = sci_name) %>% 
  group_by(valid_sci_name) %>% 
  summarise_at(vars(kingdom, phylum, class, order, family, genus), .funs = unique)
```

```{r}
FAO_spp_list <- FAO_spp_list %>% 
  left_join(filled_taxonomy_gaps_FAO, by = c("valid_sci_name" = "valid_sci_name")) %>% 
  mutate(genus = coalesce(genus.x, genus.y),
         family = coalesce(family.x, family.y),
         order = coalesce(order.x, order.y),
         class = coalesce(class.x, class.y),
         phylum = coalesce(phylum.x, phylum.y),
         kingdom = coalesce(kingdom.x, kingdom.y)
         ) %>% 
  select(-contains(".x"), -contains(".y")) 

FAO_spp_list %>% filter(is.na(kingdom))
```

### nei

```{r}
FAO_nei_resolved <- FAO_capture_list %>% 
  filter(str_detect(fao_code, "X")) %>% 
  mutate(fao_nei = stringr::word(scientific_name, 1)) %>% 
  pull(fao_nei) %>% 
  taxize::gnr_resolve(data_source_ids = 9, 
                      best_match_only = T, 
                      fields = c("all"), 
                      canonical = T)

FAO_nei_genus <- FAO_nei_resolved %>%
  mutate(matched_rank = str_extract(classification_path_ranks, "([^|]+$)")) %>% 
  filter(matched_rank == "genus") %>% 
  select(nei_genus = matched_name2) %>% 
  write_csv(file.path(DIR_GCS, "Biodiversity/data/preprocessed/FAO_nei_genus.csv"))
```

## RAM

```{r, eval = F}
load(file.path(DIR_GCS, "Biodiversity/data/raw/species-list/ram.rda")) # this loads a dataframe called ram

RAM_spp_list <- ram$taxonomy[[1]] %>% 
  filter(!is.na(species)) %>% 
  select(ram_sci_name = scientificname, ram_id = tsn, genus, family, order = ordername, class = classname, phylum, kingdom)
```

### Resolve scientific names

```{r}
RAM_resolved_names <- sources %>% 
  map(~ taxize::gnr_resolve(data_source_ids = .x, 
                            names = RAM_spp_list$ram_sci_name,
                            best_match_only = T, 
                            fields = c("all"), 
                            canonical = T ))

RAM_resolved_names_df <- RAM_resolved_names %>% 
  map(~ .x %>% 
        filter(!match_value %in% c("Could only match genus") & str_count(matched_name2, "\\w+") >= 2) %>% 
        select(user_supplied_name, matched_name2, taxon_id)) %>% 
  reduce(full_join, by = "user_supplied_name") %>% 
  set_names(c("user_supplied_name", "worms_sci_name", "worms_id", "fishbase_sci_name", "fishbase_id", "eol_sci_name", "eol_id", "col_sci_name", "col_id","gbif_sci_name", "gbif_id"))

RAM_resolved_names <- RAM_resolved_names_df %>% 
  mutate(resolved_scientific_name = case_when(
    !is.na(worms_sci_name) ~ worms_sci_name,
    !is.na(fishbase_sci_name) ~ fishbase_sci_name,
    !is.na(eol_sci_name) ~ eol_sci_name,
    !is.na(col_sci_name) ~ col_sci_name,
    TRUE ~ user_supplied_name
  )) %>% 
  select(user_supplied_name, resolved_scientific_name, everything()) 
```

```{r}
RAM_spp_list <- RAM_spp_list %>% 
  select(ram_sci_name, ram_id)

RAM_spp_list <- RAM_spp_list %>%
  left_join(RAM_resolved_names %>% 
              select(user_supplied_name, resolved_sci_name = resolved_scientific_name), 
            by = c("ram_sci_name" = "user_supplied_name")) %>% 
  select(ram_sci_name, resolved_sci_name, everything())

RAM_spp_list <- RAM_spp_list %>% 
  mutate(resolved_sci_name = coalesce(resolved_sci_name, ram_sci_name)) 
```

```{r}
RAM_spp_list$resolved_sci_name[RAM_spp_list$ram_sci_name == "Pagrus auratus"] <- 'Chrysophrys auratus'
```

### validate names and get taxonomic info from WoRMS

```{r}
plan(multiprocess)

RAM_worms_info <- RAM_spp_list %>%
  pull(resolved_sci_name) %>% 
  furrr::future_map(safely(get_worms_info)) %>% 
  transpose()

RAM_worms_info <- RAM_worms_info$result %>% 
  compact() %>% 
  bind_rows()
```
Lets take those entries that have no NAs, where the name queried matches the name supplied, and when there are only one valid name per entry

```{r}
valid_RAM_worms_info <- RAM_worms_info %>% 
  filter(!is.na(scientificname), !is.na(valid_name)) %>% # delete entries with NAs
  group_by(provided_sci_name) %>% 
  filter(provided_sci_name == scientificname) %>% # make sure to keep only rows that match the sci_name we supplied
  filter(n_distinct(valid_name, na.rm = T) == 1) %>%  # get those for which there is only one valid name
  ungroup() %>% 
  group_by(provided_sci_name) %>% 
  summarise_at(vars(valid_name, kingdom, phylum, class, order, family, genus), .funs = unique)
```

Now, from those with multiple valid names per row, lets get only the row for which status == "accepted", which means that the supplied name was already valid

```{r}
accepted_RAM_worms_info <- RAM_worms_info %>% 
  filter(!is.na(scientificname), !is.na(valid_name)) %>% # delete entries with NAs
  group_by(provided_sci_name) %>% 
  filter(provided_sci_name == scientificname) %>% # make sure to keep only rows that match the sci_name we supplied
  filter(n_distinct(valid_name, na.rm = T) > 1) %>% 
  filter(status == "accepted") %>% 
  group_by(provided_sci_name) %>% 
  summarise_at(vars(valid_name, kingdom, phylum, class, order, family, genus), .funs = unique)
```

```{r}
RAM_spp_list <- RAM_spp_list %>% 
  left_join(bind_rows(accepted_RAM_worms_info,
                      valid_RAM_worms_info),
            by = c("resolved_sci_name" = "provided_sci_name")
)

RAM_spp_list <- RAM_spp_list %>% 
  mutate(valid_sci_name = coalesce(valid_name, resolved_sci_name)) %>% 
  select(ram_sci_name, resolved_sci_name,  valid_sci_name, everything(), -valid_name)

RAM_spp_list %>% filter(is.na(kingdom))
```

### nei

```{r}
RAM_nei_genus <- ram$taxonomy[[1]] %>% 
  filter(is.na(species)) %>% 
  select(nei_genus = genus) %>% 
  write_csv(file.path(DIR_GCS, "Biodiversity/data/preprocessed/RAM_nei_genus.csv"))
```

## Watson

```{r}
Watson_data <- read_csv(file.path(DIR_GCS_data, 'Watson-catch-data-v4.0/taxa_codes.csv'))

Watson_spp_list <- Watson_data %>% 
  filter(str_count(TaxonName, '\\w+') == 2) %>% 
  filter(!TaxonName %in% c("Marine animals", "Shrimps, prawns", "Homaridae/Palinuridae"))
```

### Resolve scientific names

```{r}
Watson_resolved_names <- sources %>% 
  map(~ taxize::gnr_resolve(data_source_ids = .x, 
                            names = Watson_spp_list$TaxonName,
                            best_match_only = T, 
                            fields = c("all"), 
                            canonical = T ))

Watson_resolved_names_df <- Watson_resolved_names %>% 
  map(~ .x %>% 
        filter(!match_value %in% c("Could only match genus") & str_count(matched_name2, "\\w+") >= 2) %>% 
        select(user_supplied_name, matched_name2, taxon_id)) %>% 
  reduce(full_join, by = "user_supplied_name") %>% 
  set_names(c("user_supplied_name", "worms_sci_name", "worms_id", "fishbase_sci_name", "fishbase_id", "eol_sci_name", "eol_id", "col_sci_name", "col_id","gbif_sci_name", "gbif_id"))

Watson_resolved_names <- Watson_resolved_names_df %>% 
  mutate(resolved_scientific_name = case_when(
    !is.na(worms_sci_name) ~ worms_sci_name,
    !is.na(fishbase_sci_name) ~ fishbase_sci_name,
    !is.na(eol_sci_name) ~ eol_sci_name,
    !is.na(col_sci_name) ~ col_sci_name,
    TRUE ~ user_supplied_name
  )) %>% 
  select(user_supplied_name, resolved_scientific_name, everything()) 
```

```{r}
Watson_spp_list <- Watson_spp_list %>% 
  select(watson_sci_name = TaxonName, watson_id = TaxonKey)

Watson_spp_list <- Watson_spp_list %>%
  left_join(Watson_resolved_names %>% 
              distinct(user_supplied_name, resolved_sci_name = resolved_scientific_name), 
            by = c("watson_sci_name" = "user_supplied_name")) %>% 
  select(watson_sci_name, resolved_sci_name, everything())

Watson_spp_list <- Watson_spp_list %>% 
  mutate(resolved_sci_name = coalesce(resolved_sci_name, watson_sci_name)) 
```

```{r}
Watson_spp_list$resolved_sci_name[Watson_spp_list$watson_sci_name == "Pagrus auratus"] <- 'Chrysophrys auratus'
```

### validate names and get taxonomic info from WoRMS

```{r}
plan(multiprocess)

Watson_worms_info <- Watson_spp_list %>%
  pull(resolved_sci_name) %>% 
  furrr::future_map(safely(get_worms_info)) %>% 
  transpose()

Watson_worms_info <- Watson_worms_info$result %>% 
  compact() %>% 
  bind_rows()
```

Lets take those entries that have no NAs, where the name queried matches the name supplied, and when there are only one valid name per entry

```{r}
valid_Watson_worms_info <- Watson_worms_info %>% 
  filter(!is.na(scientificname), !is.na(valid_name)) %>% # delete entries with NAs
  group_by(provided_sci_name) %>% 
  filter(provided_sci_name == scientificname) %>% # make sure to keep only rows that match the sci_name we supplied
  filter(n_distinct(valid_name, na.rm = T) == 1) %>%  # get those for which there is only one valid name
  ungroup() %>% 
  group_by(provided_sci_name) %>% 
  summarise_at(vars(valid_name, kingdom, phylum, class, order, family, genus), .funs = unique)
```

Now, from those with multiple valid names per row, lets get only the row for which status == "accepted", which means that the supplied name was already valid

```{r}
accepted_Watson_worms_info <- Watson_worms_info %>% 
  filter(!is.na(scientificname), !is.na(valid_name)) %>% # delete entries with NAs
  group_by(provided_sci_name) %>% 
  filter(provided_sci_name == scientificname) %>% # make sure to keep only rows that match the sci_name we supplied
  filter(n_distinct(valid_name, na.rm = T) > 1) %>% 
  filter(status == "accepted") %>% 
  group_by(provided_sci_name) %>% 
  summarise_at(vars(valid_name, kingdom, phylum, class, order, family, genus), .funs = unique)
```

```{r}
Watson_spp_list <- Watson_spp_list %>% 
  left_join(bind_rows(accepted_Watson_worms_info,
                      valid_Watson_worms_info),
            by = c("resolved_sci_name" = "provided_sci_name")
)

Watson_spp_list <- Watson_spp_list %>% 
  mutate(valid_sci_name = coalesce(valid_name, resolved_sci_name)) %>% 
  select(watson_sci_name, resolved_sci_name,  valid_sci_name, everything(), -valid_name)

Watson_spp_list %>% filter(is.na(kingdom))
```

### nei

```{r}
Watson_resolved_nei <- Watson_data %>% 
  filter(str_count(TaxonName, '\\w+') == 1) %>% 
  pull(TaxonName) %>% 
  taxize::gnr_resolve(data_source_ids = 9, 
                      best_match_only = T, 
                      fields = c("all"), 
                      canonical = T)

Watson_nei_genus <- Watson_resolved_nei %>%
  mutate(matched_rank = str_extract(classification_path_ranks, "([^|]+$)")) %>% 
  filter(matched_rank == "genus") %>% 
  select(nei_genus = matched_name2) %>% 
  write_csv(file.path(DIR_GCS, "Biodiversity/data/preprocessed/Watson_nei_genus.csv"))
```


# Master list

Now we need to combine all 3 lists into a final master list of all species to be included in the analysis. When doing this, we need to be careful to avoid duplication of species. 

### exploited species

```{r}
RAM_spp_list %>% 
  filter(!valid_sci_name %in% FAO_spp_list$valid_sci_name)
```

```{r}
exploited_spp_list <- bind_rows(FAO_spp_list %>% 
                                  mutate(list = "FAO") %>% 
                                  select(valid_sci_name, list, list_id = fao_id, list_sci_name = fao_sci_name, everything()),
                                RAM_spp_list %>% 
                                  filter(!valid_sci_name %in% FAO_spp_list$valid_sci_name) %>%
                                  mutate(list = "RAM") %>% 
                                  select(valid_sci_name, list, list_id = ram_id, list_sci_name = ram_sci_name, everything()))
```

#### add Watson

```{r}
Watson_spp_list %>% 
  filter(!valid_sci_name %in% exploited_spp_list$valid_sci_name)
```

```{r}
exploited_spp_list <- bind_rows(exploited_spp_list,
                                Watson_spp_list %>% 
                                  filter(!valid_sci_name %in% exploited_spp_list$valid_sci_name) %>% 
                                  mutate(list = "Watson",
                                         watson_id = as.character(watson_id)) %>% 
                                  select(valid_sci_name, list, list_id = watson_id, list_sci_name = watson_sci_name, everything()))
```

```{r}
exploited_spp_list %>% 
  janitor::get_dupes(valid_sci_name) 
```

### Union of exploited and IUCN

```{r}
complete_spp_list <- bind_rows(exploited_spp_list %>% 
                                 mutate(exploited = T,
                                        threat_group = "fishing"),
                               IUCN_spp_list %>% 
                                 filter(!valid_sci_name %in% exploited_spp_list$valid_sci_name) %>%
                                 mutate(iucn_id = as.character(iucn_id),
                                        list = "IUCN",
                                        exploited = F) %>% 
                                 select(valid_sci_name, list, list_id = iucn_id, list_sci_name = iucn_sci_name, everything(),
                                        exploited, threat_group))
```

### trim characters

```{r}
complete_spp_list <- complete_spp_list %>% 
  map_df(~ if(class(.x) == "character"){stringr::str_trim(.x)} else (.x))
```

### Fill IUCN status

```{r}
complete_spp_list <- complete_spp_list %>% 
  left_join(IUCN_spp_list %>% 
              distinct(iucn_category, iucn_sci_name), 
            by = c("list_sci_name" = "iucn_sci_name")) %>% 
  mutate(iucn_category = coalesce(iucn_category.x, iucn_category.y)) %>% 
  select(-contains(".x"), -contains(".y"))
```

```{r}
plan(multiprocess)

missing_iucn <- complete_spp_list %>% 
  filter(is.na(iucn_category)) %>% 
  pull(valid_sci_name) %>% 
  furrr::future_map(safely(rl_search)) %>% 
  transpose()

missing_iucn <- missing_iucn$result %>% 
  map("result") %>% 
  compact() %>% 
  map(~ .x %>% 
        select(scientific_name, category)) %>% 
  bind_rows()
```

```{r}
complete_spp_list <- complete_spp_list %>% 
  left_join(missing_iucn %>% 
              distinct(scientific_name, category), by = c("valid_sci_name" = "scientific_name")) %>% 
  mutate(iucn_category = ifelse(!is.na(iucn_category), iucn_category, category)) %>% 
  select(-category)
```

```{r}
complete_spp_list <- complete_spp_list %>% 
  filter(!iucn_category %in% c("EX", "EW")) %>% 
  mutate(iucn_category = if_else(iucn_category %in% c("LR/lc"), "LC", 
                                 if_else(iucn_category == "LR/nt", 'NT', iucn_category)))

complete_spp_list %>% 
  group_by(iucn_category) %>% 
  summarize(n_spp = n()) %>% 
  arrange(desc(n_spp))
```

```{r}
complete_spp_list %>% filter(is.na(kingdom))
```

```{r}
complete_spp_list %>% 
  janitor::get_dupes(valid_sci_name) %>% 
  summarize(n_distinct(valid_sci_name))
```


```{r}
saveRDS(complete_spp_list, 
        file.path(DIR_GCS, "Biodiversity/data/preprocessed/complete_spp_list_with_taxonomy.rds"))
```

# Final Inspection

```{r}
complete_spp_list <- read_rds(file.path(DIR_GCS, "Biodiversity/data/preprocessed/complete_spp_list_with_taxonomy.rds"))
```

```{r}
complete_spp_list %>% 
  group_by(kingdom) %>% 
  summarize(n_spp = n_distinct(valid_sci_name))
```


```{r}
complete_spp_list %>% 
  group_by(iucn_category) %>% 
  summarize(n_spp = n_distinct(valid_sci_name)) %>% 
  arrange(desc(n_spp))
```

## exploited + fishing 

```{r}
complete_spp_list %>% 
  filter(threat_group %in% c("fishing")) %>% 
  group_by(exploited) %>% 
  summarize(n_spp = n_distinct(valid_sci_name))
```

```{r}
complete_spp_list %>% 
  filter(threat_group %in% c("fishing")) %>% 
  group_by(class) %>% 
  summarize(n_spp = n_distinct(valid_sci_name)) %>% 
  arrange(desc(n_spp))
```

## Other threats

```{r}
complete_spp_list %>% 
  group_by(threat_group) %>% 
  summarize(n_spp = n_distinct(valid_sci_name),
            n_threatened = sum(iucn_category %in% c("CR", "EN", "VU")))
```

### + aquaculture

```{r}
complete_spp_list %>% 
  filter(iucn_category %in% c("VU", "CR", "EN")) %>% 
  filter(threat_group %in% c("aquaculture")) %>% 
  group_by(class) %>% 
  summarize(n_spp = n_distinct(valid_sci_name)) %>% 
  arrange(desc(n_spp))
```

We would get more birds (mostly shorebirds) and plants

### + oil and gas

```{r}
complete_spp_list %>% 
  filter(iucn_category %in% c("VU", "CR", "EN")) %>% 
  filter(threat_group %in% c("oil_and_gas")) %>% 
  group_by(class) %>% 
  summarize(n_spp = n_distinct(valid_sci_name)) %>% 
  arrange(desc(n_spp))
```

### + mining

```{r}
complete_spp_list %>% 
  filter(iucn_category %in% c("VU", "CR", "EN")) %>% 
  filter(threat_group %in% c("mining")) %>% 
  group_by(class) %>% 
  summarize(n_spp = n_distinct(valid_sci_name)) %>% 
  arrange(desc(n_spp))
```


### + logging

```{r}
complete_spp_list %>% 
  filter(iucn_category %in% c("VU", "CR", "EN")) %>% 
  filter(threat_group %in% c("logging")) %>% 
  group_by(class) %>% 
  summarize(n_spp = n_distinct(valid_sci_name)) %>% 
  arrange(desc(n_spp))
```

### + development

```{r}
complete_spp_list %>% 
  filter(iucn_category %in% c("VU", "CR", "EN")) %>% 
  filter(threat_group %in% c("development")) %>% 
  group_by(class) %>% 
  summarize(n_spp = n_distinct(valid_sci_name)) %>% 
  arrange(desc(n_spp))
```

